### 2015-10-05 paper.tex
Ejemplo simplificado

\documentclass[a4paper,twoside]{article}

\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}
\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.

\subfigtopskip=0pt
\subfigcapskip=0pt
\subfigbottomskip=0pt

\begin{document}

\title{Authors' Instructions  \subtitle{Preparation of Camera-Ready Contributions to SCITEPRESS Proceedings} }

\author{\authorname{First Author Name\sup{1}, Second Author Name\sup{1} and Third Author Name\sup{2}}
\affiliation{\sup{1}Institute of Problem Solving, XYZ University, My Street, MyTown, MyCountry}
\affiliation{\sup{2}Department of Computing, Main University, MySecondTown, MyCountry}
\email{\{f\_author, s\_author\}@ips.xyz.edu, t\_author@dc.mu.edu}
}

\keywords{The paper must have at least one keyword. The text must be set to 9-point font size and without the use of bold or italic font style. For more than one keyword, please use a comma as a separator. Keywords must be titlecased.}

\abstract{The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words. The text must be set to 9-point font size.}

\onecolumn \maketitle \normalsize \vfill

\section{\uppercase{Introduction}}
\label{sec:introduction}

\noindent Your paper will be part of the conference proceedings
therefore we ask that authors follow the guidelines explained in
this example in order to achieve the highest quality possible
\cite{Smith98}.

Be advised that papers in a technically unsuitable form will be
returned for retyping. After returned the manuscript must be
appropriately modified.

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}

\rcuda~\cite{tonithesis,toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
as if they were local.
These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy~\cite{energy14} requirements.
rCUDA is structured following a client-server distributed
architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, \rcuda\ manages separate device contexts for
each client application.

The \rcuda\ 5.0 client exposes the same interface as the regular NVIDIA
CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
Therefore, applications are not aware of the fact that they are being executed
on top of a virtualization layer.
With the aim to be updated with new GPU programming models, \rcuda also supports
directive-based models such as OmpSS~\cite{repara15} and OpenACC~\{cluster15}.

The integration of remote GPGPU virtualization with global
resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
technology, making accelerator-enabled clusters more flexible and
energy efficient.

\subsection{Manuscript Setup}

\subsection{Page Setup}

\section*{\uppercase{Acknowledgements}}

\bibliographystyle{apalike}
{\small
\bibliography{example}}

\end{document}

### 2015-10-05 paper.tex
Tipografia de rCUDA cambiada

@@
 
-\rcuda~\cite{tonithesis,toniparco} is a middleware that enables transparent access
+{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
 to any NVIDIA GPU device present in a cluster from all compute
@@
 To support a concurrent scenario where GPUs are shared between
-processes\slash nodes, \rcuda\ manages separate device contexts for
+processes\slash nodes, {rCUDA} manages separate device contexts for
 each client application.
 
-The \rcuda\ 5.0 client exposes the same interface as the regular NVIDIA
+The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
 CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
@@
 on top of a virtualization layer.
-With the aim to be updated with new GPU programming models, \rcuda also supports
+With the aim to be updated with new GPU programming models, {rCUDA} also supports
 directive-based models such as OmpSS~\cite{repara15} and OpenACC~\{cluster15}.

### 2015-10-05 paper.tex
Cambiado el nombre de la bibliografia en paper.tex
Cambios en bibliografia para acentos y letra ñ

@@
 With the aim to be updated with new GPU programming models, {rCUDA} also supports
-directive-based models such as OmpSS~\cite{repara15} and OpenACC~\{cluster15}.
+directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.
 
@@
 {\small
-\bibliography{example}}
+\bibliography{paper}}
 

### 2015-10-05 paper.tex
title and authors

--- 2015-10-05 paper.tex
@@
 
-\title{Authors' Instructions  \subtitle{Preparation of Camera-Ready Contributions to SCITEPRESS Proceedings} }
-
-\author{\authorname{First Author Name\sup{1}, Second Author Name\sup{1} and Third Author Name\sup{2}}
-\affiliation{\sup{1}Institute of Problem Solving, XYZ University, My Street, MyTown, MyCountry}
-\affiliation{\sup{2}Department of Computing, Main University, MySecondTown, MyCountry}
-\email{\{f\_author, s\_author\}@ips.xyz.edu, t\_author@dc.mu.edu}
-}
+\title{GPU Virtualization in the Cloud (a mejorar el t\'itulo!!!)}
+\author{\authorname{Sergio Iserte, Francisco J. Clemente-Castell\'o, Adri\'an Castell\'o,\\Rafael Mayo and Enrique S. Quintana-Ort\'i}
+\affiliation{Department of Computer Science and Engineering\\Universitat Jaume I - Castell\'o de la Plana, Spain}
+\email{\{siserte, fclement, adcastel, mayo, quintana\}@uji.es}}
 
@@
 
-\section{\uppercase{Background}}
-\label{sec:background}
-\subsection{The rCUDA Framework}
-\label{sec:rcuda}
+\section{\uppercase{Manuscript Preparation}}
 
-\rcuda~\cite{tonithesis,toniparco} is a middleware that enables transparent access
-to any NVIDIA GPU device present in a cluster from all compute
-nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
-as if they were local.
-These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
-resource, space, and energy~\cite{energy14} requirements.
-rCUDA is structured following a client-server distributed
-architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
-acceleration services is executed, providing a transparent replacement for the
-native CUDA libraries. Furthermore, the server middleware is executed in the
-cluster nodes from which the actual GPUs provide the requested GPGPU service.
-To support a concurrent scenario where GPUs are shared between
-processes\slash nodes, \rcuda\ manages separate device contexts for
-each client application.
+\noindent We strongly encourage authors to use this document for the
+preparation of the camera-ready. Please follow the instructions
+closely in order to make the volume look as uniform as possible
+\cite{Moore99}.
 
-The \rcuda\ 5.0 client exposes the same interface as the regular NVIDIA
-CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
-APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
-Therefore, applications are not aware of the fact that they are being executed
-on top of a virtualization layer.
-With the aim to be updated with new GPU programming models, \rcuda also supports
-directive-based models such as OmpSS~\cite{repara15} and OpenACC~\{cluster15}.
+Please remember that all the papers must be in English and without
+orthographic errors.
 
-The integration of remote GPGPU virtualization with global
-resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
-technology, making accelerator-enabled clusters more flexible and
-energy efficient.
+Do not add any text to the headers (do not set running heads) and
+footers, not even page numbers, because text will be added
+electronically.
+
+For a best viewing experience the used font must be Times New
+Roman, except on special occasions, such as program code
+\ref{subsubsec:program_code}.
 
@@
 
+\noindent The template is composed by a set of 7 files, in the
+following 2 groups:\\
+\noindent {\bf Group 1.} To format your paper you will need to copy
+into your working directory, but NOT edit, the following 4 files:
+\begin{verbatim}
+  - apalike.bst
+  - apalike.sty
+  - article.cls
+  - scitepress.sty
+\end{verbatim}
+
+\noindent {\bf Group 2.} Additionally, you may wish to copy and edit
+the following 3 example files:
+\begin{verbatim}
+  - example.bib
+  - example.tex
+  - scitepress.eps
+\end{verbatim}
+
 \subsection{Page Setup}
+
+The paper size must be set to A4 (210x297 mm). The document
+margins must be the following:
+
+\begin{itemize}
+    \item Top: 3,3 cm;
+    \item Bottom: 4,2 cm;
+    \item Left: 2,6 cm;
+    \item Right: 2,6 cm.
+\end{itemize}
+
+It is advisable to keep all the given values because any text or
+material outside the aforementioned margins will not be printed.
+
+\subsection{First Section}
+
+This section must be in one column.
+
+\vfill
+\subsubsection{Title and Subtitle}
+
+Use the command \textit{$\backslash$title} and follow the given structure in "example.tex". The title and subtitle must be with initial letters
+capitalized (titlecased).
+If no subtitle is required, please remove the corresponding \textit{$\backslash$subtitle} command.
+In the title or subtitle, words like "is", "or", "then", etc.
+should not be capitalized unless they are the first word of the subtitle.
+No formulas or special characters of any form or language are allowed in the title or subtitle.
+
+\subsubsection{Authors and Affiliations}
+
+Use the command \textit{$\backslash$author} and follow the given structure in "example.tex".
+
+\subsubsection{Keywords}
+
+Use the command \textit{$\backslash$keywords} and follow the given structure in "example.tex".
+Each paper must have at least one keyword.
+If more than one is specified, please use a comma as a separator.
+The sentence must end with a period.
+
+\subsubsection{Abstract}
+
+Use the command \textit{$\backslash$abstract} and follow the given structure in "example.tex".
+Each paper must have an abstract up to 200 words. The sentence
+must end with a period.
+
+\subsection{Second Section}
+
+Files "example.tex" and "example.bib" show how to create a paper
+with a corresponding list of references.
+
+This section must be in two columns.
+
+Each column must be 7,5-centimeter wide with a column spacing
+of 0,8-centimeter.
+
+The section text must be set to 10-point.
+
+Section, subsection and sub-subsection first paragraph should not
+have the first line indent.
+
+To remove the paragraph indentation (only necessary for the
+sections), use the command \textit{$\backslash$noindent} before the
+paragraph first word.
+
+If you use other style files (.sty) you MUST include them in the
+final manuscript zip file.
+
+\subsubsection{Section Titles}
+
+The heading of a section title should be in all-capitals.
+
+Example: \textit{$\backslash$section\{FIRST TITLE\}}
+
+\vfill
+\subsubsection{Subsection Titles}
+
+The heading of a subsection title must be with initial letters
+capitalized (titlecased).
+
+Words like "is", "or", "then", etc. should not be capitalized unless
+they are the first word of the subsection title.
+
+Example: \textit{$\backslash$subsection\{First Subtitle\}}
+
+\subsubsection{Sub-Subsection Titles}
+
+The heading of a sub subsection title should be with initial letters
+capitalized (titlecased).
+
+Words like "is", "or", "then", etc should not be capitalized unless
+they are the first word of the sub subsection title.
+
+Example: \textit{$\backslash$subsubsection\{First Subsubtitle\}}
+
+\subsubsection{Tables}
+
+Tables must appear inside the designated margins or they may span
+the two columns.
+
+Tables in two columns must be positioned at the top or bottom of the
+page within the given margins. To span a table in two columns please add an asterisk (*) to the table \textit{begin} and \textit{end} command.
+
+Example: \textit{$\backslash$begin\{table*\}}
+
+\hspace*{1.5cm}\textit{$\backslash$end\{table*\}}\\
+
+Tables should be centered and should always have a caption
+positioned above it. The font size to use is 9-point. No bold or
+italic font style should be used.
+
+The final sentence of a caption should end with a period.
+
+\begin{table}[h]
+\caption{This caption has one line so it is
+centered.}\label{tab:example1} \centering
+\begin{tabular}{|c|c|}
+  \hline
+  Example column 1 & Example column 2 \\
+  \hline
+  Example text 1 & Example text 2 \\
+  \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[h]
+\caption{This caption has more than one line so it has to be
+justified.}\label{tab:example2} \centering
+\begin{tabular}{|c|c|}
+  \hline
+  Example column 1 & Example column 2 \\
+  \hline
+  Example text 1 & Example text 2 \\
+  \hline
+\end{tabular}
+\end{table}
+
+Please note that the word "Table" is spelled out.
+
+\subsubsection{Figures}
+
+Please produce your figures electronically, and integrate them into
+your document and zip file.
+
+Check that in line drawings, lines are not interrupted and have a
+constant width. Grids and details within the figures must be clearly
+readable and may not be written one on top of the other.
+
+Figure resolution should be at least 300 dpi.
+
+Figures must appear inside the designated margins or they may span
+the two columns.
+
+Figures in two columns must be positioned at the top or bottom of
+the page within the given margins. To span a figure in two columns please add an asterisk (*) to the figure \textit{begin} and \textit{end} command.
+
+Example: \textit{$\backslash$begin\{figure*\}}
+
+\hspace*{1.5cm}\textit{$\backslash$end\{figure*\}}
+
+Figures should be centered and should always have a caption
+positioned under it. The font size to use is 9-point. No bold or
+italic font style should be used.
+
+\begin{figure}[!h]
+  %\vspace{-0.2cm}
+  \centering
+   {\epsfig{file = SCITEPRESS.eps, width = 5.5cm}}
+  \caption{This caption has one line so it is centered.}
+  \label{fig:example1}
+ \end{figure}
+
+\begin{figure}[!h]
+  \vspace{-0.2cm}
+  \centering
+   {\epsfig{file = SCITEPRESS.eps, width = 5.5cm}}
+  \caption{This caption has more than one line so it has to be justified.}
+  \label{fig:example2}
+  \vspace{-0.1cm}
+\end{figure}
+
+The final sentence of a caption should end with a period.
+
+Please note that the word "Figure" is spelled out.
+
+\subsubsection{Equations}
+
+Equations should be placed on a separate line, numbered and
+centered.\\The numbers accorded to equations should appear in
+consecutive order inside each section or within the contribution,
+with the number enclosed in brackets and justified to the right,
+starting with the number 1.
+
+Example:
+
+\begin{equation}\label{eq1}
+    a=b+c
+\end{equation}
+
+\subsubsection{Program Code}\label{subsubsec:program_code}
+
+Program listing or program commands in text should be set in
+typewriter form such as Courier New.
+
+Example of a Computer Program in Pascal:
+
+\begin{small}
+\begin{verbatim}
+ Begin
+     Writeln('Hello World!!');
+ End.
+\end{verbatim}
+\end{small}
+
+The text must be aligned to the left and in 9-point type.
+
+\vfill
+\subsubsection{Reference Text and Citations}
+
+References and citations should follow the Harvard (Author, date)
+System Convention (see the References section in the compiled
+manuscript). As example you may consider the citation
+\cite{Smith98}. Besides that, all references should be cited in the
+text. No numbers with or without brackets should be used to list the
+references.
+
+References should be set to 9-point. Citations should be 10-point
+font size.
+
+You may check the structure of "example.bib" before constructing the
+references.
+
+For more instructions about the references and citations usage
+please see the appropriate link at the conference website.
+
+\section{\uppercase{Copyright Form}}
+
+\noindent For the mutual benefit and protection of Authors and
+Publishers, it is necessary that Authors provide formal written
+Consent to Publish and Transfer of Copyright before publication of
+the Book. The signed Consent ensures that the publisher has the
+Author's authorization to publish the Contribution.
+
+The copyright form is located on the authors' reserved area.
+
+The form should be completed and signed by one author on
+behalf of all the other authors.
+
+\section{\uppercase{Conclusions}}
+\label{sec:conclusion}
+
+\noindent Please note that ONLY the files required to compile your paper should be submitted. Previous versions or examples MUST be removed from the compilation directory before submission.
+
+We hope you find the information in this template useful in the preparation of your submission.
 
@@
 
+\noindent If any, should be placed before the references section
+without numbering. To do so please use the following command:
+\textit{$\backslash$section*\{ACKNOWLEDGEMENTS\}}
+
+\vfill
 \bibliographystyle{apalike}
@@
 
+\section*{\uppercase{Appendix}}
+
+\noindent If any, the appendix should appear directly after the
+references without numbering, and not on a new page. To do so please use the following command:
+\textit{$\backslash$section*\{APPENDIX\}}
+
+\vfill
 \end{document}

### 2015-10-05 paper.tex
intro

@@
 
-\keywords{The paper must have at least one keyword.
-The text must be set to 9-point font size and without the use of bold or italic font style.
-For more than one keyword, please use a comma as a separator.
-Keywords must be titlecased.}
+\keywords{Cloud Computing, GPU Virtualization, Resource Management}
 
-\abstract{The abstract should summarize the contents of the paper and should contain at least 70 and at most 200 words. The text must be set to 9-point font size.}
+\abstract{The popularity of the GPGPU computation has led several cloud vendors to provide instances of virtual machines with GPUs. 
+So that, the guest hosts must be equipped with GPUs which could be barely utilized if no GPU-enabled VM  is running in the host.
+The solution presented here is based on GPU virtualization and shareability in order to reach an equilibrium between supply and demand of accelerators from the users. 
+Hence, we propose to access any GPU from any VM getting rid of the necessity of having physical GPUs in the guest host. 
+Finally, our project goes further and it takes into account that by making logical partitions of the GPU memory we could address each partition as if it was an independent GPU device.}
 
@@
 
-\noindent Your paper will be part of the conference proceedings
-therefore we ask that authors follow the guidelines explained in
-this example in order to achieve the highest quality possible
-\cite{Smith98}.
-
-Be advised that papers in a technically unsuitable form will be
-returned for retyping. After returned the manuscript must be
-appropriately modified.
+\noindent 
+Nowadays, to provide a virtual machine with GPUs means
+that the guest host must be equipped with at least 1 GPU, and
+once it has been assigned to an instance, no other VM can use
+it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
+the conservative, to have a lot of hosts with GPUs to be
+ensured that you are likely to satisfy all the requests; or a
+more daring approach, where you will count with some GPU-
+enabled nodes hoping not to arrive a large burst of requests
+so the cluster would run out of GPU resources. Our solutions
+is based on GPU virtualization and sharing resources in order
+to reach a fair balance between supply and demand.
+Several attempts of achieving what we are pursing have been
+made, but none of them are so ambitious as our project. On
+the one hand the work done in [1] allows the VM managed
+by the hypervisor Xen to access the GPUs in the physical
+node. With the implied limitations of restricting the nodes not
+to use more GPUs than the hosted in the machine and not
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus [2] does virtualize GPUs and make
+them accessible for any VM in the cluster. However, this kind
+of virtualization strongly depends on the hypervisor used, so
+does its performance. Another similar solution is presented in
+[3] by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
+drawback is that the code of the applications must be modified
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in [4] is more mature, however,
+it is only focussed on compute intensive HPC applications.
+The main idea of our proposal goes further and apart from
+bringing solutions for all kind of HPC applications, it is aimed
+to boost the cluster flexibility in the use of GPUs.
 

### 2015-10-05 paper.tex
openstack section

@@
 
-\subsection{Manuscript Setup}
+\section{GPGPU Service for OpenStack}
+\subsection{Managing remote GPUs with OpenStack}
+The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecure.
+This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
+As it is shown in Figure~\ref{fig:internal}, we have altered the original Dashboard with a new parser, 
+which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
+Thanks to this modular approach, the Compute Module has not been modified, which led into an easily portability to other Cloud Computing environments.
 
-\subsection{Page Setup}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\linewidth]{images/os1.jpg}
+  \caption{Openstack with our GPGPU Service}
+  \label{fig:external}
+\end{figure}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=.9\linewidth]{images/os2.png}
+  \caption{Internal Communication among modules}
+  \label{fig:internal}
+\end{figure}
+
+\subsection{Development}
+We have developed an independent module that can be seen as a new shared service in the OpenStack Architecture (see Figure~\ref{fig1}).
+
+Thus, we have provided OpenStack with a new parser, which splits the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation.
+Thanks to this modular approach, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
+
+The new GPGPU Service has brought a new feature when it comes to GPUs, from now on, we can create ``GPU-pools''.
+A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
+
+Furthermore, we have extended the user interface (Horizon Project), in order to deal with the new features.
+First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
+As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration.
+Finally, with the purpose of being aware of the devices status, a new panel has been added with all the information regarding the GPUs.
+
+\subsection{Working Modes}
+The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
+The users are given two configuration options to decide whether a physical GPU will be completely allocated for an instance (first column) or the instance will address a partition of the GPU as if it were a device (second column).
+We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
+Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
+A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
+So, while in ``exclusive'' mode the instance is monopolising all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
+As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
+
+Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
+This behaviour is refereed as ``scope'' and determines that a group of instances is logically connected to a pool of GPUs.
+Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
+Again, the GPU pool can be made of ``exclusive'' GPUs or ``shared''.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=.5\textwidth]{images/workingmodes.jpg}
+  \caption{Examples of working modes}
+  \label{fig2}
+\end{figure}
+
+\subsection{User Interface}
+Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
+First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
+As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration (see Subfigure~\ref{fig:ui-launch}).
+Furthermore, a new panel with the existent GPUs has been added to the Interface (see Subfigure~\ref{fig:ui-rgpus}), where all the information of the GPUs is displayed.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=.9\linewidth]{images/UI-launch.pdf}
+  \caption{Launching Instances and assigning GPUs}
+  \label{fig:ui-launch}
+\end{figure}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=.85\linewidth]{images/UI-rgpus.pdf}
+  \caption{GPU Information Panel}
+  \label{fig:ui-rgpus}
+\end{figure}
+
+\subsection{Experimental Results}
 

### 2015-10-05 paper.tex
openstack section
Primera descripcion de AWS añadida

@@
 
+\subsection{Amazon Web Services}
+\label{sec:aws}
+AWS~\cite{aws} is a public cloud computing provider
+and it is composed by several cloud services such as 
+ cloud-based computation, storage and other func-
+tionality that enables organizations and/or individuals to deploy
+services and applications on an on-demand basis. 
+These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
+perfectly with their softwares requirements.
+
+From the point of view of HPC, AWS offers high performance facilities including 
+instances equiped with GPUs and high performance network interconnection which 
+are common resources used in this research field.
+
 \section{GPGPU Service for OpenStack}
@@
   \centering
-  \includegraphics[width=.9\linewidth]{images/UI-launch.pdf}
+  \includegraphics[width=.9\linewidth]{images/UI-launch.jpg}
   \caption{Launching Instances and assigning GPUs}
@@
   \centering
-  \includegraphics[width=.85\linewidth]{images/UI-rgpus.pdf}
+  \includegraphics[width=.85\linewidth]{images/UI-rgpus.jpg}
   \caption{GPU Information Panel}

### 2015-10-05 paper.tex
structura cambiada
Añadidos agradecimientos

@@
 
+\subsection{OpenStack}
+\label{sec:openstack}
+\begin{itemize}
+\item Descripcion basica del openstack (Sisco?)
+\end{itemize}
+
+\section{Motivation}
+\label{sec:motivation}
+
+\begin{itemize}
+\item Actual estado de las tecnolog\'ias de cloud
+\item Limitaciones que nos llevan a realizar este trabajo
+\item Beneficios de la unión de ambas tecnolog\'ias
+\end{itemize}
+
+\section{Remote GPUs in AWS}
+\label{sec:rgpuaws}
+
+\begin{itemize}
+\item Escenarios creados
+\item Experimentos
+\item An\'alisis basados en flexibilidad y coste econ\'omico
+\end{itemize}
+
 \section{GPGPU Service for OpenStack}
@@
 
+\section{Discussion}
+\label{sec:discussion}
+Volver a hacer un resumen
+
 \section*{\uppercase{Acknowledgements}}
+The researchers from the Universidad Jaume~I were supported by Universitat Jaume I research project (P11B2013-21), project
+TIN2011-23283 and~FEDER.
 

### 2015-10-05 paper.tex
openstack module section

--- 2015-10-05 paper.tex
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecure.
+The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
 This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
@@
   \centering
-  \includegraphics[width=.9\linewidth]{images/os2.png}
+  \includegraphics[width=\linewidth]{images/os2.png}
   \caption{Internal Communication among modules}

### 2015-10-05 paper.tex
Abstract modificado

--- 2015-10-05 paper.tex
@@
 
-\abstract{The popularity of the GPGPU computation has led several cloud vendors to provide instances of virtual machines with GPUs. 
-So that, the guest hosts must be equipped with GPUs which could be barely utilized if no GPU-enabled VM  is running in the host.
-The solution presented here is based on GPU virtualization and shareability in order to reach an equilibrium between supply and demand of accelerators from the users. 
-Hence, we propose to access any GPU from any VM getting rid of the necessity of having physical GPUs in the guest host. 
-Finally, our project goes further and it takes into account that by making logical partitions of the GPU memory we could address each partition as if it was an independent GPU device.}
+\abstract{The use of accelerators in order to reduce the execution time of some 
+application has become more popular during last years. These devices increment 
+the computational power of a computer thanks to its parallel architecture. %The popularity of the GPGPU computation 
+This trend has led several cloud vendors such as Amazon Web Services or some 
+middlewares such as OpenStack to add to their current facilities, or to allow, 
+instances of virtual machines equipped with NVIDIA GPUs. 
+Accomplishing this requirements needs that the guest hosts must be equipped with 
+GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability 
+in order to reach an equilibrium between supply and demand of accelerators from the users. 
+Hence, we propose to decouple, from the nodes, real GPUs by using the virtualization technology. 
+So, these GPUs can be accessed from any VM getting rid of the necessity of having physical 
+GPUs in the guest host. 
+Moreover, our project studies the viability of these join by using a public cloud 
+service configuration and develops a module for OpenStack middleware in order to 
+add the needed support for the virtualized devices.  
+%goes further and it takes into account that by making logical 
+%partitions of the GPU memory we could address each partition as if it was an 
+%independent GPU device.
+}
 
@@
 From the point of view of HPC, AWS offers high performance facilities including 
-instances equiped with GPUs and high performance network interconnection which 
+instances equipped with GPUs and high performance network interconnection which 
 are common resources used in this research field.
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecure.
+The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
 This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
@@
 A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
-So, while in ``exclusive'' mode the instance is monopolising all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
+So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
 As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
@@
 
+\section{Conclusions}
+\label{sec:conclusions}
+
+\section{Future work}
+\label{sec:future}
+
 \section*{\uppercase{Acknowledgements}}

### 2015-10-05 paper.tex
Titulo modificado
experimental OS

--- 2015-10-05 paper.tex
@@
 \noindent
-Nowadays, to provide a virtual machine with GPUs means
-that the guest host must be equipped with at least 1 GPU, and
-once it has been assigned to an instance, no other VM can use
-it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
-the conservative, to have a lot of hosts with GPUs to be
-ensured that you are likely to satisfy all the requests; or a
-more daring approach, where you will count with some GPU-
-enabled nodes hoping not to arrive a large burst of requests
-so the cluster would run out of GPU resources. Our solutions
-is based on GPU virtualization and sharing resources in order
-to reach a fair balance between supply and demand.
-Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project. On
-the one hand the work done in [1] allows the VM managed
-by the hypervisor Xen to access the GPUs in the physical
-node. With the implied limitations of restricting the nodes not
-to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented
-by the project gVirtus [2] does virtualize GPUs and make
-them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in
-[3] by the name of gCloud. This one continues without being
-integrated in a Cloud Computing Manager, though its main
-drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. On the
-other hand, the work exposed in [4] is more mature, however,
-it is only focussed on compute intensive HPC applications.
-The main idea of our proposal goes further and apart from
-bringing solutions for all kind of HPC applications, it is aimed
-to boost the cluster flexibility in the use of GPUs.
 
@@
 \label{sec:motivation}
+Nowadays, to provide a virtual machine with GPUs means
+that the guest host must be equipped with at least a GPU, and
+once it has been assigned to an instance, no other VM can use
+it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
+the conservative, to have a lot of hosts with GPUs to be
+ensured that you are likely to satisfy all the requests; or a
+more daring approach, where you will count with some GPU-
+enabled nodes hoping not to arrive a large burst of requests
+so the cluster would run out of GPU resources.
+
+Our solutions
+is based on GPU virtualization and sharing resources in order
+to reach a fair balance between supply and demand.
+Several attempts of achieving what we are pursing have been
+made, but none of them are so ambitious as our project. On
+the one hand the work done in~\cite{younge2013enabling} allows the VM managed
+by the hypervisor Xen to access the GPUs in the physical
+node. With the implied limitations of restricting the nodes not
+to use more GPUs than the hosted in the machine and not
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+them accessible for any VM in the cluster. However, this kind
+of virtualization strongly depends on the hypervisor used, so
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
+drawback is that the code of the applications must be modified
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+it is only focussed on compute intensive HPC applications.
+The main idea of our proposal goes further and apart from
+bringing solutions for all kind of HPC applications, it is aimed
+to boost the cluster flexibility in the use of GPUs.
 
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
+Our contribution reaches a solution that drive us to a scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
 This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
-As it is shown in Figure~\ref{fig:internal}, we have altered the original Dashboard with a new parser, 
+As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
 which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
-Thanks to this modular approach, the Compute Module has not been modified, which led into an easily portability to other Cloud Computing environments.
+The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
+A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
 
@@
 
-\subsection{Development}
-We have developed an independent module that can be seen as a new shared service in the OpenStack Architecture (see Figure~\ref{fig1}).
-
-Thus, we have provided OpenStack with a new parser, which splits the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation.
-Thanks to this modular approach, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
-
-The new GPGPU Service has brought a new feature when it comes to GPUs, from now on, we can create ``GPU-pools''.
-A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
-
-Furthermore, we have extended the user interface (Horizon Project), in order to deal with the new features.
-First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration.
-Finally, with the purpose of being aware of the devices status, a new panel has been added with all the information regarding the GPUs.
+Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
 
@@
 Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
-First of all, the Instance Launch Panel has been provided with a new field, where the user can assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' has been chosen, will appear the proper fields for the pool configuration (see Subfigure~\ref{fig:ui-launch}).
-Furthermore, a new panel with the existent GPUs has been added to the Interface (see Subfigure~\ref{fig:ui-rgpus}), where all the information of the GPUs is displayed.
+First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
+As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
+Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
 
@@
   \centering
-  \includegraphics[width=.9\linewidth]{images/UI-launch.jpg}
+  \includegraphics[width=\linewidth]{images/UI-launch.jpg}
   \caption{Launching Instances and assigning GPUs}
@@
   \centering
-  \includegraphics[width=.85\linewidth]{images/UI-rgpus.jpg}
+  \includegraphics[width=\linewidth]{images/UI-rgpus.jpg}
   \caption{GPU Information Panel}
@@
 \subsection{Experimental Results}
+The test were executed on a 5-node cluster where a machine
+equipped with a Intel Xeon E7420 quadcore processor at
+2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
+MHz, was guest host for the VMs. While the other 4 nodes
+were auxiliary servers with a Tesla C1090 GPU each one. 
+In the software stack the testbed system operates a
+Centos 6.6; the cloud platform is deployed by OpenStack; and
+the GPUs use CUDA 6.5. The GPU virtualization support is
+based on rCUDA v5.0b.
+We have designed 6 different set-ups which can be divided
+in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
+“Exclusive” mode will provide, at most, 4 accelerators. Whilst
+the number of available GPU in “Shared” mode will depend
+on the partition size. In this particular case, we have halved
+the memory, so we can use up to 8 GPUs. For each group, we
+have deployed virtual clusters of 1, 2 and 3 nodes where will
+run the processes of the application. The instances are based
+on the flavor “m1.medium” which determines VM of 2 cores
+and 4096 MB of memory.
+MCUDA-MEME~\cite{Liu2010} has been the application executed with
+different configurations in order to test the set-ups. MCUDA-
+MEME is GPGPU MPI application, so the number of GPUs
+will determine the number of executable processes.
+The Figure~\ref{fig3} shows and compares the execution time of
+the application with different configurations over different set-
+ups. We can determine that with 1 node, the performance is
+lower, with 2 and 3 is quite similar, though. Hence, the greatest
+difference can be appreciated between modes, where the
+highest performance is achieved by sharing the accelerators.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.5\textwidth]{images/chart.jpg}
+  \caption{Execution time of MCUDA-MEME}
+  \label{fig3}
+\end{figure}
 
@@
 \label{sec:discussion}
-Volver a hacer un resumen
+-Volver a hacer un resumen
+-posibles futuras aplicaciones (lab)
+-pruebas con otros proveedores o con otras plataformas (azure)
 

### 2015-10-05 paper.tex

@@
 
-\title{GPU Virtualization in the Cloud (a mejorar el t\'itulo!!!)}
+\title{Enabling the Use of GPU Virtualization in Cloud Environments}
 \author{\authorname{Sergio Iserte, Francisco J. Clemente-Castell\'o, Adri\'an Castell\'o,\\Rafael Mayo and Enrique S. Quintana-Ort\'i}
@@
 \subsection{Managing remote GPUs with OpenStack}
-Our contribution reaches a solution that drive us to a scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part into the OpenStack Architecure.
+The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
+>>>>>>> b29156d83b82f6d0c9490cadc106cdbd54178d8e
 This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.

### 2015-10-06 paper.tex
Abstract modificado

@@
 application has become more popular during last years. These devices increment 
-the computational power of a computer thanks to its parallel architecture. %The popularity of the GPGPU computation 
+the computational power of a node thanks to its parallel architecture. %The popularity of the GPGPU computation 
 This trend has led several cloud vendors such as Amazon Web Services or some 
@@
 in order to reach an equilibrium between supply and demand of accelerators from the users. 
-Hence, we propose to decouple, from the nodes, real GPUs by using the virtualization technology. 
-So, these GPUs can be accessed from any VM getting rid of the necessity of having physical 
-GPUs in the guest host. 
+Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
+With this software configuration, these GPUs can be accessed from any VM getting rid 
+of the necessity of having physical GPUs in the guest host. 
 Moreover, our project studies the viability of these join by using a public cloud 
@@
 add the needed support for the virtualized devices. 
+The results show a feasible configuration which adds flexibility to the current, well-known 
+cloud solutions. 
 %goes further and it takes into account that by making logical 

### 2015-10-06 paper.tex
ispell pasado

@@
 add the needed support for the virtualized devices. 
-The results show a feasible configuration which adds flexibility to the current, well-known 
+The results show a feasible configuration which adds flexibility to the current and well-known 
 cloud solutions. 
@@
 and it is composed by several cloud services such as 
- cloud-based computation, storage and other func-
-tionality that enables organizations and/or individuals to deploy
+ cloud-based computation, storage and other functionality 
+that enables organizations and/or individuals to deploy
 services and applications on an on-demand basis. 
 These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
-perfectly with their softwares requirements.
+perfectly with their software requirements.
 

### 2015-10-06 paper.tex
image AWS

--- 2015-10-05 paper.tex
@@
 service configuration and develops a module for OpenStack middleware in order to 
-add the needed support for the virtualized devices.  
+add the needed support for the virtualized devices and the logic to manage them.  
 %goes further and it takes into account that by making logical 
@@
 
-\section{Remote GPUs in AWS}
+\section{Working with Remote GPUs in AWS}
 \label{sec:rgpuaws}
@@
 \begin{itemize}
-\item Escenarios creados
+\item Esc\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\linewidth]{images/AWS.pdf}
+  \caption{An AWS instance with 13 remote GPUs}
+  \label{fig:aws1}
+\end{figure}enarios creados
+
 \item Experimentos
@@
 
-\section{GPGPU Service for OpenStack}
+\section{GPGPU Shared Service in OpenStack}
+In this section is explained all the work related to OpenStack that has been done, from the development of an external module to handle remote GPUs to the physical setup and its testing.
 \subsection{Managing remote GPUs with OpenStack}
-The main idea of this project is to reach a solution that drive us to an scenario pretty similar the one that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
->>>>>>> b29156d83b82f6d0c9490cadc106cdbd54178d8e
+The main idea of this solution is to reach the scenario that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
 This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.

### 2015-10-06 paper.tex
pdfs images

--- 2015-10-06 paper.tex
@@
 service configuration and develops a module for OpenStack middleware in order to 
-add the needed support for the virtualized devices. 
+add the needed support for the virtualized devices and the logic to manage them.. 
 The results show a feasible configuration which adds flexibility to the current and well-known 
 cloud solutions. 
-%goes further and it takes into account that by making logical 
-%partitions of the GPU memory we could address each partition as if it was an 
-%independent GPU device.
 }
@@
   \centering
-  \includegraphics[width=\linewidth]{images/UI-launch.jpg}
+  \includegraphics[width=\linewidth]{images/UI-launch.pdf}
   \caption{Launching Instances and assigning GPUs}
@@
   \centering
-  \includegraphics[width=\linewidth]{images/UI-rgpus.jpg}
+  \includegraphics[width=\linewidth]{images/UI-rgpus.pdf}
   \caption{GPU Information Panel}

### 2015-10-06 paper.tex
aws images side by side
motivacion modificada

@@
 service configuration and develops a module for OpenStack middleware in order to 
-add the needed support for the virtualized devices and the logic to manage them.. 
+add the needed support for the virtualized devices and the logic to manage them. 
 The results show a feasible configuration which adds flexibility to the current and well-known 
@@
 \label{sec:motivation}
-Nowadays, to provide a virtual machine with GPUs means
-that the guest host must be equipped with at least a GPU, and
+
+Cloud computing, storage and other facilities are becoming 
+more popular during the last years thanks to the easy set up 
+process and the low required IT invest. The features provided 
+by the cloud vendors are useful for companies and individuals 
+which have access to the last technology without spending a 
+huge amount of money. Moreover, the systems offered by the cloud 
+companies matches with the ones required thanks to its flexibility.
+
+The power of current cloud vendors is based on the use 
+of Virtual Machines (VM) which resides inside the host machine and 
+emulates a physical node with its own 
+cpus, storage, network interconnection, OS and GPU. 
+However, to provide a VM with GPUs means
+that the guest host must be equipped with at least one GPU, and
 once it has been assigned to an instance, no other VM can use
-it. Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
+it.
+
+Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
 the conservative, to have a lot of hosts with GPUs to be

### 2015-10-06 paper.tex
aws tabla iperf
motivacion modificada

@@
 \noindent 
+Our solutions
+is based on GPU virtualization and sharing resources in order
+to reach a fair balance between supply and demand.
+Several attempts of achieving what we are pursing have been
+made, but none of them are so ambitious as our project. On
+the one hand the work done in~\cite{younge2013enabling} allows the VM managed
+by the hypervisor Xen to access the GPUs in the physical
+node. With the implied limitations of restricting the nodes not
+to use more GPUs than the hosted in the machine and not
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+them accessible for any VM in the cluster. However, this kind
+of virtualization strongly depends on the hypervisor used, so
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
+drawback is that the code of the applications must be modified
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+it is only focussed on compute intensive HPC applications.
+The main idea of our proposal goes further and apart from
+bringing solutions for all kind of HPC applications, it is aimed
+to boost the cluster flexibility in the use of GPUs.
 
@@
 The power of current cloud vendors is based on the use 
-of Virtual Machines (VM) which resides inside the host machine and 
+of Virtual Machines (VM) which  
 emulates a physical node with its own 
@@
 once it has been assigned to an instance, no other VM can use
-it.
-
-Depending on the users’ necessities, two approaches to introduce GPGPU computation in the cluster can be determined:
+it. Therefore, the number of VM that are allowed to use a GPU is 
+limited by the number of the installed devices. This lack of 
+flexibility is also noticed in the cluster configurations, where 
+depending on the users’ necessities, 
+two approaches to introduce GPGPU computation can be determined:
 the conservative, to have a lot of hosts with GPUs to be
@@
 
-Our solutions
-is based on GPU virtualization and sharing resources in order
-to reach a fair balance between supply and demand.
-Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project. On
-the one hand the work done in~\cite{younge2013enabling} allows the VM managed
-by the hypervisor Xen to access the GPUs in the physical
-node. With the implied limitations of restricting the nodes not
-to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented
-by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
-them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
-integrated in a Cloud Computing Manager, though its main
-drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. On the
-other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
-it is only focussed on compute intensive HPC applications.
-The main idea of our proposal goes further and apart from
-bringing solutions for all kind of HPC applications, it is aimed
-to boost the cluster flexibility in the use of GPUs.
+The use of a virtualization technology such as {rCUDA} adds 
+the desired flexibility because this software detachs the GPUs 
+from the nodes where they are installed allowing all nodes in a cluster 
+not only use them as if they were local but also to share them 
+achieving a higher utilization rate and reducing the idle status time.
+
+Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
+environment should be able to use and share all the GPUs installed in 
+the physical nodes removing the current limitation. But, in order to achieve this goal, 
+the cloud facilities need to be provided with the {rCUDA} software and 
+in the case of OpenStack, a module for configuring virtualized GPUs 
+ has to be implemented.
 
@@
 
+\begin{figure*}[ht]
+\centering
+\subfigure[An instance of 5 remote GPUs]{%
+\includegraphics[width=0.31\linewidth]{images/aws3.png}
+\label{subfig:aws1}}
+\quad
+\subfigure[An instance of 8 remote GPUs]{%
+\includegraphics[width=0.31\linewidth]{images/aws2.png}
+\label{subfig:aws2}}
+\subfigure[An instance of 13 remote GPUs]{%
+\includegraphics[width=0.31\linewidth]{images/aws1.png}
+\label{subfig:aws3}}
+\caption{AWS scenarios}
+\label{fig:aws}
+\end{figure*}
+
 \begin{itemize}
-\item Esc\begin{figure}[!t]
-  \centering
-  \includegraphics[width=\linewidth]{images/AWS.pdf}
-  \caption{An AWS instance with 13 remote GPUs}
-  \label{fig:aws1}
-\end{figure}enarios creados
-
+\item Escenarios creados
 \item Experimentos

### 2015-10-06 paper.tex
aws tabla instances

--- 2015-10-06 paper.tex
@@
 
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{AWS Instances}
+\label{table:awsInstances}
+\tabcolsep=0.09cm
+\begin{center}\begin{tabular}{cccccc}
+Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
+c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.21\\ \hline
+c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
+g2.2xlarge & 8 & 15 GiB & High & 1 & \$ 0.65\\ \hline
+g2.8xlarge & 32 & 60 GiB & 10 Gb & 4 & \$ 2.6 \\ \hline
+\end{tabular}\end{center}\end{table}
+
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{IPERF results between instances}
+\label{table:iperf}
+\tabcolsep=0.4cm
+\begin{center}\begin{tabular}{ccc}
+Server & Client & BandWidth Gb/s\\ \hline \hline
+g2.8xlarge & c3.2xlarge & 1 \\ \hline
+g2.8xlarge & c3.8xlarge & 7.5 \\ \hline
+g2.8xlarge & g2.2xlarge & 1 \\ \hline
+g2.8xlarge & g2.8xlarge & 7.5 \\ \hline
+\end{tabular}\end{center}\end{table}
+
 \section{Working with Remote GPUs in AWS}
@@
 
+\begin{figure*}[ht]
+\centering
+\subfigure[5 remote GPUs over 1GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws3.png}
+\label{subfig:aws1}}
+\quad
+\subfigure[8 remote GPUs over 10GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws2.png}
+\label{subfig:aws2}}
+\subfigure[13 remote GPUs over 1GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws1.png}
+\label{subfig:aws3}}
+\caption{An AWS instance with several remote GPUs}
+\label{fig:aws}
+\end{figure*}
+
 \begin{itemize}
-\item Esc\begin{figure}[!t]
-  \centering
-  \includegraphics[width=\linewidth]{images/AWS.pdf}
-  \caption{An AWS instance with 13 remote GPUs}
-  \label{fig:aws1}
-\end{figure}enarios creados
-
+\item Escenarios creados
 \item Experimentos

### 2015-10-06 paper.tex

--- 2015-10-06 paper.tex
@@
 \centering
-\subfigure[An instance of 5 remote GPUs]{%
+\subfigure[5 remote GPUs over 1GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws3.png}
@@
 \quad
-\subfigure[An instance of 8 remote GPUs]{%
+\subfigure[8 remote GPUs over 10GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws2.png}
 \label{subfig:aws2}}
-\subfigure[An instance of 13 remote GPUs]{%
+\subfigure[13 remote GPUs over 1GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws1.png}
 \label{subfig:aws3}}
-\caption{AWS scenarios}
+\caption{An AWS instance with several remote GPUs}
 \label{fig:aws}

### 2015-10-06 paper.tex
ispell pasado

--- 2015-10-06 paper.tex
@@
 emulates a physical node with its own 
-cpus, storage, network interconnection, OS and GPU. 
+CPUs, storage, network interconnection, OS and GPU. 
 However, to provide a VM with GPUs means
@@
 The use of a virtualization technology such as {rCUDA} adds 
-the desired flexibility because this software detachs the GPUs 
+the desired flexibility because this software detaches the GPUs 
 from the nodes where they are installed allowing all nodes in a cluster 

### 2015-10-06 paper.tex
aws index

--- 2015-10-06 paper.tex
@@
 
+\section{Working with AWS}
+\label{sec:aws}
+
+\subsection{Current Features}
+
+\subsubsection{Instances}
 \begin{table}[!t]
@@
 
+\subsubsection{Network}
 \begin{table}[!t]
@@
 \label{table:iperf}
-\tabcolsep=0.4cm
-\begin{center}\begin{tabular}{ccc}
-Server & Client & BandWidth Gb/s\\ \hline \hline
-g2.8xlarge & c3.2xlarge & 1 \\ \hline
-g2.8xlarge & c3.8xlarge & 7.5 \\ \hline
-g2.8xlarge & g2.2xlarge & 1 \\ \hline
-g2.8xlarge & g2.8xlarge & 7.5 \\ \hline
+\tabcolsep=0.24cm
+\begin{center}\begin{tabular}{cccc}
+Server & Client & Network & BandWidth\\ \hline \hline
+g2.8xlarge & c3.2xlarge & High & 1  Gb/s\\ \hline
+g2.8xlarge & c3.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
+g2.8xlarge & g2.2xlarge & High & 1 Gb/s\\ \hline
+g2.8xlarge & g2.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
 \end{tabular}\end{center}\end{table}
 
-\section{Working with Remote GPUs in AWS}
-\label{sec:rgpuaws}
+\subsubsection{GPUs}
 
@@
 
-\begin{itemize}
-\item Escenarios creados
-\item Experimentos
-\item An\'alisis basados en flexibilidad y coste econ\'omico
-\end{itemize}
+\subsection{Testbed Scenarios}
+\subsection{Experiments}
+\subsection{Results}
+Flexibility and economical expenses
 

### 2015-10-06 paper.tex
AWS iniciado

@@
 the cloud facilities need to be provided with the {rCUDA} software and 
-in the case of OpenStack, a module for configuring virtualized GPUs 
+in the case of OpenStack, a module for manageing virtualized GPUs 
  has to be implemented.
-
-\begin{itemize}
-\item Actual estado de las tecnolog\'ias de cloud
-\item Limitaciones que nos llevan a realizar este trabajo
-\item Beneficios de la unión de ambas tecnolog\'ias
-\end{itemize}
 
@@
 \label{sec:aws}
+In this section, first of all, the current user-level features 
+provided by the Amazon Web Services are explained and analyzed 
+from the point of view of the GPU usage. 
+Finally, the scenarios and experiments where virtualized GPUs are utilized are 
+described and first results are showed.
 
 \subsection{Current Features}
+This subsection explains the most common features checked during 
+the virtual machine creation step.
 
 \subsubsection{Instances}
+
+An instance is a pre-configured virtual machine focused on an 
+specific target. Inside the large list of instances offered by AWS, 
+we can find: general pourpose (T2, M4 and M3); computer science optimized 
+(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
+GPU capable instances. Each kind of instance has its own pourpose and its 
+own price. Moreover, each instance has different number of CPUs and network 
+interconnection which can be: low, medium, high or 10Gb.
+
+For this study, the instances presented in the Table~\ref{table:awsInstances} 
+have been selected. The C3 family instances, which are not equipped with 
+GPUs are going to be used as client whereas the instances of the G2 family are 
+acting as servers.
+
 \begin{table}[!t]
@@
 
-\subsubsection{Network}
+\subsubsection{Network behaviour}
+As it is shown in the Table~\ref{table:awsInstances}, each instance has different 
+interconnection and taking into account that the use of the network 
+is critical when GPU virtualization is applied, a simple test to verify the 
+real network bandwidth is needed.
+
 \begin{table}[!t]
@@
 
+The IPERF tool has been executed between the instances showed in 
+the Table~\ref{table:iperf} in order to know the real bandwidth.
+Therefore, we can assure that the nomenclature ''High'' used in the 
+network column correspond to a 1Gb interconnection and the 10Gb has 
+a real bandwidth of 7.5Gb/s.
+
 \subsubsection{GPUs}
 
+\subsection{Testbed Scenarios}
 \begin{figure*}[ht]
@@
 
-\subsection{Testbed Scenarios}
 \subsection{Experiments}

### 2015-10-06 paper.tex
AWS GPU subseccioniniciado

@@
 \subsubsection{GPUs}
+As it has been said before, an instance is based in a VM 
+which is running on a real node with its own virtualized 
+components, so it is possible that AWS could use a virtualization 
+framework to offer GPU services to all the instances.
+Therefore, we need to verify if the GPUs showed in the instances are 
+non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
+bandwidthtest. As it is shown in the Table~\ref{}, as the bandwidth achieved 
+in this test is higher than the network bandwith, it reveals that it is a 
+real GPU.
 

### 2015-10-06 paper.tex
Titulos cambiados
acks

--- 2015-10-06 paper.tex
@@
 \section{GPGPU Shared Service in OpenStack}
-In this section is explained all the work related to OpenStack that has been done, from the development of an external module to handle remote GPUs to the physical setup and its testing.
+In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
 \subsection{Managing remote GPUs with OpenStack}
@@
 \section*{\uppercase{Acknowledgements}}
-The researchers from the Universidad Jaume~I were supported by Universitat Jaume I research project (P11B2013-21), project
+The authors want to thank the IT members of the department Vicente Roca and Gustavo Edo for the given help.
+The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
 TIN2011-23283 and~FEDER.

### 2015-10-06 paper.tex

--- 2015-10-06 paper.tex
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{AWS Instances}
+\caption{AWS available HPC instances}
 \label{table:awsInstances}
@@
 \label{subfig:aws3}}
-\caption{An AWS instance with several remote GPUs}
+\caption{Different scenario configurations where an AWS instance acting as a client is using several remote GPUs}
 \label{fig:aws}

### 2015-10-06 paper.tex
bandwidth test añadido

--- 2015-10-06 paper.tex
@@
 non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-bandwidthtest. As it is shown in the Table~\ref{}, as the bandwidth achieved 
+bandwidthtest. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
 in this test is higher than the network bandwith, it reveals that it is a 
 real GPU. 
+\begin{table}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{NVIDIA SDK bandwidthtest execution transfering 32MB using pageable memory }
+\label{table:bwt}
+\tabcolsep=0.05cm
+\begin{center}\begin{tabular}{cccc}
+Name &  Data Movement & Network & BandWidth (MB/s)\\ \hline \hline
+g2.2xlarge & Host to Device & Hight& 3,004.5\\ \hline
+g2.2xlarge & Device to Host & Hight& 2,809.2\\ \hline
+g2.8xlarge & Host to Device & 10 Gb& 2,814.7\\ \hline
+g2.8xlarge & Device to Host & 10 Gb& 3,182.3\\ \hline
+\end{tabular}\end{center}\end{table}
 

### 2015-10-06 paper.tex
Tabla del bandwidthtest modificada

@@
 \begin{center}\begin{tabular}{cccc}
-Server & Client & Network & BandWidth\\ \hline \hline
+Server & Client & Network & Bandwidth\\ \hline \hline
 g2.8xlarge & c3.2xlarge & High & 1  Gb/s\\ \hline
@@
 \label{table:bwt}
-\tabcolsep=0.05cm
+\tabcolsep=0.09cm
 \begin{center}\begin{tabular}{cccc}
-Name &  Data Movement & Network & BandWidth (MB/s)\\ \hline \hline
-g2.2xlarge & Host to Device & Hight& 3,004.5\\ \hline
-g2.2xlarge & Device to Host & Hight& 2,809.2\\ \hline
-g2.8xlarge & Host to Device & 10 Gb& 2,814.7\\ \hline
-g2.8xlarge & Device to Host & 10 Gb& 3,182.3\\ \hline
+Name &  Data Movement & Network & BandWidth \\ \hline \hline
+g2.2xlarge & Host to Device & Hight& 3,004 MB/s \\ \hline
+g2.2xlarge & Device to Host & Hight& 2,809 MB/s\\ \hline
+g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
+g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
 \end{tabular}\end{center}\end{table}

### 2015-10-06 paper.tex
Tabla del bandwidthtest modificada

@@
 the Table~\ref{table:iperf} in order to know the real bandwidth.
-Therefore, we can assure that the nomenclature ''High'' used in the 
+Therefore, we can assure that the nomenclature ``High'' used in the 
 network column correspond to a 1Gb interconnection and the 10Gb has 
 a real bandwidth of 7.5Gb/s.
+Moreover, it looks like the instances equipped with a ``High'' interconnection 
+are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
+match perfectly. The gap between real and theorical can be observed with 
+the 10Gb interconnection which reaches up to 7.5 Gb/s.
 

### 2015-10-06 paper.tex
montecarlo aws

--- 2015-10-06 paper.tex
@@
 non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-bandwidthtest. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
+bandwidthtest. As it is shown in the Table~\ref{}, as the bandwidth achieved 
 in this test is higher than the network bandwith, it reveals that it is a 
 real GPU.
-\begin{table}[!t]
-\renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK bandwidthtest execution transfering 32MB using pageable memory }
-\label{table:bwt}
-\tabcolsep=0.05cm
-\begin{center}\begin{tabular}{cccc}
-Name &  Data Movement & Network & BandWidth (MB/s)\\ \hline \hline
-g2.2xlarge & Host to Device & Hight& 3,004.5\\ \hline
-g2.2xlarge & Device to Host & Hight& 2,809.2\\ \hline
-g2.8xlarge & Host to Device & 10 Gb& 2,814.7\\ \hline
-g2.8xlarge & Device to Host & 10 Gb& 3,182.3\\ \hline
-\end{tabular}\end{center}\end{table}
 
 \subsection{Testbed Scenarios}
-\begin{figure*}[ht]
+\begin{figure*}[!t]
 \centering
@@
 \label{subfig:aws3}}
-\caption{Different scenario configurations where an AWS instance acting as a client is using several remote GPUs}
+\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs}
 \label{fig:aws}
@@
 \subsection{Results}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\linewidth]{images/aws-mont2.png}
+  \caption{Montecarlo execution time}
+  \label{fig:internal}
+\end{figure}
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=\linewidth]{images/aws-mont1.png}
+  \caption{Montecarlo options per second}
+  \label{fig:internal}
+\end{figure}
+
 Flexibility and economical expenses
@@
 \section*{\uppercase{Acknowledgements}}
-The researchers from the Universidad Jaume~I were supported by Universitat Jaume I research project (P11B2013-21), project
+The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
+
+The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
 TIN2011-23283 and~FEDER.

### 2015-10-06 paper.tex

--- 2015-10-06 paper.tex
@@
 \subsection{Testbed Scenarios}
-\begin{figure*}[ht]
+\begin{figure*}[!t]
 \centering

### 2015-10-06 paper.tex
Localizacion de las figuras y tablas modificado

@@
 
-\begin{table}[!t]
+\begin{table}[htb]
 \renewcommand{\arraystretch}{1.3}
@@
 non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-bandwidthtest. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
+{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
 in this test is higher than the network bandwith, it reveals that it is a 
 real GPU. 
-\begin{table}[!t]
+\begin{table}[htb]
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK bandwidthtest execution transfering 32MB using pageable memory }
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory }
 \label{table:bwt}
@@
 \subsection{Testbed Scenarios}
-\begin{figure*}[!t]
+\begin{figure*}[ht]
 \centering
@@
 
-\begin{figure}[!t]
+\begin{figure}[htb]
   \centering
@@
 \end{figure}
-\begin{figure}[!t]
+\begin{figure}[htb]
   \centering

### 2015-10-06 paper.tex
AWS escenarios añadidos

@@
 \subsection{Testbed Scenarios}
+In this section, the different client-server configurations 
+are explained. All the scenarios are based in the maximum number of instance 
+that a user can freely select without submiting a formal request.
+
+The first scenarion (Figure~\ref{subfig:aws1}) shows the most common 
+and used hardware configuration in real clusters, where each node is 
+populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
+using the ``High'' interconnection.
+
 \begin{figure*}[ht]
@@
 \end{figure*}
+
+The second configuration (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
+with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
+a 10Gb network and the cliend is able to execute application using up to 8 GPUs.
+
+The last scenario (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
+using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
 

### 2015-10-06 paper.tex
results montecarlo

--- 2015-10-06 paper.tex
@@
 \noindent
-Our solutions
-is based on GPU virtualization and sharing resources in order
-to reach a fair balance between supply and demand.
-Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project. On
-the one hand the work done in~\cite{younge2013enabling} allows the VM managed
-by the hypervisor Xen to access the GPUs in the physical
-node. With the implied limitations of restricting the nodes not
-to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented
-by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
-them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
-integrated in a Cloud Computing Manager, though its main
-drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. On the
-other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
-it is only focussed on compute intensive HPC applications.
-The main idea of our proposal goes further and apart from
-bringing solutions for all kind of HPC applications, it is aimed
-to boost the cluster flexibility in the use of GPUs.
 
@@
 \end{itemize}
+
+\section{State of Art}
+\label{sec:state}
+Our solutions
+is based on GPU virtualization and sharing resources in order
+to reach a fair balance between supply and demand.
+Several attempts of achieving what we are pursing have been
+made, but none of them are so ambitious as our project. On
+the one hand the work done in~\cite{younge2013enabling} allows the VM managed
+by the hypervisor Xen to access the GPUs in the physical
+node. With the implied limitations of restricting the nodes not
+to use more GPUs than the hosted in the machine and not
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+them accessible for any VM in the cluster. However, this kind
+of virtualization strongly depends on the hypervisor used, so
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
+drawback is that the code of the applications must be modified
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+it is only focussed on compute intensive HPC applications.
+The main idea of our proposal goes further and apart from
+bringing solutions for all kind of HPC applications, it is aimed
+to boost the cluster flexibility in the use of GPUs.
 
@@
 
-\subsection{Experiments}
-\subsection{Results}
+\subsection{Experimental results}
+In order to assess that GPU virtualization with AWS instances was working as expected, two applications were executed to analyze the results and their scalability.
 
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=\linewidth]{images/aws-mont2.png}
-  \caption{Montecarlo execution time}
-  \label{fig:internal}
-\end{figure}
+The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of~\ref{fig:aws} and on the instances with local GPUs. 
+We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
+In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
+Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
+However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
+unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
+
+% \begin{figure}[!t]
+%   \centering
+%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
+%   \caption{Montecarlo execution time}
+%   \label{fig:mont-exe}
+% \end{figure}
 \begin{figure}[!t]
@@
   \caption{Montecarlo options per second}
-  \label{fig:internal}
+  \label{fig:mont-opt}
 \end{figure}
@@
 the GPUs use CUDA 6.5. The GPU virtualization support is
-based on rCUDA v5.0b.
+based on rCUDA v5.0.
+
 We have designed 6 different set-ups which can be divided
@@
 and 4096 MB of memory.
+
 MCUDA-MEME~\cite{Liu2010} has been the application executed with

### 2015-10-06 paper.tex

@@
 \subsection{Testbed Scenarios}
-\begin{figure*}[!t]
+In this section, the different client-server configurations 
+are explained. All the scenarios are based in the maximum number of instance 
+that a user can freely select without submiting a formal request.
+
+The first scenarion (Figure~\ref{subfig:aws1}) shows the most common 
+and used hardware configuration in real clusters, where each node is 
+populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
+using the ``High'' interconnection.
+
+The second configuration (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
+with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
+a 10Gb network and the cliend is able to execute application using up to 8 GPUs.
+
+The last scenario (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
+using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
+
+\begin{figure*}[ht]
 \centering
@@
 % \end{figure}
-\begin{figure}[!t]
+
+\begin{figure}[htb]
   \centering

### 2015-10-06 paper.tex
OpenStack description

@@
 \label{sec:openstack}
-\begin{itemize}
-\item Descripcion basica del openstack (Sisco?)
-\end{itemize}
+
+One of the most prominent open-source cloud computing platform for both private and public clouds is OpenStack.
+The OpenStack project was announced in 2010 by Rackspace and NASA, and currently developed and distributed by the OpenStack Foundation with the support of more than 150 relevant companies.
+OpenStack is a cloud operating system that provides Infrastructure as a Service (IaaS) able to control large pools of compute, storage, and networking resources throughout a datacenter, all managed through a dashboard or an API that gives administrators control while empowering their users to provision resources through a web interface or a command-line.
 

### 2015-10-06 paper.tex
OpenStack description
OpenStack corregido
results lammps

--- 2015-10-06 paper.tex
@@
 The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
-The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of~\ref{fig:aws} and on the instances with local GPUs. 
+The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
 We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
@@
 unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
-
 % \begin{figure}[!t]
@@
 % \end{figure}
-
 \begin{figure}[htb]
@@
 
+The other application, LAMMPS\footnote{\url{http://lammps.sandia.gov}}, is a classical molecular dynamics simulator that can be applied at the atomic, meso, or continuum scale. 
+From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
+The Figure shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
+Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
+Thus, if we compare the results of a LAMMPS execution with a larger problem (see Figure), while CUDA performs better, the interesting point is the execution time on remote GPUs. 
+They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
+pays off the drawback of a slow network.
+
+\begin{figure}[htb]
+\centering
+\subfigure[first]{%
+\includegraphics[width=0.455\linewidth]{images/lammps100.png}
+\label{subfig:lammps1}}
+\quad
+\subfigure[second]{%
+\includegraphics[width=0.455\linewidth]{images/lammps2000.png}
+\label{subfig:lammps2}}
+\caption{Execution time of LAMMPS}
+\label{fig:lammps}
+\end{figure}
+
+\subsection{Discussion}
 Flexibility and economical expenses

### 2015-10-06 paper.tex

@@
 \end{document}
+
+=======
+\documentclass[a4paper,twoside]{article}
+
+\usepackage{epsfig}
+\usepackage{subfigure}
+\usepackage{calc}
+\usepackage{amssymb}
+\usepackage{amstext}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{multicol}
+\usepackage{pslatex}
+\usepackage{apalike}
+\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
+
+\subfigtopskip=0pt
+\subfigcapskip=0pt
+\subfigbottomskip=0pt
+
+\begin{document}
+
+\title{Enabling the Use of GPU Virtualization in Cloud Environments}
+\author{\authorname{Sergio Iserte, Francisco J. Clemente-Castell\'o, Adri\'an Castell\'o,\\Rafael Mayo and Enrique S. Quintana-Ort\'i}
+\affiliation{Department of Computer Science and Engineering\\Universitat Jaume I - Castell\'o de la Plana, Spain}
+\email{\{siserte, fclement, adcastel, mayo, quintana\}@uji.es}}
+
+\keywords{Cloud Computing, GPU Virtualization, Resource Management}
+
+\abstract{The use of accelerators in order to reduce the execution time of some 
+application has become more popular during last years. These devices increment 
+the computational power of a node thanks to its parallel architecture. %The popularity of the GPGPU computation 
+This trend has led several cloud vendors such as Amazon Web Services or some 
+middlewares such as OpenStack to add to their current facilities, or to allow, 
+instances of virtual machines equipped with NVIDIA GPUs. 
+Accomplishing this requirements needs that the guest hosts must be equipped with 
+GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability 
+in order to reach an equilibrium between supply and demand of accelerators from the users. 
+Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
+With this software configuration, these GPUs can be accessed from any VM getting rid 
+of the necessity of having physical GPUs in the guest host. 
+Moreover, our project studies the viability of these join by using a public cloud 
+service configuration and develops a module for OpenStack middleware in order to 
+add the needed support for the virtualized devices and the logic to manage them. 
+The results show a feasible configuration which adds flexibility to the current and well-known 
+cloud solutions. 
+}
+
+\onecolumn \maketitle \normalsize \vfill
+
+\section{\uppercase{Introduction}}
+\label{sec:introduction}
+
+\noindent
+
+\section{\uppercase{Background}}
+\label{sec:background}
+\subsection{The rCUDA Framework}
+\label{sec:rcuda}
+
+{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
+to any NVIDIA GPU device present in a cluster from all compute
+nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
+as if they were local.
+These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
+resource, space, and energy~\cite{energy14} requirements.
+rCUDA is structured following a client-server distributed
+architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
+acceleration services is executed, providing a transparent replacement for the
+native CUDA libraries. Furthermore, the server middleware is executed in the
+cluster nodes from which the actual GPUs provide the requested GPGPU service.
+To support a concurrent scenario where GPUs are shared between
+processes\slash nodes, {rCUDA} manages separate device contexts for
+each client application.
+
+The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
+CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
+APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
+Therefore, applications are not aware of the fact that they are being executed
+on top of a virtualization layer.
+With the aim to be updated with new GPU programming models, {rCUDA} also supports
+directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.
+
+The integration of remote GPGPU virtualization with global
+resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
+technology, making accelerator-enabled clusters more flexible and
+energy efficient.
+
+\subsection{Amazon Web Services}
+\label{sec:aws}
+AWS~\cite{aws} is a public cloud computing provider
+and it is composed by several cloud services such as 
+ cloud-based computation, storage and other functionality 
+that enables organizations and/or individuals to deploy
+services and applications on an on-demand basis. 
+These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
+perfectly with their software requirements.
+
+From the point of view of HPC, AWS offers high performance facilities including 
+instances equipped with GPUs and high performance network interconnection which 
+are common resources used in this research field.
+
+\subsection{OpenStack}
+\label{sec:openstack}
+
+One of the most prominent open-source cloud computing platform for both private 
+and public clouds is OpenStack. The OpenStack project was announced in 2010 by 
+Rackspace and NASA, and currently developed and distributed by the OpenStack
+Foundation with the support of more than 150 relevant companies. OpenStack is a 
+cloud operating system that provides Infrastructure as a Service (IaaS) able to 
+control large pools of compute, storage, and networking resources throughout a 
+datacenter, all managed through a dashboard or an API that gives administrators 
+control while empowering their users to provision resources through a web interface
+or a command-line. It supports most available hypervisors and handles provisioning 
+and life-cycle management of VMs.
+
+\section{State of Art}
+\label{sec:state}
+Our solutions
+is based on GPU virtualization and sharing resources in order
+to reach a fair balance between supply and demand.
+Several attempts of achieving what we are pursing have been
+made, but none of them are so ambitious as our project. On
+the one hand the work done in~\cite{younge2013enabling} allows the VM managed
+by the hypervisor Xen to access the GPUs in the physical
+node. With the implied limitations of restricting the nodes not
+to use more GPUs than the hosted in the machine and not
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+them accessible for any VM in the cluster. However, this kind
+of virtualization strongly depends on the hypervisor used, so
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
+drawback is that the code of the applications must be modified
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+it is only focussed on compute intensive HPC applications.
+The main idea of our proposal goes further and apart from
+bringing solutions for all kind of HPC applications, it is aimed
+to boost the cluster flexibility in the use of GPUs.
+
+\section{Motivation}
+\label{sec:motivation}
+
+Cloud computing, storage and other facilities are becoming 
+more popular during the last years thanks to the easy set up 
+process and the low required IT invest. The features provided 
+by the cloud vendors are useful for companies and individuals 
+which have access to the last technology without spending a 
+huge amount of money. Moreover, the systems offered by the cloud 
+companies matches with the ones required thanks to its flexibility.
+
+The power of current cloud vendors is based on the use 
+of Virtual Machines (VM) which  
+emulates a physical node with its own 
+CPUs, storage, network interconnection, OS and GPU. 
+However, to provide a VM with GPUs means
+that the guest host must be equipped with at least one GPU, and
+once it has been assigned to an instance, no other VM can use
+it. Therefore, the number of VM that are allowed to use a GPU is 
+limited by the number of the installed devices. This lack of 
+flexibility is also noticed in the cluster configurations, where 
+depending on the users’ necessities, 
+two approaches to introduce GPGPU computation can be determined:
+the conservative, to have a lot of hosts with GPUs to be
+ensured that you are likely to satisfy all the requests; or a
+more daring approach, where you will count with some GPU-
+enabled nodes hoping not to arrive a large burst of requests
+so the cluster would run out of GPU resources.
+
+The use of a virtualization technology such as {rCUDA} adds 
+the desired flexibility because this software detaches the GPUs 
+from the nodes where they are installed allowing all nodes in a cluster 
+not only use them as if they were local but also to share them 
+achieving a higher utilization rate and reducing the idle status time.
+
+Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
+environment should be able to use and share all the GPUs installed in 
+the physical nodes removing the current limitation. But, in order to achieve this goal, 
+the cloud facilities need to be provided with the {rCUDA} software and 
+in the case of OpenStack, a module for manageing virtualized GPUs 
+ has to be implemented.
+
+\section{Working with AWS}
+\label{sec:aws}
+In this section, first of all, the current user-level features 
+provided by the Amazon Web Services are explained and analyzed 
+from the point of view of the GPU usage. 
+Finally, the scenarios and experiments where virtualized GPUs are utilized are 
+described and first results are showed.
+
+\subsection{Current Features}
+This subsection explains the most common features checked during 
+the virtual machine creation step.
+
+\subsubsection{Instances}
+
+An instance is a pre-configured virtual machine focused on an 
+specific target. Inside the large list of instances offered by AWS, 
+we can find: general pourpose (T2, M4 and M3); computer science optimized 
+(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
+GPU capable instances. Each kind of instance has its own pourpose and its 
+own price. Moreover, each instance has different number of CPUs and network 
+interconnection which can be: low, medium, high or 10Gb.
+
+For this study, the instances presented in the Table~\ref{table:awsInstances} 
+have been selected. The C3 family instances, which are not equipped with 
+GPUs are going to be used as client whereas the instances of the G2 family are 
+acting as servers.
+
+\begin{table}[htb]
+\renewcommand{\arraystretch}{1.3}
+\caption{AWS available HPC instances}
+\label{table:awsInstances}
+\tabcolsep=0.09cm
+\begin{center}\begin{tabular}{cccccc}
+Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
+c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.21\\ \hline
+c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
+g2.2xlarge & 8 & 15 GiB & High & 1 & \$ 0.65\\ \hline
+g2.8xlarge & 32 & 60 GiB & 10 Gb & 4 & \$ 2.6 \\ \hline
+\end{tabular}\end{center}\end{table}
+
+\subsubsection{Network behaviour}
+As it is shown in the Table~\ref{table:awsInstances}, each instance has different 
+interconnection and taking into account that the use of the network 
+is critical when GPU virtualization is applied, a simple test to verify the 
+real network bandwidth is needed.
+
+\begin{table}[htb]
+\renewcommand{\arraystretch}{1.3}
+\caption{IPERF results between instances}
+\label{table:iperf}
+\tabcolsep=0.24cm
+\begin{center}\begin{tabular}{cccc}
+Server & Client & Network & Bandwidth\\ \hline \hline
+g2.8xlarge & c3.2xlarge & High & 1  Gb/s\\ \hline
+g2.8xlarge & c3.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
+g2.8xlarge & g2.2xlarge & High & 1 Gb/s\\ \hline
+g2.8xlarge & g2.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
+\end{tabular}\end{center}\end{table}
+
+The IPERF tool has been executed between the instances showed in 
+the Table~\ref{table:iperf} in order to know the real bandwidth.
+Therefore, we can assure that the nomenclature ``High'' used in the 
+network column correspond to a 1Gb interconnection and the 10Gb has 
+a real bandwidth of 7.5Gb/s.
+Moreover, it looks like the instances equipped with a ``High'' interconnection 
+are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
+match perfectly. The gap between real and theorical can be observed with 
+the 10Gb interconnection which reaches up to 7.5 Gb/s.
+
+\subsubsection{GPUs}
+As it has been said before, an instance is based in a VM 
+which is running on a real node with its own virtualized 
+components, so it is possible that AWS could use a virtualization 
+framework to offer GPU services to all the instances.
+Therefore, we need to verify if the GPUs showed in the instances are 
+non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
+{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
+in this test is higher than the network bandwith, it reveals that it is a 
+real GPU. 
+\begin{table}[htb]
+\renewcommand{\arraystretch}{1.3}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory }
+\label{table:bwt}
+\tabcolsep=0.09cm
+\begin{center}\begin{tabular}{cccc}
+Name &  Data Movement & Network & BandWidth \\ \hline \hline
+g2.2xlarge & Host to Device & Hight& 3,004 MB/s \\ \hline
+g2.2xlarge & Device to Host & Hight& 2,809 MB/s\\ \hline
+g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
+g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
+\end{tabular}\end{center}\end{table}
+
+\subsection{Testbed Scenarios}
+In this section, the different client-server configurations 
+are explained. All the scenarios are based in the maximum number of instance 
+that a user can freely select without submiting a formal request.
+
+The first scenarion (Figure~\ref{subfig:aws1}) shows the most common 
+and used hardware configuration in real clusters, where each node is 
+populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
+using the ``High'' interconnection.
+
+The second configuration (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
+with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
+a 10Gb network and the cliend is able to execute application using up to 8 GPUs.
+
+The last scenario (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
+using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
+
+\begin{figure*}[ht]
+\centering
+\subfigure[5 remote GPUs over 1GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws3.png}
+\label{subfig:aws1}}
+\quad
+\subfigure[8 remote GPUs over 10GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws2.png}
+\label{subfig:aws2}}
+\subfigure[13 remote GPUs over 1GbE]{%
+\includegraphics[width=0.31\linewidth]{images/aws1.png}
+\label{subfig:aws3}}
+\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs}
+\label{fig:aws}
+\end{figure*}
+
+\subsection{Experimental results}
+In order to assess that GPU virtualization with AWS instances was working as expected, two applications were executed to analyze the results and their scalability.
+
+The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of~\ref{fig:aws} and on the instances with local GPUs. 
+We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
+In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
+Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
+However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
+unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
+
+% \begin{figure}[!t]
+%   \centering
+%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
+%   \caption{Montecarlo execution time}
+%   \label{fig:mont-exe}
+% \end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/aws-mont1.png}
+  \caption{Montecarlo options per second}
+  \label{fig:mont-opt}
+\end{figure}
+
+Flexibility and economical expenses
+
+\section{GPGPU Shared Service in OpenStack}
+In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
+\subsection{Managing remote GPUs with OpenStack}
+The main idea of this solution is to reach the scenario that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
+This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
+As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
+which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
+The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
+A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/os1.jpg}
+  \caption{Openstack with our GPGPU Service}
+  \label{fig:external}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/os2.png}
+  \caption{Internal Communication among modules}
+  \label{fig:internal}
+\end{figure}
+
+Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
+
+\subsection{Working Modes}
+The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
+The users are given two configuration options to decide whether a physical GPU will be completely allocated for an instance (first column) or the instance will address a partition of the GPU as if it were a device (second column).
+We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
+Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
+A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
+So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
+As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
+
+Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
+This behaviour is refereed as ``scope'' and determines that a group of instances is logically connected to a pool of GPUs.
+Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
+Again, the GPU pool can be made of ``exclusive'' GPUs or ``shared''.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=.5\textwidth]{images/workingmodes.jpg}
+  \caption{Examples of working modes}
+  \label{fig2}
+\end{figure}
+
+\subsection{User Interface}
+Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
+First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
+As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
+Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/UI-launch.pdf}
+  \caption{Launching Instances and assigning GPUs}
+  \label{fig:ui-launch}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/UI-rgpus.pdf}
+  \caption{GPU Information Panel}
+  \label{fig:ui-rgpus}
+\end{figure}
+
+\subsection{Experimental Results}
+The test were executed on a 5-node cluster where a machine
+equipped with a Intel Xeon E7420 quadcore processor at
+2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
+MHz, was guest host for the VMs. While the other 4 nodes
+were auxiliary servers with a Tesla C1090 GPU each one. 
+In the software stack the testbed system operates a
+Centos 6.6; the cloud platform is deployed by OpenStack; and
+the GPUs use CUDA 6.5. The GPU virtualization support is
+based on rCUDA v5.0.
+
+We have designed 6 different set-ups which can be divided
+in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
+“Exclusive” mode will provide, at most, 4 accelerators. Whilst
+the number of available GPU in “Shared” mode will depend
+on the partition size. In this particular case, we have halved
+the memory, so we can use up to 8 GPUs. For each group, we
+have deployed virtual clusters of 1, 2 and 3 nodes where will
+run the processes of the application. The instances are based
+on the flavor “m1.medium” which determines VM of 2 cores
+and 4096 MB of memory.
+
+MCUDA-MEME~\cite{Liu2010} has been the application executed with
+different configurations in order to test the set-ups. MCUDA-
+MEME is GPGPU MPI application, so the number of GPUs
+will determine the number of executable processes.
+The Figure~\ref{fig3} shows and compares the execution time of
+the application with different configurations over different set-
+ups. We can determine that with 1 node, the performance is
+lower, with 2 and 3 is quite similar, though. Hence, the greatest
+difference can be appreciated between modes, where the
+highest performance is achieved by sharing the accelerators.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{images/chart.jpg}
+  \caption{Execution time of MCUDA-MEME}
+  \label{fig3}
+\end{figure}
+
+\section{Discussion}
+\label{sec:discussion}
+-Volver a hacer un resumen
+-posibles futuras aplicaciones (lab)
+-pruebas con otros proveedores o con otras plataformas (azure)
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+\section{Future work}
+\label{sec:future}
+
+\section*{\uppercase{Acknowledgements}}
+The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
+
+The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
+TIN2011-23283 and~FEDER.
+
+\bibliographystyle{apalike}
+{\small
+\bibliography{paper}}
+
+\end{document}

### 2015-10-06 paper.tex
uppercase

@@
 \usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
+
+\usepackage{url}% http://ctan.org/pkg/url
 
@@
 
-\noindent
+%\noindent
 
@@
 \label{sec:openstack}
-\begin{itemize}
-\item Descripcion basica del openstack (Sisco?)
-\end{itemize}
+One of the most prominent open-source cloud computing platform for both private 
+and public clouds is OpenStack. The OpenStack project was announced in 2010 by 
+Rackspace and NASA, and currently developed and distributed by the OpenStack
+Foundation with the support of more than 150 relevant companies. OpenStack is a 
+cloud operating system that provides Infrastructure as a Service (IaaS) able to 
+control large pools of compute, storage, and networking resources throughout a 
+datacenter, all managed through a dashboard or an API that gives administrators 
+control while empowering their users to provision resources through a web interface
+or a command-line. It supports most available hypervisors and handles provisioning 
+and life-cycle management of VMs.
 
-\section{State of Art}
+\section{\uppercase{State of Art}}
 \label{sec:state}
@@
 
-\section{Motivation}
+\section{\uppercase{Motivation}}
 \label{sec:motivation}
@@
 
-\section{Working with AWS}
+\section{\uppercase{Working with AWS}}
 \label{sec:aws}
@@
 
-\section{GPGPU Shared Service in OpenStack}
+\section{\uppercase{GPGPU Shared Service in OpenStack}}
 In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
@@
 
-\section{Discussion}
+\section{\uppercase{Discussion}}
 \label{sec:discussion}
@@
 
-\section{Conclusions}
+\section{\uppercase{Conclusions}}
 \label{sec:conclusions}
 
-\section{Future work}
+\section{\uppercase{Future work}}
 \label{sec:future}
@@
 \end{document}
-
-=======
-\documentclass[a4paper,twoside]{article}
-
-\usepackage{epsfig}
-\usepackage{subfigure}
-\usepackage{calc}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{multicol}
-\usepackage{pslatex}
-\usepackage{apalike}
-\usepackage{SCITEPRESS}     % Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
-
-\subfigtopskip=0pt
-\subfigcapskip=0pt
-\subfigbottomskip=0pt
-
-\begin{document}
-
-\title{Enabling the Use of GPU Virtualization in Cloud Environments}
-\author{\authorname{Sergio Iserte, Francisco J. Clemente-Castell\'o, Adri\'an Castell\'o,\\Rafael Mayo and Enrique S. Quintana-Ort\'i}
-\affiliation{Department of Computer Science and Engineering\\Universitat Jaume I - Castell\'o de la Plana, Spain}
-\email{\{siserte, fclement, adcastel, mayo, quintana\}@uji.es}}
-
-\keywords{Cloud Computing, GPU Virtualization, Resource Management}
-
-\abstract{The use of accelerators in order to reduce the execution time of some 
-application has become more popular during last years. These devices increment 
-the computational power of a node thanks to its parallel architecture. %The popularity of the GPGPU computation 
-This trend has led several cloud vendors such as Amazon Web Services or some 
-middlewares such as OpenStack to add to their current facilities, or to allow, 
-instances of virtual machines equipped with NVIDIA GPUs. 
-Accomplishing this requirements needs that the guest hosts must be equipped with 
-GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
-The solution presented in this work is based on GPU virtualization and shareability 
-in order to reach an equilibrium between supply and demand of accelerators from the users. 
-Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
-With this software configuration, these GPUs can be accessed from any VM getting rid 
-of the necessity of having physical GPUs in the guest host. 
-Moreover, our project studies the viability of these join by using a public cloud 
-service configuration and develops a module for OpenStack middleware in order to 
-add the needed support for the virtualized devices and the logic to manage them. 
-The results show a feasible configuration which adds flexibility to the current and well-known 
-cloud solutions. 
-}
-
-\onecolumn \maketitle \normalsize \vfill
-
-\section{\uppercase{Introduction}}
-\label{sec:introduction}
-
-\noindent
-
-\section{\uppercase{Background}}
-\label{sec:background}
-\subsection{The rCUDA Framework}
-\label{sec:rcuda}
-
-{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
-to any NVIDIA GPU device present in a cluster from all compute
-nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
-as if they were local.
-These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
-resource, space, and energy~\cite{energy14} requirements.
-rCUDA is structured following a client-server distributed
-architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
-acceleration services is executed, providing a transparent replacement for the
-native CUDA libraries. Furthermore, the server middleware is executed in the
-cluster nodes from which the actual GPUs provide the requested GPGPU service.
-To support a concurrent scenario where GPUs are shared between
-processes\slash nodes, {rCUDA} manages separate device contexts for
-each client application.
-
-The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
-CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
-APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
-Therefore, applications are not aware of the fact that they are being executed
-on top of a virtualization layer.
-With the aim to be updated with new GPU programming models, {rCUDA} also supports
-directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.
-
-The integration of remote GPGPU virtualization with global
-resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
-technology, making accelerator-enabled clusters more flexible and
-energy efficient.
-
-\subsection{Amazon Web Services}
-\label{sec:aws}
-AWS~\cite{aws} is a public cloud computing provider
-and it is composed by several cloud services such as 
- cloud-based computation, storage and other functionality 
-that enables organizations and/or individuals to deploy
-services and applications on an on-demand basis. 
-These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
-perfectly with their software requirements.
-
-From the point of view of HPC, AWS offers high performance facilities including 
-instances equipped with GPUs and high performance network interconnection which 
-are common resources used in this research field.
-
-\subsection{OpenStack}
-\label{sec:openstack}
-
-One of the most prominent open-source cloud computing platform for both private 
-and public clouds is OpenStack. The OpenStack project was announced in 2010 by 
-Rackspace and NASA, and currently developed and distributed by the OpenStack
-Foundation with the support of more than 150 relevant companies. OpenStack is a 
-cloud operating system that provides Infrastructure as a Service (IaaS) able to 
-control large pools of compute, storage, and networking resources throughout a 
-datacenter, all managed through a dashboard or an API that gives administrators 
-control while empowering their users to provision resources through a web interface
-or a command-line. It supports most available hypervisors and handles provisioning 
-and life-cycle management of VMs.
-
-\section{State of Art}
-\label{sec:state}
-Our solutions
-is based on GPU virtualization and sharing resources in order
-to reach a fair balance between supply and demand.
-Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project. On
-the one hand the work done in~\cite{younge2013enabling} allows the VM managed
-by the hypervisor Xen to access the GPUs in the physical
-node. With the implied limitations of restricting the nodes not
-to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented
-by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
-them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
-integrated in a Cloud Computing Manager, though its main
-drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. On the
-other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
-it is only focussed on compute intensive HPC applications.
-The main idea of our proposal goes further and apart from
-bringing solutions for all kind of HPC applications, it is aimed
-to boost the cluster flexibility in the use of GPUs.
-
-\section{Motivation}
-\label{sec:motivation}
-
-Cloud computing, storage and other facilities are becoming 
-more popular during the last years thanks to the easy set up 
-process and the low required IT invest. The features provided 
-by the cloud vendors are useful for companies and individuals 
-which have access to the last technology without spending a 
-huge amount of money. Moreover, the systems offered by the cloud 
-companies matches with the ones required thanks to its flexibility.
-
-The power of current cloud vendors is based on the use 
-of Virtual Machines (VM) which  
-emulates a physical node with its own 
-CPUs, storage, network interconnection, OS and GPU. 
-However, to provide a VM with GPUs means
-that the guest host must be equipped with at least one GPU, and
-once it has been assigned to an instance, no other VM can use
-it. Therefore, the number of VM that are allowed to use a GPU is 
-limited by the number of the installed devices. This lack of 
-flexibility is also noticed in the cluster configurations, where 
-depending on the users’ necessities, 
-two approaches to introduce GPGPU computation can be determined:
-the conservative, to have a lot of hosts with GPUs to be
-ensured that you are likely to satisfy all the requests; or a
-more daring approach, where you will count with some GPU-
-enabled nodes hoping not to arrive a large burst of requests
-so the cluster would run out of GPU resources.
-
-The use of a virtualization technology such as {rCUDA} adds 
-the desired flexibility because this software detaches the GPUs 
-from the nodes where they are installed allowing all nodes in a cluster 
-not only use them as if they were local but also to share them 
-achieving a higher utilization rate and reducing the idle status time.
-
-Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
-environment should be able to use and share all the GPUs installed in 
-the physical nodes removing the current limitation. But, in order to achieve this goal, 
-the cloud facilities need to be provided with the {rCUDA} software and 
-in the case of OpenStack, a module for manageing virtualized GPUs 
- has to be implemented.
-
-\section{Working with AWS}
-\label{sec:aws}
-In this section, first of all, the current user-level features 
-provided by the Amazon Web Services are explained and analyzed 
-from the point of view of the GPU usage. 
-Finally, the scenarios and experiments where virtualized GPUs are utilized are 
-described and first results are showed.
-
-\subsection{Current Features}
-This subsection explains the most common features checked during 
-the virtual machine creation step.
-
-\subsubsection{Instances}
-
-An instance is a pre-configured virtual machine focused on an 
-specific target. Inside the large list of instances offered by AWS, 
-we can find: general pourpose (T2, M4 and M3); computer science optimized 
-(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable instances. Each kind of instance has its own pourpose and its 
-own price. Moreover, each instance has different number of CPUs and network 
-interconnection which can be: low, medium, high or 10Gb.
-
-For this study, the instances presented in the Table~\ref{table:awsInstances} 
-have been selected. The C3 family instances, which are not equipped with 
-GPUs are going to be used as client whereas the instances of the G2 family are 
-acting as servers.
-
-\begin{table}[htb]
-\renewcommand{\arraystretch}{1.3}
-\caption{AWS available HPC instances}
-\label{table:awsInstances}
-\tabcolsep=0.09cm
-\begin{center}\begin{tabular}{cccccc}
-Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
-c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.21\\ \hline
-c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
-g2.2xlarge & 8 & 15 GiB & High & 1 & \$ 0.65\\ \hline
-g2.8xlarge & 32 & 60 GiB & 10 Gb & 4 & \$ 2.6 \\ \hline
-\end{tabular}\end{center}\end{table}
-
-\subsubsection{Network behaviour}
-As it is shown in the Table~\ref{table:awsInstances}, each instance has different 
-interconnection and taking into account that the use of the network 
-is critical when GPU virtualization is applied, a simple test to verify the 
-real network bandwidth is needed.
-
-\begin{table}[htb]
-\renewcommand{\arraystretch}{1.3}
-\caption{IPERF results between instances}
-\label{table:iperf}
-\tabcolsep=0.24cm
-\begin{center}\begin{tabular}{cccc}
-Server & Client & Network & Bandwidth\\ \hline \hline
-g2.8xlarge & c3.2xlarge & High & 1  Gb/s\\ \hline
-g2.8xlarge & c3.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
-g2.8xlarge & g2.2xlarge & High & 1 Gb/s\\ \hline
-g2.8xlarge & g2.8xlarge & 10Gb & 7.5  Gb/s\\ \hline
-\end{tabular}\end{center}\end{table}
-
-The IPERF tool has been executed between the instances showed in 
-the Table~\ref{table:iperf} in order to know the real bandwidth.
-Therefore, we can assure that the nomenclature ``High'' used in the 
-network column correspond to a 1Gb interconnection and the 10Gb has 
-a real bandwidth of 7.5Gb/s.
-Moreover, it looks like the instances equipped with a ``High'' interconnection 
-are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
-match perfectly. The gap between real and theorical can be observed with 
-the 10Gb interconnection which reaches up to 7.5 Gb/s.
-
-\subsubsection{GPUs}
-As it has been said before, an instance is based in a VM 
-which is running on a real node with its own virtualized 
-components, so it is possible that AWS could use a virtualization 
-framework to offer GPU services to all the instances.
-Therefore, we need to verify if the GPUs showed in the instances are 
-non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
-in this test is higher than the network bandwith, it reveals that it is a 
-real GPU. 
-\begin{table}[htb]
-\renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory }
-\label{table:bwt}
-\tabcolsep=0.09cm
-\begin{center}\begin{tabular}{cccc}
-Name &  Data Movement & Network & BandWidth \\ \hline \hline
-g2.2xlarge & Host to Device & Hight& 3,004 MB/s \\ \hline
-g2.2xlarge & Device to Host & Hight& 2,809 MB/s\\ \hline
-g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
-g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
-\end{tabular}\end{center}\end{table}
-
-\subsection{Testbed Scenarios}
-In this section, the different client-server configurations 
-are explained. All the scenarios are based in the maximum number of instance 
-that a user can freely select without submiting a formal request.
-
-The first scenarion (Figure~\ref{subfig:aws1}) shows the most common 
-and used hardware configuration in real clusters, where each node is 
-populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
-using the ``High'' interconnection.
-
-The second configuration (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
-with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
-a 10Gb network and the cliend is able to execute application using up to 8 GPUs.
-
-The last scenario (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
-using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
-
-\begin{figure*}[ht]
-\centering
-\subfigure[5 remote GPUs over 1GbE]{%
-\includegraphics[width=0.31\linewidth]{images/aws3.png}
-\label{subfig:aws1}}
-\quad
-\subfigure[8 remote GPUs over 10GbE]{%
-\includegraphics[width=0.31\linewidth]{images/aws2.png}
-\label{subfig:aws2}}
-\subfigure[13 remote GPUs over 1GbE]{%
-\includegraphics[width=0.31\linewidth]{images/aws1.png}
-\label{subfig:aws3}}
-\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs}
-\label{fig:aws}
-\end{figure*}
-
-\subsection{Experimental results}
-In order to assess that GPU virtualization with AWS instances was working as expected, two applications were executed to analyze the results and their scalability.
-
-The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
-The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of~\ref{fig:aws} and on the instances with local GPUs. 
-We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
-In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
-Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
-However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
-unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
-
-% \begin{figure}[!t]
-%   \centering
-%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
-%   \caption{Montecarlo execution time}
-%   \label{fig:mont-exe}
-% \end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{images/aws-mont1.png}
-  \caption{Montecarlo options per second}
-  \label{fig:mont-opt}
-\end{figure}
-
-Flexibility and economical expenses
-
-\section{GPGPU Shared Service in OpenStack}
-In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
-\subsection{Managing remote GPUs with OpenStack}
-The main idea of this solution is to reach the scenario that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
-This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
-As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
-which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
-The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
-A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{images/os1.jpg}
-  \caption{Openstack with our GPGPU Service}
-  \label{fig:external}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{images/os2.png}
-  \caption{Internal Communication among modules}
-  \label{fig:internal}
-\end{figure}
-
-Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
-
-\subsection{Working Modes}
-The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
-The users are given two configuration options to decide whether a physical GPU will be completely allocated for an instance (first column) or the instance will address a partition of the GPU as if it were a device (second column).
-We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
-Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
-A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
-So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
-As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
-
-Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
-This behaviour is refereed as ``scope'' and determines that a group of instances is logically connected to a pool of GPUs.
-Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
-Again, the GPU pool can be made of ``exclusive'' GPUs or ``shared''.
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=.5\textwidth]{images/workingmodes.jpg}
-  \caption{Examples of working modes}
-  \label{fig2}
-\end{figure}
-
-\subsection{User Interface}
-Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
-First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
-Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{images/UI-launch.pdf}
-  \caption{Launching Instances and assigning GPUs}
-  \label{fig:ui-launch}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{images/UI-rgpus.pdf}
-  \caption{GPU Information Panel}
-  \label{fig:ui-rgpus}
-\end{figure}
-
-\subsection{Experimental Results}
-The test were executed on a 5-node cluster where a machine
-equipped with a Intel Xeon E7420 quadcore processor at
-2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
-MHz, was guest host for the VMs. While the other 4 nodes
-were auxiliary servers with a Tesla C1090 GPU each one. 
-In the software stack the testbed system operates a
-Centos 6.6; the cloud platform is deployed by OpenStack; and
-the GPUs use CUDA 6.5. The GPU virtualization support is
-based on rCUDA v5.0.
-
-We have designed 6 different set-ups which can be divided
-in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
-“Exclusive” mode will provide, at most, 4 accelerators. Whilst
-the number of available GPU in “Shared” mode will depend
-on the partition size. In this particular case, we have halved
-the memory, so we can use up to 8 GPUs. For each group, we
-have deployed virtual clusters of 1, 2 and 3 nodes where will
-run the processes of the application. The instances are based
-on the flavor “m1.medium” which determines VM of 2 cores
-and 4096 MB of memory.
-
-MCUDA-MEME~\cite{Liu2010} has been the application executed with
-different configurations in order to test the set-ups. MCUDA-
-MEME is GPGPU MPI application, so the number of GPUs
-will determine the number of executable processes.
-The Figure~\ref{fig3} shows and compares the execution time of
-the application with different configurations over different set-
-ups. We can determine that with 1 node, the performance is
-lower, with 2 and 3 is quite similar, though. Hence, the greatest
-difference can be appreciated between modes, where the
-highest performance is achieved by sharing the accelerators.
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{images/chart.jpg}
-  \caption{Execution time of MCUDA-MEME}
-  \label{fig3}
-\end{figure}
-
-\section{Discussion}
-\label{sec:discussion}
--Volver a hacer un resumen
--posibles futuras aplicaciones (lab)
--pruebas con otros proveedores o con otras plataformas (azure)
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-\section{Future work}
-\label{sec:future}
-
-\section*{\uppercase{Acknowledgements}}
-The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
-
-The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
-TIN2011-23283 and~FEDER.
-
-\bibliographystyle{apalike}
-{\small
-\bibliography{paper}}
-
-\end{document}

### 2015-10-07 paper.tex
Background de openstack modificado

@@
 \label{sec:openstack}
-One of the most prominent open-source cloud computing platform for both private 
-and public clouds is OpenStack. The OpenStack project was announced in 2010 by 
-Rackspace and NASA, and currently developed and distributed by the OpenStack
-Foundation with the support of more than 150 relevant companies. OpenStack is a 
-cloud operating system that provides Infrastructure as a Service (IaaS) able to 
-control large pools of compute, storage, and networking resources throughout a 
-datacenter, all managed through a dashboard or an API that gives administrators 
+
+OpenStack is a cloud operating system that provides Infrastructure as a Service (IaaS). 
+It is able to control large pools of compute, storage, and networking resources throughout a 
+datacenter. All this resources are managed through a dashboard or an API that gives administrators 
 control while empowering their users to provision resources through a web interface
-or a command-line. It supports most available hypervisors and handles provisioning 
+or a command-line. 
+It supports most available hypervisors and handles provisioning 
 and life-cycle management of VMs.

### 2015-10-07 paper.tex
anotaciones

--- 2015-10-06 paper.tex
@@
 Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
-Thus, if we compare the results of a LAMMPS execution with a larger problem (see Figure), while CUDA performs better, the interesting point is the execution time on remote GPUs. 
+Thus, if we compare the results of a LAMMPS execution with a larger problem (see Figure), although CUDA continues performing better, the interesting point is the execution time on remote GPUs. 
 They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
-pays off the drawback of a slow network.
+pays off the drawback of a slower network.
 
@@
 \subfigure[first]{%
-\includegraphics[width=0.455\linewidth]{images/lammps100.png}
+\includegraphics[width=\linewidth]{images/lammps100.png}
 \label{subfig:lammps1}}
@@
 \subfigure[second]{%
-\includegraphics[width=0.455\linewidth]{images/lammps2000.png}
+\includegraphics[width=\linewidth]{images/lammps2000.png}
 \label{subfig:lammps2}}
@@
 In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
+
+Version de open stack con la que se trabaja y tipo de instalacion ya que está toda en un nodo...
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{images/os-orig.png}
+  \caption{Openstack Architecture}
+  \label{fig:os-orig}
+\end{figure}
+
 \subsection{Managing remote GPUs with OpenStack}

### 2015-10-07 paper.tex
Ancho de banda con AWS y rCUDA añadido

--- 2015-10-07 paper.tex
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{AWS available HPC instances}
+\caption{AWS available HPC instances.}
 \label{table:awsInstances}
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{IPERF results between instances}
+\caption{IPERF results between instances.}
 \label{table:iperf}
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory }
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using local GPU.}
 \label{table:bwt}
@@
 \centering
-\subfigure[5 remote GPUs over 1GbE]{%
+\subfigure[5 remote GPUs over 1GbE.]{%
 \includegraphics[width=0.31\linewidth]{images/aws3.png}
@@
 \quad
-\subfigure[8 remote GPUs over 10GbE]{%
+\subfigure[8 remote GPUs over 10GbE.]{%
 \includegraphics[width=0.31\linewidth]{images/aws2.png}
 \label{subfig:aws2}}
-\subfigure[13 remote GPUs over 1GbE]{%
+\subfigure[13 remote GPUs over 1GbE.]{%
 \includegraphics[width=0.31\linewidth]{images/aws1.png}
 \label{subfig:aws3}}
-\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs}
+\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
 \label{fig:aws}
 \end{figure*}
+
+Once the scenarios are configured from the point of view of hardware, 
+the {rCUDA} software needs to be installed in order to add 
+the required flexibility to the system. So, the {rCUDA} server has been 
+executed in the server nodes and the {rCUDA} libraries have been installed 
+in the node that is acting as a client.
+
+Moreover, with the aim to know the expected network bandwidth using 
+a remote GPU, we have executed again the NVIDIA {\tt bandwidthtest} sample.
+The results observed in the Table~\ref{} shows how the bandwidth achieved is limited 
+by the network.
+
+\begin{table}[htb]
+\renewcommand{\arraystretch}{1.3}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using {rCUDA}.}
+\label{table:bwt}
+\tabcolsep=0.09cm
+\begin{center}\begin{tabular}{cccc}
+Scenario &  Data Movement & Network & BandWidth \\ \hline \hline
+A & Host to Device & Hight& 127 MB/s \\ \hline
+A & Device to Host & Hight& 126 MB/s\\ \hline
+B & Host to Device & 10 Gb& 858 MB/s\\ \hline
+B & Device to Host & 10 Gb& 843 MB/s\\ \hline
+\end{tabular}\end{center}\end{table}
 
@@
 
-The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The first of them was {\tt MonteCarloMultigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
@@
   \includegraphics[width=\linewidth]{images/aws-mont1.png}
-  \caption{Montecarlo options per second}
+  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
   \label{fig:mont-opt}

### 2015-10-07 paper.tex
Puntuacion añadida en las figuras
Escenarios en AWS como lista

@@
 
-The first scenarion (Figure~\ref{subfig:aws1}) shows the most common 
+\begin{itemize}
+\item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
 and used hardware configuration in real clusters, where each node is 
@@
 
-The second configuration (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
+\item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
 with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
@@
 
-The last scenario (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
+\item Scenario C (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
 using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
+\end{itemize}
 
@@
 a remote GPU, we have executed again the NVIDIA {\tt bandwidthtest} sample.
-The results observed in the Table~\ref{} shows how the bandwidth achieved is limited 
+The results observed in the Table~\ref{table:bwtrcuda} shows how the bandwidth achieved is limited 
 by the network.
@@
 \caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using {rCUDA}.}
-\label{table:bwt}
+\label{table:bwtrcuda}
 \tabcolsep=0.09cm
@@
 \label{subfig:lammps2}}
-\caption{Execution time of LAMMPS}
+\caption{Execution time of LAMMPS.}
 \label{fig:lammps}
@@
   \includegraphics[width=\linewidth]{images/os-orig.png}
-  \caption{Openstack Architecture}
+  \caption{Openstack Architecture.}
   \label{fig:os-orig}
@@
   \includegraphics[width=\linewidth]{images/os1.jpg}
-  \caption{Openstack with our GPGPU Service}
+  \caption{Openstack with our GPGPU Service.}
   \label{fig:external}
@@
   \includegraphics[width=\linewidth]{images/os2.png}
-  \caption{Internal Communication among modules}
+  \caption{Internal Communication among modules.}
   \label{fig:internal}
@@
   \includegraphics[width=.5\textwidth]{images/workingmodes.jpg}
-  \caption{Examples of working modes}
+  \caption{Examples of working modes.}
   \label{fig2}
@@
   \includegraphics[width=\linewidth]{images/UI-launch.pdf}
-  \caption{Launching Instances and assigning GPUs}
+  \caption{Launching Instances and assigning GPUs.}
   \label{fig:ui-launch}
@@
   \includegraphics[width=\linewidth]{images/UI-rgpus.pdf}
-  \caption{GPU Information Panel}
+  \caption{GPU Information Panel.}
   \label{fig:ui-rgpus}
@@
   \includegraphics[width=0.5\textwidth]{images/chart.jpg}
-  \caption{Execution time of MCUDA-MEME}
+  \caption{Execution time of MCUDA-MEME.}
   \label{fig3}
@@
 The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
-
 The researchers were supported by Universitat Jaume I research project (P11B2013-21), project

### 2015-10-07 paper.tex
Escenarios en AWS como lista

@@
 
-\item Scenario C (Figure~\ref{subfig:aws3}), joins both previous ones and a single client, 
-using a 1Gb network interconnection has 13 GPUs that can be used as if they were locally installed.
+\item Scenario C (Figure~\ref{subfig:aws3}) joins both scenarios A and B. So a 
+single client, using a 1Gb network interconnection, has 13 GPUs that can be used as if they were locally installed.
 \end{itemize}

### 2015-10-07 paper.tex
lammps charts

--- 2015-10-07 paper.tex
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{AWS available HPC instances.}
+\caption{AWS available HPC instances}
 \label{table:awsInstances}
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{IPERF results between instances.}
+\caption{IPERF results between instances}
 \label{table:iperf}
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using local GPU.}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory }
 \label{table:bwt}
@@
 \centering
-\subfigure[5 remote GPUs over 1GbE.]{%
+\subfigure[5 remote GPUs over 1GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws3.png}
@@
 \quad
-\subfigure[8 remote GPUs over 10GbE.]{%
+\subfigure[8 remote GPUs over 10GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws2.png}
 \label{subfig:aws2}}
-\subfigure[13 remote GPUs over 1GbE.]{%
+\subfigure[13 remote GPUs over 1GbE]{%
 \includegraphics[width=0.31\linewidth]{images/aws1.png}
 \label{subfig:aws3}}
-\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
+\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs}
 \label{fig:aws}
 \end{figure*}
-
-Once the scenarios are configured from the point of view of hardware, 
-the {rCUDA} software needs to be installed in order to add 
-the required flexibility to the system. So, the {rCUDA} server has been 
-executed in the server nodes and the {rCUDA} libraries have been installed 
-in the node that is acting as a client.
-
-Moreover, with the aim to know the expected network bandwidth using 
-a remote GPU, we have executed again the NVIDIA {\tt bandwidthtest} sample.
-The results observed in the Table~\ref{} shows how the bandwidth achieved is limited 
-by the network.
-
-\begin{table}[htb]
-\renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using {rCUDA}.}
-\label{table:bwt}
-\tabcolsep=0.09cm
-\begin{center}\begin{tabular}{cccc}
-Scenario &  Data Movement & Network & BandWidth \\ \hline \hline
-A & Host to Device & Hight& 127 MB/s \\ \hline
-A & Device to Host & Hight& 126 MB/s\\ \hline
-B & Host to Device & 10 Gb& 858 MB/s\\ \hline
-B & Device to Host & 10 Gb& 843 MB/s\\ \hline
-\end{tabular}\end{center}\end{table}
 
@@
 
-The first of them was {\tt MonteCarloMultigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The first of them was {\it Montecarlo Multigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
@@
   \includegraphics[width=\linewidth]{images/aws-mont1.png}
-  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
+  \caption{Montecarlo execution results}
   \label{fig:mont-opt}
@@
 \centering
-\subfigure[first]{%
-\includegraphics[width=\linewidth]{images/lammps100.png}
+\subfigure[Run size = 100.]{%
+\includegraphics[width=\linewidth]{images/lammps1.pdf}
 \label{subfig:lammps1}}
 \quad
-\subfigure[second]{%
-\includegraphics[width=\linewidth]{images/lammps2000.png}
+\subfigure[Run size = 2000.]{%
+\includegraphics[width=\linewidth]{images/lammps2.pdf}
 \label{subfig:lammps2}}
@@
 
+\subsection{Managing remote GPUs with OpenStack}
+The main idea is to move from the original Openstack architecture (see Figure~\ref{fig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Figure~\ref{fig:external}).
+This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
+As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
+which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
+The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
+A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
+
 \begin{figure}[htb]
@@
 \end{figure}
-
-\subsection{Managing remote GPUs with OpenStack}
-The main idea of this solution is to reach the scenario that could be seen in Figure~\ref{fig:external}, where a new shared service takes part in to the OpenStack Architecture.
-This new service is in charge of GPGPU computation, bringing more flexibility when it comes to GPUs.
-As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
-which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
-The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
-A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
 

### 2015-10-07 paper.tex
Ampliacion OpenStack

@@
 
-OpenStack is a cloud operating system that provides Infrastructure as a Service (IaaS). 
+OpenStack \cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
 It is able to control large pools of compute, storage, and networking resources throughout a 
@@
 and life-cycle management of VMs. 
+The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
+or software requirements and the ability to integrate with legacy systems and third party technologies.
+
+From the HPC perspective, OpenStack offers high performance computing (HPC) configurations using
+multiple supported hypervisors and different hardware architectures. However, despite to be able to
+specify some extra parameters for special resources like GPUs, choosing a host based on the existence
+of a GPU is currently unsupported in OpenStack \cite{OpenStackGPU}.
 

### 2015-10-07 paper.tex
lammps charts merge solved

--- 2015-10-07 paper.tex
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea is to move from the original Openstack architecture (see Figure~\ref{fig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Figure~\ref{fig:external}).
+The main idea is to move from the original Openstack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
 This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
@@
   \centering
+  \subfigure[Original in version XXXXXXXXXXXXXXXXXXXXXX]{
     \includegraphics[width=\linewidth]{images/os-orig.png}
-  \caption{Openstack Architecture.}
-  \label{fig:os-orig}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
+    \label{subfig:os-orig}
+   }
+   \quad
+  \subfigure[With GPGPU developed module]{
     \includegraphics[width=\linewidth]{images/os1.jpg}
-  \caption{Openstack with our GPGPU Service.}
-  \label{fig:external}
+    \label{subfig:os-mod}
+   }
+  \caption{Openstack Architecture}
+  \label{fig:os}
 \end{figure}

### 2015-10-07 paper.tex
OpenStack ampliacion

@@
 and life-cycle management of VMs. 
+The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
+or software requirements and the ability to integrate with legacy systems and third party technologies.
+
+From the HPC perspective, OpenStack offers high performance computing (HPC) configurations using
+multiple supported hypervisors and different hardware architectures. However, despite to be able to
+specify some extra parameters for special resources like GPUs, chooding a host based on the existence
+of a GPU is currently unsupported in OpenStack \cite{OpenStackGPU}.
 

### 2015-10-07 paper.tex
OpenStack ampliacion

@@
 multiple supported hypervisors and different hardware architectures. However, despite to be able to
-specify some extra parameters for special resources like GPUs, chooding a host based on the existence
+specify some extra parameters for special resources like GPUs, choosing a host based on the existence
 of a GPU is currently unsupported in OpenStack \cite{OpenStackGPU}.

### 2015-10-07 paper.tex
fichero de compilacion añadido

@@
 
-OpenStack \cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
+OpenStack~\cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
 It is able to control large pools of compute, storage, and networking resources throughout a 
-datacenter. All this resources are managed through a dashboard or an API that gives administrators 
+datacenter. All these resources are managed through a dashboard or an API that gives administrators 
 control while empowering their users to provision resources through a web interface
@@
 and life-cycle management of VMs.
+
 The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
@@
 specify some extra parameters for special resources like GPUs, choosing a host based on the existence
-of a GPU is currently unsupported in OpenStack \cite{OpenStackGPU}.
+of a GPU is currently unsupported in OpenStack~\cite{OpenStackGPU}.
 

### 2015-10-07 paper.tex
mcudameme chart

--- 2015-10-07 paper.tex
@@
 \usepackage{url}% http://ctan.org/pkg/url
+\usepackage{flushend}
 
@@
 
-OpenStack \cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
+OpenStack is a cloud operating system that provides Infrastructure as a Service (IaaS). 
 It is able to control large pools of compute, storage, and networking resources throughout a 
@@
 and life-cycle management of VMs.
-The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
-or software requirements and the ability to integrate with legacy systems and third party technologies.
-
-From the HPC perspective, OpenStack offers high performance computing (HPC) configurations using
-multiple supported hypervisors and different hardware architectures. However, despite to be able to
-specify some extra parameters for special resources like GPUs, choosing a host based on the existence
-of a GPU is currently unsupported in OpenStack \cite{OpenStackGPU}.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{images/aws-mont1.png}
+  \includegraphics[width=\linewidth]{images/mont.pdf}
   \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{images/chart.jpg}
-  \caption{Execution time of MCUDA-MEME.}
+  \includegraphics[width=\linewidth]{images/mcudameme-os.pdf}
+  \caption{Scalability results of executing MCUDAMEME with a different number of MPI processes}
   \label{fig3}
@@
 The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
+
 The researchers were supported by Universitat Jaume I research project (P11B2013-21), project

### 2015-10-07 paper.tex
fichero de compilacion añadido

--- 2015-10-07 paper.tex
@@
 Name &  Data Movement & Network & BandWidth \\ \hline \hline
-g2.2xlarge & Host to Device & Hight& 3,004 MB/s \\ \hline
-g2.2xlarge & Device to Host & Hight& 2,809 MB/s\\ \hline
+g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
+g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
 g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
@@
 Scenario &  Data Movement & Network & BandWidth \\ \hline \hline
-A & Host to Device & Hight& 127 MB/s \\ \hline
-A & Device to Host & Hight& 126 MB/s\\ \hline
+A & Host to Device & High& 127 MB/s \\ \hline
+A & Device to Host & High& 126 MB/s\\ \hline
 B & Host to Device & 10 Gb& 858 MB/s\\ \hline

### 2015-10-07 paper.tex
future work

--- 2015-10-07 paper.tex
@@
   \centering
-  \subfigure[Original in version XXXXXXXXXXXXXXXXXXXXXX]{
+  \subfigure[Original in version XXXXXXXXXXXX]{
     \includegraphics[width=\linewidth]{images/os-orig.png}
@@
 \label{sec:future}
+The first task that we should have keep in mind after this study is to upgrade the interconnection network to one more focused on HPC.
+So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
+
+Looking for real situations where the performance is not so important as the flexibility to provide more accelerators. 
+Because of that, to take the first steps through how to deploy programming computer labs easily, could be really interesting.
+
+Finally, to bring in more policies to decide the strategy of creating and assigning remote GPUs in the physical device, would be a task to increase the flexibility given by the GPGPU module for Openstack.
 

### 2015-10-07 paper.tex
future work

--- 2015-10-07 paper.tex
@@
 \label{sec:future}
+The first task that we should have keep in mind after this study is to upgrade the interconnection network to one more focused on HPC.
+So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
+
+Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
+would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPU.
+
+Finally, to bring in more strategies to decide where a remote GPUs was created and assigned to a physical device, would be an interesting task to carry out. 
+With more scheduling policies the flexibility given by the GPGPU module for Openstack should increase.
 

### 2015-10-07 paper.tex
mcudameme better explained

--- 2015-10-07 paper.tex
@@
 From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
-The Figure shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
+The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
 Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
-Thus, if we compare the results of a LAMMPS execution with a larger problem (see Figure), although CUDA continues performing better, the interesting point is the execution time on remote GPUs. 
+Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs.
 They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
@@
 \subsection{Experimental Results}
-The test were executed on a 5-node cluster where a machine
+The tests were executed on a 5-node cluster where a machine
 equipped with a Intel Xeon E7420 quadcore processor at
@@
 different configurations in order to test the set-ups. MCUDA-
-MEME is GPGPU MPI application, so the number of GPUs
-will determine the number of executable processes.
+MEME is an MPI application which is accelerated by GPUs
+Besides, each of its processes must have a GPU to not abort 
+So, the number of GPUs will determine the number of processes that we can launch.
+
 The Figure~\ref{fig3} shows and compares the execution time of
-the application with different configurations over different set-
-ups. We can determine that with 1 node, the performance is
-lower, with 2 and 3 is quite similar, though. Hence, the greatest
-difference can be appreciated between modes, where the
-highest performance is achieved by sharing the accelerators.
+the application with different configurations over different setups. 
+We have used up to 3 nodes to spread the processes and launched up to 8 processes (only 4 in exclusive mode), one for each remote GPU.
+Firstly, we can state that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
+Secondly, the lowest execution time is obtained by both modes (exclusive and shared) when running their maximum number of processes with more than 1 node.
+It could mean that it is not worthy to continue giving resources to the application to scale, because the performance growth rate is barely increasing.
+Although, the time is lower sharing the GPUs the setup cannot take advantage of doubling the devices.
 
@@
 -Volver a hacer un resumen
--posibles futuras aplicaciones (lab)
--pruebas con otros proveedores o con otras plataformas (azure)
 
@@
 So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
+The same thought, would guide us to try other Cloud vendors which more support for HPC.
 
-Looking for real situations where the performance is not so important as the flexibility to provide more accelerators. 
-Because of that, to take the first steps through how to deploy programming computer labs easily, could be really interesting.
+Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
+would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPUs.
 
-Finally, to bring in more policies to decide the strategy of creating and assigning remote GPUs in the physical device, would be a task to increase the flexibility given by the GPGPU module for Openstack.
+Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
+With more scheduling policies the flexibility given by the GPGPU module for Openstack would increase.
 

### 2015-10-07 paper.tex
aws discussion

@@
 \subsection{Discussion}
-Flexibility and economical expenses
+Putting all together,  it is cristal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
+while running intensive parallel applications. 
+Hence, AWS's target is more oriented to offer the service than to provide performance.
+Also with the aim of plainly offering the service of GPGPU, AWS lacks in flexibility. 
+AWS forces the user to choose between a poor instance with a GPU and a powerful instance with 4 GPUs.
+
+Coming back to the Table~\ref{table:awsInstances} can be seen how not only the resources of the ``g2.8xlarge'' has been quadrupled, but also the cost per hour.
+So, in the case of having other necessities (talking in terms of the instances type), with remotes GPUs we could attach an accelerator to any kind of instance.
+Furthermore, reducing the budget in VM could be possible by customizing the resources of the available instances.
+If needed, we could work on an instance of 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs.
+Avoiding to pay the double for features that we are not willing to use in ``g2.8xlarge''.
+
+The last conclusion we drew was related to GPU-shareability. 
+As we deduced, AWS has GPU-capable physical machines reserved which could be idle until a GPU-instance request were made.
+To have such an expensive accelerators in a standby state is counter-productive. So, it would make more sense if we had less devices, accessible from any machine, though.
 

### 2015-10-08 paper.tex
nombre test modificado

--- 2015-10-07 paper.tex
@@
 
-The first of them was {\tt MonteCarloMultigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 

### 2015-10-08 paper.tex
aws discussion rev
nombre instancia modificado

@@
 (C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable instances. Each kind of instance has its own pourpose and its 
+GPU capable instances (G2). Each kind of instance has its own pourpose and its 
 own price. Moreover, each instance has different number of CPUs and network 

### 2015-10-08 paper.tex
OpenStack

--- 2015-10-08 paper.tex
@@
 
-The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The first of them was {\tt MonteCarloMultigpu} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
@@
 From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
-The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
+The Figure shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
 Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
-Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs.
+Thus, if we compare the results of a LAMMPS execution with a larger problem (see Figure), although CUDA continues performing better, the interesting point is the execution time on remote GPUs. 
 They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
@@
 In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
-
-Version de open stack con la que se trabaja y tipo de instalacion ya que está toda en un nodo...
 
@@
 \subsection{Experimental Results}
-The tests were executed on a 5-node cluster where a machine
+The test were executed on a 5-node cluster where a machine
 equipped with a Intel Xeon E7420 quadcore processor at
@@
 In the software stack the testbed system operates a
-Centos 6.6; the cloud platform is deployed by OpenStack; and
-the GPUs use CUDA 6.5. The GPU virtualization support is
-based on rCUDA v5.0.
+Centos 6.6; and the GPUs use CUDA 6.5. The GPU virtualization support is
+based on rCUDA v5.0. In order to deploy an infrastructure as a service cloud,
+we installed OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisor in one
+of these 5-node cluster. A fully-featured OpenStack deployment usually requires
+at least three nodes: one controller node that manages the compute nodes 
+where the VMs are hosted, one network node that manages the logic
+virtual network for the VMs, and one or more compute nodes that 
+run the hypervisor and virtual machines. However, our OpenStack deployment
+was deployed in one-node architecture making better use of our resources.
+Thus, the main OpenStack services such as identity, image, compute, dashboard, and 
+network services were installed in the first machine without GPU described above.
 
@@
 different configurations in order to test the set-ups. MCUDA-
-MEME is an MPI application which is accelerated by GPUs
-Besides, each of its processes must have a GPU to not abort 
-So, the number of GPUs will determine the number of processes that we can launch.
-
+MEME is GPGPU MPI application, so the number of GPUs
+will determine the number of executable processes.
 The Figure~\ref{fig3} shows and compares the execution time of
-the application with different configurations over different setups. 
-We have used up to 3 nodes to spread the processes and launched up to 8 processes (only 4 in exclusive mode), one for each remote GPU.
-Firstly, we can state that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
-Secondly, the lowest execution time is obtained by both modes (exclusive and shared) when running their maximum number of processes with more than 1 node.
-It could mean that it is not worthy to continue giving resources to the application to scale, because the performance growth rate is barely increasing.
-Although, the time is lower sharing the GPUs the setup cannot take advantage of doubling the devices.
+the application with different configurations over different set-
+ups. We can determine that with 1 node, the performance is
+lower, with 2 and 3 is quite similar, though. Hence, the greatest
+difference can be appreciated between modes, where the
+highest performance is achieved by sharing the accelerators.
 
@@
 -Volver a hacer un resumen
+-posibles futuras aplicaciones (lab)
+-pruebas con otros proveedores o con otras plataformas (azure)
 
@@
 So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
-The same thought, would guide us to try other Cloud vendors which more support for HPC.
 
 Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
-would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPUs.
+would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPU.
 
-Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
-With more scheduling policies the flexibility given by the GPGPU module for Openstack would increase.
+Finally, to bring in more strategies to decide where a remote GPUs was created and assigned to a physical device, would be an interesting task to carry out. 
+With more scheduling policies the flexibility given by the GPGPU module for Openstack should increase.
 

### 2015-10-08 paper.tex

--- 2015-10-08 paper.tex
@@
 Also with the aim of plainly offering the service of GPGPU, AWS lacks in flexibility. 
-AWS forces the user to choose between a poor instance with a GPU and a powerful instance with 4 GPUs.
+AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
 

### 2015-10-08 paper.tex
nombre instancia modificado

--- 2015-10-08 paper.tex
@@
 Therefore, we can assure that the nomenclature ``High'' used in the 
-network column correspond to a 1Gb interconnection and the 10Gb has 
-a real bandwidth of 7.5Gb/s.
+network column correspond to a 1~Gb interconnection and the 10~Gb has 
+a real bandwidth of 7.5~Gb/s.
 Moreover, it looks like the instances equipped with a ``High'' interconnection 
-are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
+are software-limited up to 1~Gb/s due to the theoretical and real bandwith 
 match perfectly. The gap between real and theorical can be observed with 
-the 10Gb interconnection which reaches up to 7.5 Gb/s.
+the 10~Gb interconnection which reaches up to 7.5~Gb/s.
 

### 2015-10-08 paper.tex
comentado el discussion
varios

@@
 
-\keywords{Cloud Computing, GPU Virtualization, Resource Management}
+\keywords{Cloud Computing, GPU Virtualization, AWS, Resource Management}
 
@@
 application has become more popular during last years. These devices increment 
-the computational power of a node thanks to its parallel architecture. %The popularity of the GPGPU computation 
+the computational power of a node thanks to its parallel architecture.
 This trend has led several cloud vendors such as Amazon Web Services or some 
@@
 \label{sec:rcuda}
-
 {rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
@@
 (C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable instances (G2). Each kind of instance has its own pourpose and its 
+GPU capable instances. Each kind of instance has its own pourpose and its 
 own price. Moreover, each instance has different number of CPUs and network 
@@
 Therefore, we can assure that the nomenclature ``High'' used in the 
-network column correspond to a 1~Gb interconnection and the 10~Gb has 
-a real bandwidth of 7.5~Gb/s.
+network column correspond to a 1Gb interconnection and the 10Gb has 
+a real bandwidth of 7.5Gb/s.
 Moreover, it looks like the instances equipped with a ``High'' interconnection 
-are software-limited up to 1~Gb/s due to the theoretical and real bandwith 
+are software-limited up to 1 Gb/s due to the theoretical and real bandwith 
 match perfectly. The gap between real and theorical can be observed with 
-the 10~Gb interconnection which reaches up to 7.5~Gb/s.
+the 10Gb interconnection which reaches up to 7.5 Gb/s.
 
@@
 In this section, the different client-server configurations 
-are explained. All the scenarios are based in the maximum number of instance 
-that a user can freely select without submiting a formal request.
+are explained. All the scenarios are based in the maximum number of instances 
+that a user can freely select without submitting a formal request. 
+In particular, can be launched 5 of ``g2.2xlarge'' and 2 of ``g2.8xlarge''.
 
@@
 
-The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
+The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, an application purely computational. 
+It was launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
@@
 MHz, was guest host for the VMs. While the other 4 nodes
-were auxiliary servers with a Tesla C1090 GPU each one. 
+were auxiliary servers with a Tesla C1060 GPU each one. 
 In the software stack the testbed system operates a

### 2015-10-08 paper.tex

@@
 
-\section{\uppercase{Discussion}}
-\label{sec:discussion}
--Volver a hacer un resumen
+%\section{\uppercase{Discussion}}
+%\label{sec:discussion}
+%-Volver a hacer un resumen
 

### 2015-10-08 paper.tex

--- 2015-10-08 paper.tex
@@
 
-Version de open stack con la que se trabaja y tipo de instalacion ya que está toda en un nodo...
-
 \subsection{Managing remote GPUs with OpenStack}

### 2015-10-08 paper.tex
testbed

--- 2015-10-08 paper.tex
@@
 \subsection{Experimental Results}
-The tests were executed on a 5-node cluster where a machine
-equipped with a Intel Xeon E7420 quadcore processor at
-2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667
-MHz, was guest host for the VMs. While the other 4 nodes
-were auxiliary servers with a Tesla C1060 GPU each one. 
-In the software stack the testbed system operates a
-Centos 6.6; the cloud platform is deployed by OpenStack; and
-the GPUs use CUDA 6.5. The GPU virtualization support is
-based on rCUDA v5.0.
+The tests were executed on 2 sets of nodes interconnected through a 1Gb Ethernet network.
+
+The first set, in charge of providing the cloud environment,
+was composed of 3 nodes, each one equipped with a Intel Xeon E7420 quadcore processor at
+2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
+To deploy an infrastructure as a service cloud, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
+A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
+where the VMs are hosted, one network node that manages the logic
+virtual network for the VMs, and one or more compute nodes that 
+run the hypervisor and virtual machines.
+
+The second set, composed by 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
+Their system operates a Centos 6.6; and the GPUs use CUDA 6.5. 
+The GPU virtualization support is based on rCUDA v5.0.
 

### 2015-10-08 paper.tex
references
aws details

@@
 specific target. Inside the large list of instances offered by AWS, 
-we can find: general pourpose (T2, M4 and M3); computer science optimized 
+we can find: general purpose (T2, M4 and M3); computer science optimized 
 (C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable instances (G2). Each kind of instance has its own pourpose and its 
+GPU capable instances (G2). Each kind of instance has its own purpose and its 
 own price. Moreover, each instance has different number of CPUs and network 
@@
 
-For this study, the instances presented in the Table~\ref{table:awsInstances} 
-have been selected. The C3 family instances, which are not equipped with 
-GPUs are going to be used as client whereas the instances of the G2 family are 
-acting as servers.
+For this study, carried out in June 2015, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
+So, at this moment the instances used had the features presented in the Table~\ref{table:awsInstances}.
 
@@
 \end{tabular}\end{center}\end{table}
+
+The selection of these concrete types have the aim of using C3 family instances, which are not equipped with GPUs, as client whereas the instances of the G2 family are acting as servers.
 
@@
 \subsubsection{GPUs}
-As it has been said before, an instance is based in a VM 
-which is running on a real node with its own virtualized 
-components, so it is possible that AWS could use a virtualization 
-framework to offer GPU services to all the instances.
-Therefore, we need to verify if the GPUs showed in the instances are 
-non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
-in this test is higher than the network bandwith, it reveals that it is a 
-real GPU. 
+As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
+Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}.
+As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
 \begin{table}[htb]

### 2015-10-08 paper.tex
Related work

--- 2015-10-08 paper.tex
@@
 \label{sec:state}
-Our solutions
-is based on GPU virtualization and sharing resources in order
+Our solutions is based on GPU virtualization and sharing resources in order
 to reach a fair balance between supply and demand.
@@
 to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented
-by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+to share idle GPUs to other machines, several solutions have been presented.
+The solution presented by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
 them accessible for any VM in the cluster. However, this kind
 of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
-integrated in a Cloud Computing Manager, though its main
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
+This one continues without being integrated in a Cloud Computing Manager, though its main
 drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. On the
-other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
+and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
+to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
+approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
+On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
 it is only focussed on compute intensive HPC applications. 
@@
 
-The IPERF tool has been executed between the instances showed in 
+The IPERF\footnote{\url{http://iperf.fr/}} tool has been executed between the instances showed in 
 the Table~\ref{table:iperf} in order to know the real bandwidth.

### 2015-10-08 paper.tex
ispell pasado

--- 2015-10-08 paper.tex
@@
 the cloud facilities need to be provided with the {rCUDA} software and 
-in the case of OpenStack, a module for manageing virtualized GPUs 
+in the case of OpenStack, a module for managing virtualized GPUs 
  has to be implemented.
@@
 
-For this study, carried out in June 2015, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
-So, at this moment the instances used had the features presented in the Table~\ref{table:awsInstances}.
+For this study, the instances presented in the Table~\ref{table:awsInstances} 
+have been selected. The C3 family instances, which are not equipped with 
+GPUs are going to be used as client whereas the instances of the G2 family are 
+acting as servers.
 
@@
 \end{tabular}\end{center}\end{table}
-
-The selection of these concrete types have the aim of using C3 family instances, which are not equipped with GPUs, as client whereas the instances of the G2 family are acting as servers.
 
@@
 Moreover, it looks like the instances equipped with a ``High'' interconnection 
-are software-limited up to 1~Gb/s due to the theoretical and real bandwith 
-match perfectly. The gap between real and theorical can be observed with 
+are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
+match perfectly. The gap between real and theoretical can be observed with 
 the 10~Gb interconnection which reaches up to 7.5~Gb/s.
@@
 \subsubsection{GPUs}
-As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
-Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}.
-As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
+As it has been said before, an instance is based in a VM 
+which is running on a real node with its own virtualized 
+components, so it is possible that AWS could use a virtualization 
+framework to offer GPU services to all the instances.
+Therefore, we need to verify if the GPUs showed in the instances are 
+non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
+{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
+in this test is higher than the network bandwidth, it reveals that it is a 
+real GPU. 
 \begin{table}[htb]
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using local GPU.}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
 \label{table:bwt}
@@
 \begin{center}\begin{tabular}{cccc}
-Name &  Data Movement & Network & BandWidth \\ \hline \hline
+Name &  Data Movement & Network & Bandwidth \\ \hline \hline
 g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
@@
 with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
-a 10Gb network and the cliend is able to execute application using up to 8 GPUs.
+a 10Gb network and the client is able to execute application using up to 8 GPUs.
 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using {rCUDA}.}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using {rCUDA}.}
 \label{table:bwtrcuda}
@@
 \begin{center}\begin{tabular}{cccc}
-Scenario &  Data Movement & Network & BandWidth \\ \hline \hline
+Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
 A & Host to Device & High& 127 MB/s \\ \hline
@@
 \subsection{Discussion}
-Putting all together,  it is cristal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
+Putting all together,  it is crystal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
 while running intensive parallel applications. 
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea is to move from the original Openstack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
+The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
 This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
@@
    }
-  \caption{Openstack Architecture}
+  \caption{OpenStack Architecture}
   \label{fig:os}
@@
 Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
-With more scheduling policies the flexibility given by the GPGPU module for Openstack would increase.
+With more scheduling policies the flexibility given by the GPGPU module for OpenStack would increase.
 

### 2015-10-08 paper.tex

@@
 \subsubsection{GPUs}
-As it has been said before, an instance is based in a VM 
-which is running on a real node with its own virtualized 
-components, so it is possible that AWS could use a virtualization 
-framework to offer GPU services to all the instances.
-Therefore, we need to verify if the GPUs showed in the instances are 
-non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
-{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
-in this test is higher than the network bandwidth, it reveals that it is a 
-real GPU. 
+As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
+Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}.
+As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
 \begin{table}[htb]

### 2015-10-08 paper.tex
energy

--- 2015-10-08 paper.tex
@@
 \label{sec:state}
-Our solutions is based on GPU virtualization and sharing resources in order
+Our solutions
+is based on GPU virtualization and sharing resources in order
 to reach a fair balance between supply and demand.
@@
 to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines, several solutions have been presented.
-The solution presented by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+to share idle GPUs to other machines. The solution presented
+by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
 them accessible for any VM in the cluster. However, this kind
 of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
-This one continues without being integrated in a Cloud Computing Manager, though its main
+does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. This one continues without being
+integrated in a Cloud Computing Manager, though its main
 drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
-and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
-to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
-approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
-On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
+in order to be run in their virtual-GPU environment. On the
+other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
 it is only focussed on compute intensive HPC applications.
@@
 For this study, carried out in June 2015, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
-So, at this moment the instances used had the features presented in the Table~\ref{table:awsInstances}.
+So, at that moment the instances used had the features presented in the Table~\ref{table:awsInstances}.
 
@@
 
-The IPERF\footnote{\url{http://iperf.fr/}} tool has been executed between the instances showed in 
+The IPERF tool has been executed between the instances showed in 
 the Table~\ref{table:iperf} in order to know the real bandwidth.
@@
 In particular, can be launched 5 of ``g2.2xlarge'' and 2 of ``g2.8xlarge''.
+
+All the instances deployed operate the RHEL 7.1 64 bit system and the version 6.5 of CUDA.
 
@@
 
+Other interesting approach would be to make the system energy-aware. 
+Provided that a GPU could throw power information, a similar study based on the energy consumption could be done.
+
 Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 

### 2015-10-08 paper.tex
state of art modificado

@@
 to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines, several solutions have been presented.
+to share idle GPUs to other machines, several approaches have been implemented.
 The solution presented by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make

### 2015-10-08 paper.tex
mod
revisado por adrian

--- 2015-10-08 paper.tex
@@
 Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
-With this software configuration, these GPUs can be accessed from any VM getting rid 
+With this software configuration, the GPUs can be accessed from any VM getting rid 
 of the necessity of having physical GPUs in the guest host. 
@@
 services and applications on an on-demand basis. 
-These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
+These services replace IT infrastructure for most companies and provide agility and instant elasticity matching 
 perfectly with their software requirements.
@@
 by the hypervisor Xen to access the GPUs in the physical
-node. With the implied limitations of restricting the nodes not
+node, with the implied limitations of restricting the nodes not
 to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines, several solutions have been presented.
-The solution presented by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
+to share idle GPUs to other machines.
+The solution presented by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
 them accessible for any VM in the cluster. However, this kind
@@
 \subsection{Managing remote GPUs with OpenStack}
-The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
+The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig})
+ to a solution where a new shared service takes part into the architecture and it 
+is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
 This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
@@
 Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
-This behaviour is refereed as ``scope'' and determines that a group of instances is logically connected to a pool of GPUs.
+This behaviour is refereed as ``scope'' and it determines that a group of instances is logically connected to a pool of GPUs.
 Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
-Again, the GPU pool can be made of ``exclusive'' GPUs or ``shared''.
+Again, the GPU pool can be composed by ``exclusive'' or ``shared'' GPUs.
 
@@
 Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
-First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
-Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
+First of all, the Instance Launch Panel should be provided with a new field, 
+where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
+As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration appears(see Figure~\ref{fig:ui-launch}).
+Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
 
@@
 The first set, in charge of providing the cloud environment,
-was composed of 3 nodes, each one equipped with a Intel Xeon E7420 quadcore processor at
+was composed of 3 nodes, each one equipped with an Intel Xeon E7420 quadcore processor at
 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
-To deploy an infrastructure as a service cloud, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
+To deploy an infrastructure as a cloud service, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
 A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
@@
 We have designed 6 different set-ups which can be divided
-in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
+into 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
 “Exclusive” mode will provide, at most, 4 accelerators. Whilst
@@
 different configurations in order to test the set-ups. MCUDA-
-MEME is an MPI application which is accelerated by GPUs
+MEME is an MPI application which is accelerated by GPUs.
 Besides, each of its processes must have a GPU to not abort 
-So, the number of GPUs will determine the number of processes that we can launch.
+so, the number of GPUs will determine the number of processes that we can launch.
 

### 2015-10-08 paper.tex
rev sergio

@@
 
-\abstract{The use of accelerators in order to reduce the execution time of some 
-application has become more popular during last years. These devices increment 
-the computational power of a node thanks to its parallel architecture.
-This trend has led several cloud vendors such as Amazon Web Services or some 
-middlewares such as OpenStack to add to their current facilities, or to allow, 
+\abstract{The use of accelerators, such as GPUs, to reduce the execution time of GPGPU applications has become more popular during last years. 
+These devices increment the computational power of a node thanks to its parallel architecture.
+This trend has led several cloud vendors such as Amazon Web Services or some middlewares such as OpenStack to add to their current facilities, or to allow, 
 instances of virtual machines equipped with NVIDIA GPUs. 
-Accomplishing this requirements needs that the guest hosts must be equipped with 
-GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
-The solution presented in this work is based on GPU virtualization and shareability 
-in order to reach an equilibrium between supply and demand of accelerators from the users. 
+Accomplishing these needs, the guest hosts must be equipped with GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between supply and demand of accelerators. 
 Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
@@
 
+****** FALTA UNIR AMBAS PARTES DEL TRABAJO ******
+
 \section{\uppercase{Working with AWS}}
@@
 
-\subsubsection{Network behaviour}
-As it is shown in the Table~\ref{table:awsInstances}, each instance has different 
-interconnection and taking into account that the use of the network 
+\subsubsection{Networking}
+As it is shown in the Table~\ref{table:awsInstances}, each instance has a different 
+network and taking into account that the use of the bandwidth 
 is critical when GPU virtualization is applied, a simple test to verify the 
@@
 Therefore, we can assure that the nomenclature ``High'' used in the 
-network column correspond to a 1~Gb interconnection and the 10~Gb has 
+network column correspond to a 1~Gb interconnection network and the 10~Gb has 
 a real bandwidth of 7.5~Gb/s.
-Moreover, it looks like the instances equipped with a ``High'' interconnection 
+Moreover, it looks like the instances equipped with a ``High'' interconnection network
 are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
@@
 \subsubsection{GPUs}
-As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
+As it has been said before, an instance is based on a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
 Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}.
-As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
+As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
 \begin{table}[htb]
@@
 In this section, the different client-server configurations 
-are explained. All the scenarios are based in the maximum number of instances 
+are explained. All the scenarios are based on the maximum number of instances 
 that a user can freely select without submitting a formal request. 
-In particular, can be launched 5 of ``g2.2xlarge'' and 2 of ``g2.8xlarge''.
+In particular, the maximum number of ``g2.2xlarge'' is 5 and 2 for the ``g2.8xlarge'' instances.
 
@@
 populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
-using the ``High'' interconnection.
+using the ``High'' network.
 
@@
 the {rCUDA} software needs to be installed in order to add 
-the required flexibility to the system. So, the {rCUDA} server has been 
-executed in the server nodes and the {rCUDA} libraries have been installed 
+the required flexibility to the system. So, the {rCUDA} server is 
+executed in the server nodes and the {rCUDA} libraries are installed 
 in the node that is acting as a client.
@@
 The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
-We have removed the results of the Scenario~\ref{subfig:aws2} for clarity ought to they are exactly the same that the obtained from Scenario~\ref{subfig:aws3} up to 8 GPUs. 
-In other words, we could state that the results until the 8th GPU would correspond to the Scenario~\ref{subfig:aws2} and thereafter to the Scenario~\ref{subfig:aws3}.
+We have removed the results of the Scenario B for clarity ought to they are exactly the same that the obtained from Scenario C up to 8 GPUs. 
+In other words, we could state that the results until the 8th GPU would correspond to the Scenario B and thereafter to the Scenario C.
 Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
-However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario~\ref{subfig:aws1} could increase the throughput because
+However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario A could increase the throughput because
 unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
@@
 The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
-Furthermore, for the execution on remote GPUs, can be seen a little difference between both networks, the ``High'' network performs worse than ``10 Gb'' network.
+Furthermore, for the execution on remote GPUs, a little difference between both networks can be appreciated, so, the ``High'' network performs worse than ``10 Gb'' network.
 Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs.
@@
 
-The last conclusion we drew was related to GPU-shareability. 
+The last conclusion we will draw is related to GPU-shareability. 
 As we deduced, AWS has GPU-capable physical machines reserved which could be idle until a GPU-instance request were made.
@@
 
-%\section{\uppercase{Discussion}}
-%\label{sec:discussion}
+\subsection{Discussion}
+\label{sec:discussion}
 %-Volver a hacer un resumen

### 2015-10-08 paper.tex
estructurado el state of art
Motivation improved

@@
 Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project. On
+made, but none of them are so ambitious as our project.
+
+On
 the one hand the work done in~\cite{younge2013enabling} allows the VM managed
@@
 to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines.
-The solution presented by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
+to share idle GPUs to other machines. The solution presented 
+by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
 them accessible for any VM in the cluster. However, this kind
@@
 it is only focussed on compute intensive HPC applications.
+
 The main idea of our proposal goes further and apart from
@@
 process and the low required IT invest. The features provided 
-by the cloud vendors are useful for companies and individuals 
+by the cloud services are useful for companies and individuals 
 which have access to the last technology without spending a 
@@
 
-The power of current cloud vendors is based on the use 
+Two different kinds of cloud approaches are available:
+on the one hand, public facilities such as Amazon Web Services or 
+Microsoft Azure allows users to chose a by default system configuration 
+or to compose its own system features. On the other hand, institutions/companies 
+are able to deploy their own cloud infrastructure by using software products 
+like OpenStack.
+
+In both cases, the power of current cloud facilities is based on the use 
 of Virtual Machines (VM) which  
@@
 once it has been assigned to an instance, no other VM can use
-it. Therefore, the number of VM that are allowed to use a GPU is 
+it. Therefore, the number of VM that are allowed to use GPUs is 
 limited by the number of the installed devices. This lack of 
@@
 achieving a higher utilization rate and reducing the idle status time.
-
 Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
 environment should be able to use and share all the GPUs installed in 
-the physical nodes removing the current limitation. But, in order to achieve this goal, 
-the cloud facilities need to be provided with the {rCUDA} software and 
+the physical nodes removing the current limitations.
+
+The main goals of this work are to study current cloud solutions from 
+the point of view of HPC field and GPU usage, and to analyze and improve 
+them by adding flexibility with GPU virtualization.
+
+Our work have been focused on Amazon Web Service as example of public 
+solution and on OpenStack software acting as a private cloud environment. 
+In order to achieve it, these cloud facilities need to be provided with the {rCUDA} software and 
 in the case of OpenStack, a module for managing virtualized GPUs 
- has to be implemented.
-
-****** FALTA UNIR AMBAS PARTES DEL TRABAJO ******
+ requires to be implemented.
 

### 2015-10-08 paper.tex
discussion os

--- 2015-10-08 paper.tex
@@
 As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
+% \begin{table}[htb]
+% \renewcommand{\arraystretch}{1.3}
+% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
+% \label{table:bwt}
+% \tabcolsep=0.09cm
+% \begin{center}\begin{tabular}{cccc}
+% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
+% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
+% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
+% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
+% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
+% \end{tabular}\end{center}\end{table}
+
 \begin{table}[htb]
@@
 \tabcolsep=0.09cm
-\begin{center}\begin{tabular}{cccc}
-Name &  Data Movement & Network & Bandwidth \\ \hline \hline
-g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
-g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
-g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
-g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
+\begin{center}\begin{tabular}{ccc}
+Name &  Data Movement & Bandwidth \\ \hline \hline
+g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
+g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
+g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
+g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
 \end{tabular}\end{center}\end{table}
@@
  to a solution where a new shared service takes part into the architecture and it 
-is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
+is responsible for all about GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
 This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
 As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
-which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
-The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
+which splits the HTTP query in order to make use of both, the GPGPU API for GPU-related operations, and the nova API for the rest of the computation. 
+The new GPGPU Service has brought a new feature which brings more flexibility, apart from the regular idea of having GPUs in a VM, from now on, we can create ``GPU-pools''. 
 A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
@@
   \centering
-  \subfigure[Original in version XXXXXXXXXXXX]{
+  \subfigure[Version Icehouse]{
     \includegraphics[width=\linewidth]{images/os-orig.png}
@@
 The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
-The users are given two configuration options to decide whether a physical GPU will be completely allocated for an instance (first column) or the instance will address a partition of the GPU as if it were a device (second column).
+The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
 We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
@@
 where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration appears(see Figure~\ref{fig:ui-launch}).
+As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
 Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
@@
 \subsection{Discussion}
-\label{sec:discussion}
-%-Volver a hacer un resumen
+Again, the networking issue is restricting the performance of our executions, but it has been discussed during virtually all the paper.
+However, it is not a new topic and in~\cite{tonithesis} is explained that investing in the network infrastructure could make the difference.
 
@@
 \label{sec:conclusions}
+We have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
 

### 2015-10-08 paper.tex
Motivation improved

--- 2015-10-08 paper.tex
@@
 on the one hand, public facilities such as Amazon Web Services or 
-Microsoft Azure allows users to chose a by default system configuration 
-or to compose its own system features. On the other hand, institutions/companies 
+Microsoft Azure that allow users to chose a pre-configured system  
+and on the other hand, institutions/companies 
 are able to deploy their own cloud infrastructure by using software products 

### 2015-10-08 paper.tex
discussion os

--- 2015-10-08 paper.tex
@@
 Again, the networking issue is restricting the performance of our executions, but it has been discussed during virtually all the paper.
-However, it is not a new topic and in~\cite{tonithesis} is explained that investing in the network infrastructure could make the difference.
+However, it is not a new topic and in~\cite{tonithesis} is explained that improving the network infrastructure could make the difference.
+
+The most remarkable achievement is the wide range of possible configuration and the freedom that would have an administrator to set and reset a system to fit the 
+user requirements.
+
+With an small investment in infrastructure and in maintenance can be fulfilled the requests of GPU-capable hosts in a completely transparent way for the user.
+Energy can be saved, not only Thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
 

### 2015-10-08 paper.tex
discussion os

@@
 
-With an small investment in infrastructure and in maintenance can be fulfilled the requests of GPU-capable hosts in a completely transparent way for the user.
-Energy can be saved, not only Thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
+The requests of GPU-capable hosts in a completely transparent way for the user can be fulfilled with an small investment in infrastructure and in maintenance.
+Energy can be saved, not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
 

### 2015-10-08 paper.tex
discussion openstack touched

@@
 The requests of GPU-capable hosts in a completely transparent way for the user can be fulfilled with an small investment in infrastructure and in maintenance.
-Energy can be saved, not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
+Moreover, energy can be saved, not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
 

### 2015-10-20 paper.tex
conclusions

--- 2015-10-08 paper.tex
@@
 \section{\uppercase{Working with AWS}}
-\label{sec:aws}
+\label{sec:workingaws}
 In this section, first of all, the current user-level features 
@@
 \label{sec:conclusions}
-We have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
+We have seen a thoroughly study of the possibilities offered by AWS in everything to do with GPUs.
+Its limitations drove us to deploy our own private cloud to have more flexibility when working with these accelerators.
+So that, we have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
+
+Leaving aside the performance of the executions which we knew in advance that would be poor due to the interconnection network features, we have created new operation modes.
+They have brought interesting new ways of using GPUs in situations where the simple fact of having accessing a GPU is more important than having a powerful GPU to boost the performance.
 
@@
 \section*{\uppercase{Acknowledgements}}
-The authors would like to thank the IT members of the department Vicente Roca and Gustavo Edo for their help.
+The authors would like to thank the IT members of the department Gustavo Edo and Vicente Roca for their help.
 

### 2015-10-21 paper.tex
primera version introduccion

@@
 \label{sec:introduction}
+The growth in the use of coprocessor and accelerators has been continuous
+in the past several years. In the most recent TOP500
+ranking~\cite{nov14top500},  dated June 2015, 90 supercomputers
+were equipped with accelerators of some sort.  These devices offer
+performance improvements over traditional processors for
+computational kernels which match their architecture; moreover, 
+they are more energy efficient achieving an important MFlop/Watt ratio.
+
+The generalized adoption of coprocessors to accelerate general-purpose 
+codes started with the release of the CUDA programming model for
+NVIDIA GPUs~\cite{cuda65}. This libraries include both high- and low-level application programming interfaces (APIs)
+to allow the use of NVIDIA devices as general-purpose
+accelerators.  
+%Prior to that, researchers matched 
+%GPU architectures to computational kernels, but these had to
+%use graphics-oriented APIs such as OpenGL~\cite{opengl} or
+%Cg~\cite{cg} to do so. 
+Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
+to offer a cross-vendor solution. The third-generation programming
+model for accelerators, based on compiler directives, started with
+OpenACC~\cite{openaccweb}. This standard defines a collection of 
+directives to facilitate platform-independent accelerator
+usage. This generation is also completed with other task-oriented, directive-based programming model such as OmpSs~\cite{ompssweb} 
+which can target directly CPUs or accelerators and recently, OpenMP~\cite{openmp4} incorporated some
+accelerator-oriented directives in its version 4.
+%, although 
+%no available compiler supports them to date.
+
+The current approach for creating clusters with GPUs has been to
+populate every compute node with one or more of these devices. 
+However, these configurations present a low
+utilization rate of the computational resources available in the
+hardware accelerators, not only because of mismatches between the application's type of parallelism
+and the coprocessor architecture, but also because of applications can be hardly parallelized more than 
+80\% of the total code. 
+In order to address this problem, remote virtualization was proposed.
+Among the virtualization frameworks, the most
+prominent is \rcuda~\cite{tonithesis,toniparco}, which
+enables cluster configurations with fewer GPUs than nodes.  With this
+approach, GPU-equipped nodes act as GPGPU (general-purpose GPU) servers for the rest of the
+compute nodes of the cluster, enabling a CUDA-sharing solution that potentially achieves 
+ a higher overall GPU load in the system. Compared with 
+other CUDA and OpenCL virtualization frameworks (e.g.
+DS-CUDA~\cite{dscuda}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}),
+ \rcuda\ is a mature production-ready framework
+ that offers support for the latest CUDA revisions and provides wide
+coverage of current GPGPU APIs.
 

### 2015-10-21 paper.tex
primera version introduccion

--- 2015-10-20 paper.tex
@@
 \label{sec:introduction}
+The use of coprocessor and accelerators has been continuously
+growing in the past several years and 90 supercomputers
+were equipped with accelerators of some sort in the most 
+recent TOP500 list~\cite{nov14top500}, dated June 2015.  
+These devices achieve better performance than traditional processors for
+codes which match their architecture.
+
+These devices were general adopted with the release of the CUDA programming model for
+NVIDIA GPUs~\cite{cuda65}. This libraries include both high- and low-level application programming interfaces (APIs)
+to allow the use of NVIDIA devices as general-purpose
+accelerators.  
+Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
+to offer a cross-vendor solution.
+
+The current approach for creating clusters with GPUs has been to
+install one or more of these devices in each compute node. 
+However, these configurations present a low
+utilization rate of the computational resources  
+ because of mismatches between the application's type of parallelism
+and the accelerators' architecture.
+
+In order to address this problem, remote virtualization was proposed 
+and several solutions have been presented (e.g. {rCUDA}~\cite{tonithesis,toniparco},
+DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).
+
+Among the virtualization frameworks, the most
+prominent is {rCUDA}, which
+enables cluster configurations with fewer GPUs than nodes.  With this
+approach, GPU-equipped nodes act as GPGPU (general-purpose GPU), enabling a CUDA-sharing solution that potentially achieves 
+ a higher overall GPU load in the system.
 

### 2015-10-21 paper.tex
intro cloud

--- 2015-10-21 paper.tex
@@
 
-%\noindent
+Many cloud vendors have started offering VM with GPUs in order to provide
+
+Accelerators such as GPUs have been adopted by major cloud vendors,
+
+as GPUs provide orders-of-magnitude speedup for computation-intensive data-parallel applications.
+
+In cloud environments, sharing GPU resources among multiple virtual machines (VMs) is not straightforward which is translated into a low utilization rate.
+
+accelerate general-purpose workloads in many applica-
+tion domains such as supercomputing systems. Unfor-
+tunately, enterprise and cloud computing domains have
+
+inefficient access to GPU virtualization technology be-
+cause the required resource isolation of multiple vir-
+tual machines (VMs) accessing the same set of GPU re-
+sources cannot be efficiently guaranteed.
+
+Efficiently utilizing GPU resources in the cloud envi-
+ronment is not so straightforward. Current commercial
+
+cloud providers (e.g., Amazon Elastic Compute Cloud—
+
+EC2) dedicate an individual physical instance of GPUs
+
+to a VM, which causes resource under-utilization. A fea-
+sible solution to make GPU a truly shared and schedula-
+ble resource in the cloud is through GPU virtualization,
+
+which allows multiple VMs to access the same set of
+
+GPUs simultaneously. Recent research [8, 9, 14, 20, 26,
+
+27] implements several GPU virtualization technologies
+
+for graphics rendering applications as well as GPGPU
+
+computing workloads.
 

### 2015-10-21 paper.tex
primera version introduccion
merda q plou

--- 2015-10-21 paper.tex
@@
  a higher overall GPU load in the system.
+
+ Many cloud vendors have started offering VM with GPUs in order to provide
+
+Accelerators such as GPUs have been adopted by major cloud vendors,
+
+as GPUs provide orders-of-magnitude speedup for computation-intensive data-parallel applications.
+
+In cloud environments, sharing GPU resources among multiple virtual machines (VMs) is not straightforward which is translated into a low utilization rate.
+
+accelerate general-purpose workloads in many applica-
+tion domains such as supercomputing systems. Unfor-
+tunately, enterprise and cloud computing domains have
+
+inefficient access to GPU virtualization technology be-
+cause the required resource isolation of multiple vir-
+tual machines (VMs) accessing the same set of GPU re-
+sources cannot be efficiently guaranteed.
+
+Efficiently utilizing GPU resources in the cloud envi-
+ronment is not so straightforward. Current commercial
+
+cloud providers (e.g., Amazon Elastic Compute Cloud—
+
+EC2) dedicate an individual physical instance of GPUs
+
+to a VM, which causes resource under-utilization. A fea-
+sible solution to make GPU a truly shared and schedula-
+ble resource in the cloud is through GPU virtualization,
+
+which allows multiple VMs to access the same set of
+
+GPUs simultaneously. Recent research [8, 9, 14, 20, 26,
+
+27] implements several GPU virtualization technologies
 

### 2015-10-21 paper.tex
merda q plou

--- 2015-10-21 paper.tex
@@
 \label{sec:introduction}
-The use of coprocessor and accelerators has been continuously
-growing in the past several years and 90 supercomputers
-were equipped with accelerators of some sort in the most 
-recent TOP500 list~\cite{nov14top500}, dated June 2015.  
-These devices achieve better performance than traditional processors for
-codes which match their architecture.
-
-These devices were general adopted with the release of the CUDA programming model for
-NVIDIA GPUs~\cite{cuda65}. This libraries include both high- and low-level application programming interfaces (APIs)
-to allow the use of NVIDIA devices as general-purpose
-accelerators.  
-Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
-to offer a cross-vendor solution.
-
-The current approach for creating clusters with GPUs has been to
-install one or more of these devices in each compute node. 
-However, these configurations present a low
-utilization rate of the computational resources  
- because of mismatches between the application's type of parallelism
-and the accelerators' architecture.
-
-In order to address this problem, remote virtualization was proposed 
-and several solutions have been presented (e.g. {rCUDA}~\cite{tonithesis,toniparco},
-DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).
-
-Among the virtualization frameworks, the most
-prominent is {rCUDA}, which
-enables cluster configurations with fewer GPUs than nodes.  With this
-approach, GPU-equipped nodes act as GPGPU (general-purpose GPU), enabling a CUDA-sharing solution that potentially achieves 
- a higher overall GPU load in the system.
-
-%\noindent
 

### 2015-10-21 paper.tex

--- 2015-10-21 paper.tex
@@
 a higher overall GPU load in the system.
-
- Many cloud vendors have started offering VM with GPUs in order to provide
-
-Accelerators such as GPUs have been adopted by major cloud vendors,
-
-as GPUs provide orders-of-magnitude speedup for computation-intensive data-parallel applications.
-
-In cloud environments, sharing GPU resources among multiple virtual machines (VMs) is not straightforward which is translated into a low utilization rate.
-
-accelerate general-purpose workloads in many applica-
-tion domains such as supercomputing systems. Unfor-
-tunately, enterprise and cloud computing domains have
-
-inefficient access to GPU virtualization technology be-
-cause the required resource isolation of multiple vir-
-tual machines (VMs) accessing the same set of GPU re-
-sources cannot be efficiently guaranteed.
-
-Efficiently utilizing GPU resources in the cloud envi-
-ronment is not so straightforward. Current commercial
-
-cloud providers (e.g., Amazon Elastic Compute Cloud—
-
-EC2) dedicate an individual physical instance of GPUs
-
-to a VM, which causes resource under-utilization. A fea-
-sible solution to make GPU a truly shared and schedula-
-ble resource in the cloud is through GPU virtualization,
-
-which allows multiple VMs to access the same set of
-
-GPUs simultaneously. Recent research [8, 9, 14, 20, 26,
-
-27] implements several GPU virtualization technologies
-
-%\noindent
 

### 2015-10-21 paper.tex
intro cloud

@@
 a higher overall GPU load in the system.
+
+Nowadays, many cloud vendors have started offering virtual machines (VM) with GPUs in order to provide
+GPGPU computation service to users who need it. An examples of that are:
+Amazon Web Services (AWS), Penguin Computing, Softlayer and Microsoft Azure.
+In the pubic scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
+Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
+Openstack or Opennebula.
+
+However, in cloud environments sharing GPU resources among multiple VMs
+is not straightforward as would be in a physical servers which is translated into a low utilization rate.
+On the one hand, instances in public clouds are not customizable enough. 
+On the other, in private cloud, although the instances can be customized in many aspects, when referring to GPUs
+the number of options is reduced.
+Moreover, neither vendors nor tools are offering GPGPU as a service and that is what we are pursing:
+to offer flexibility and boost the utilization of GPU resources either in private or public clouds.
 

### 2015-10-21 paper.tex
intro sections

@@
 Amazon Web Services (AWS), Penguin Computing, Softlayer and Microsoft Azure.
-In the pubic scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
+In the public scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
 Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
@@
 to offer flexibility and boost the utilization of GPU resources either in private or public clouds.
+
+The main goals of this work are to study current cloud solutions from 
+the point of view of HPC field and GPU usage, and to analyze and improve 
+them by adding flexibility with GPU virtualization.
+
+The rest of the paper is structured as follows: 
+in Section~\ref{sec:background} the technologies used are explained;
+Section~\ref{sec:state} summarizes previous work in this field; 
+Section~\ref{sec:motivation} states the motivation to carry out this work;
+all the effort made using AWS is explained in Section~\ref{sec:workingaws};
+while the effort made in Openstack is described in Section~\ref{sec:openstack};
+finally Section~\ref{sec:conclusions} gives an overview of the advances made and~\ref{sec:future} propose the next steps
+should be take in this research.
 
@@
 the physical nodes removing the current limitations.
-
-The main goals of this work are to study current cloud solutions from 
-the point of view of HPC field and GPU usage, and to analyze and improve 
-them by adding flexibility with GPU virtualization.
 

### 2015-10-21 paper.tex
citas modificadas

@@
 In order to address this problem, remote virtualization was proposed
-and several solutions have been presented (e.g. {rCUDA}~\cite{tonithesis,toniparco},
+and several solutions have been presented (e.g. {rCUDA}~\cite{tonithesis},
 DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).
@@
 \label{sec:rcuda}
-{rCUDA}~\cite{tonithesis,toniparco} is a middleware that enables transparent access
+{rCUDA}~\cite{toniparco} is a middleware that enables transparent access
 to any NVIDIA GPU device present in a cluster from all compute

### 2015-10-21 paper.tex
intro footnotes

--- 2015-10-21 paper.tex
@@
 GPGPU computation service to users who need it. An examples of that are:
-Amazon Web Services (AWS), Penguin Computing, Softlayer and Microsoft Azure.
+Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
+Penguin Computing\footnote{\url{http://www.penguincomputing.com}}, 
+Softlayer\footnote{\url{http://www.softlayer.com}} and 
+Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
 In the public scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
 Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
-Openstack or Opennebula.
+Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
 

### 2015-10-21 paper.tex
intro footnotes

@@
 all the effort made using AWS is explained in Section~\ref{sec:workingaws};
-while the effort made in Openstack is described in Section~\ref{sec:openstack};
+while the effort made in Openstack is described in Section~\ref{sec:gpgpuOS};
 finally Section~\ref{sec:conclusions} gives an overview of the advances made and~\ref{sec:future} propose the next steps
@@
 \section{\uppercase{GPGPU Shared Service in OpenStack}}
+\label{sec:gpgpuOS}
 In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.

### 2015-10-22 paper.tex
introduccion modificada

@@
 \label{sec:introduction}
-The use of coprocessor and accelerators has been continuously
-growing in the past several years and 90 supercomputers
-were equipped with accelerators of some sort in the most
-recent TOP500 list~\cite{nov14top500}, dated June 2015.
-These devices achieve better performance than traditional processors for
-codes which match their architecture.
+Nowadays, many cloud vendors have started offering virtual machines (VM) with GPUs in order to provide
+GPGPU computation service to users who need it. An examples of that are:
+Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
+Penguin Computing\footnote{\url{http://www.penguincomputing.com}}, 
+Softlayer\footnote{\url{http://www.softlayer.com}} and 
+Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
+In the public scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
+Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
+Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
+
+However, in cloud environments sharing GPU resources among multiple VMs
+is not straightforward as would be in a physical servers which is translated into a low utilization rate.
+On the one hand, instances in public clouds are not customizable enough. 
+On the other, in private cloud, although the instances can be customized in many aspects, when referring to GPUs
+the number of options is reduced.
+Moreover, neither vendors nor tools are offering GPGPU as a service and that is what we are pursing:
+to offer flexibility and boost the utilization of GPU resources either in private or public clouds.
 
@@
 a higher overall GPU load in the system.
-
-Nowadays, many cloud vendors have started offering virtual machines (VM) with GPUs in order to provide
-GPGPU computation service to users who need it. An examples of that are:
-Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
-Penguin Computing\footnote{\url{http://www.penguincomputing.com}}, 
-Softlayer\footnote{\url{http://www.softlayer.com}} and 
-Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
-In the public scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
-Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
-Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
-
-However, in cloud environments sharing GPU resources among multiple VMs
-is not straightforward as would be in a physical servers which is translated into a low utilization rate.
-On the one hand, instances in public clouds are not customizable enough. 
-On the other, in private cloud, although the instances can be customized in many aspects, when referring to GPUs
-the number of options is reduced.
-Moreover, neither vendors nor tools are offering GPGPU as a service and that is what we are pursing:
-to offer flexibility and boost the utilization of GPU resources either in private or public clouds.
 

### 2015-10-23 paper.tex
abstract revisado

@@
 
-\title{Enabling the Use of GPU Virtualization in Cloud Environments}
+\title{Enabling GPU Virtualization in Cloud Environments}
 \author{\authorname{Sergio Iserte, Francisco J. Clemente-Castell\'o, Adri\'an Castell\'o,\\Rafael Mayo and Enrique S. Quintana-Ort\'i}
@@
 
-\abstract{The use of accelerators, such as GPUs, to reduce the execution time of GPGPU applications has become more popular during last years. 
+\abstract{The use of accelerators, such as graphics processing units (GPUs), to reduce the execution time of compute-intensive applications has become more popular during the past few years. 
 These devices increment the computational power of a node thanks to its parallel architecture.
-This trend has led several cloud vendors such as Amazon Web Services or some middlewares such as OpenStack to add to their current facilities, or to allow, 
-instances of virtual machines equipped with NVIDIA GPUs. 
-Accomplishing these needs, the guest hosts must be equipped with GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
-The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between supply and demand of accelerators. 
-Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
-With this software configuration, the GPUs can be accessed from any VM getting rid 
-of the necessity of having physical GPUs in the guest host. 
-Moreover, our project studies the viability of these join by using a public cloud 
-service configuration and develops a module for OpenStack middleware in order to 
-add the needed support for the virtualized devices and the logic to manage them. 
-The results show a feasible configuration which adds flexibility to the current and well-known 
+This trend has led cloud vendors, as Amazon Web Services, or middlewares such as OpenStack to add to their current facilities   
+instances of virtual machines including GPUs. 
+To fullfil these needs, the guest hosts must be equipped with GPUs which will be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and applications' demand of accelerators. 
+Hence, we propose to decouple real GPUs from the nodes by using the virtualization technology rCUDA. 
+With this software configuration, GPUs can be accessed from any VM avoiding the need  
+of having a physical GPUs in the guest host. 
+Moreover, we study the viability of this approach by using a public cloud 
+service configuration and develop a module for OpenStack middleware in order to 
+add support for the virtualized devices and the logic to manage them. 
+The results demonstrate this is a viable configuration which adds flexibility to the current and well-known 
 cloud solutions. 

### 2015-10-23 paper.tex
revision 6.1

--- 2015-10-22 paper.tex
@@
 \label{sec:gpgpuOS}
-In this section all the work related to OpenStack that has been done is explained, from the development of an external module to handle remote GPUs to the physical setup and its testing.
+In this section we review the work related with OpenStack, from the development of an external module to handle remote GPUs to the physical setup and its testing.
 
 \subsection{Managing remote GPUs with OpenStack}
-The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig})
- to a solution where a new shared service takes part into the architecture and it 
-is responsible for all about GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
-This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
-As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
-which splits the HTTP query in order to make use of both, the GPGPU API for GPU-related operations, and the nova API for the rest of the computation. 
-The new GPGPU Service has brought a new feature which brings more flexibility, apart from the regular idea of having GPUs in a VM, from now on, we can create ``GPU-pools''. 
-A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.
+The main idea is to evolve from the original OpenStack architecture (see Figure~\ref{subfig:os-orig})
+ to a solution where a new shared service, responsible for all about GPU accelerators, becomes integrated into the architecture (see Figure~\ref{subfig:os-mod}).
+This new service brings more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
+As illustrated in Figure~\ref{fig:internal}, we alter the original OpenStack Dashboard with a new parser, 
+which splits the HTTP query in order to make use of both the GPGPU API for GPU-related operations, and the Nova API for the rest of the computations. 
+The new GPGPU Service allows to access GPUs in a VM, but also to create ``GPU-pools'', a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more VMs.
 
@@
    \quad
-  \subfigure[With GPGPU developed module]{
+  \subfigure[With GPGPU module]{
     \includegraphics[width=\linewidth]{images/os1.jpg}
@@
    }
-  \caption{OpenStack Architecture}
+  \caption{OpenStack Architecture.}
   \label{fig:os}
@@
 
-Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
+Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified and the tool can be easily ported to other Cloud Computing solutions.
 

### 2015-10-23 paper.tex
revision 6.2

@@
 \subsection{Working Modes}
-The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
+The developed module allows users to create any of the scenarios displayed in Figure~\ref{fig2}. 
 The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
-We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
-Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
-A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
-So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
-As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
+We refer to this as the ``mode'', with possible values being: ``exclusive'' or ``shared''. 
+Let us assume there are 4 GPU devices in the cluster (independently where they were hosted. 
+An example of those scenarios is shown in Figure~\ref{fig2}.
+There, in ``exclusive'' mode the instance monopolizes all the GPUs, while in the ``shared'' mode, the GPUs are partitioned. 
+As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
 
@@
 This behaviour is refereed as ``scope'' and it determines that a group of instances is logically connected to a pool of GPUs.
-Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
-Again, the GPU pool can be composed by ``exclusive'' or ``shared'' GPUs.
+Working with ``public scope'' (bottom row of Figure~\ref{fig2}) implies that the GPUs of a pool can be used simultaneously by all the instances linked to it.
+Again, the GPU pool can be composed of ``exclusive'' or ``shared'' GPUs.
 

### 2015-10-23 paper.tex
introduction revisado

--- 2015-10-23 paper.tex
@@
 \label{sec:introduction}
-Nowadays, many cloud vendors have started offering virtual machines (VM) with GPUs in order to provide
-GPGPU computation service to users who need it. An examples of that are:
+Nowadays, many cloud vendors have started offering virtual machines (VMs) with GPUs in order to provide
+GPGPU computation services to users. A few relevant examples include 
 Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
@@
 Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
-In the public scope, one of the most popular cloud vendors is AWS that has a wide range of preconfigured instances ready to be launched.
-Otherwise, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
+In the public scope, one of the most popular cloud vendors is AWS, that offers a wide range of preconfigured instances ready to be launched.
+Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
 Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
 
-However, in cloud environments sharing GPU resources among multiple VMs
-is not straightforward as would be in a physical servers which is translated into a low utilization rate.
-On the one hand, instances in public clouds are not customizable enough. 
-On the other, in private cloud, although the instances can be customized in many aspects, when referring to GPUs
+Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
+is not straightforward as is in a physical servers resulting into a low utilization rate.
+On one hand, instances in public clouds are not easily customizable. 
+On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
 the number of options is reduced.
-Moreover, neither vendors nor tools are offering GPGPU as a service and that is what we are pursing:
-to offer flexibility and boost the utilization of GPU resources either in private or public clouds.
+Moreover, neither vendors nor tools are offer GPGPU wasting the opportunity of providing flexibility and boosty the utilization of GPU resources in both private or public clouds.
 
-These devices were general adopted with the release of the CUDA programming model for
-NVIDIA GPUs~\cite{cuda65}. This libraries include both high- and low-level application programming interfaces (APIs)
+These devices were adopted with the release of the CUDA programming model for
+NVIDIA GPUs~\cite{cuda65}, which includes both high- and low-level application programming interfaces (APIs)
 to allow the use of NVIDIA devices as general-purpose
@@
 
-The current approach for creating clusters with GPUs has been to
-install one or more of these devices in each compute node.
-However, these configurations present a low
+The current approach to assemble a cluster with accelerators consist in installing 
+one or more of these devices in each compute node.
+However, these configurations often present a low
 utilization rate of the computational resources
@@
 
-In order to address this problem, remote virtualization was proposed
-and several solutions have been presented (e.g. {rCUDA}~\cite{tonithesis},
+Remote virtualization has been recently proposed to deal with the low-usage problem 
+(e.g. {rCUDA}~\cite{tonithesis},
 DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).
 
-Among the virtualization frameworks, the most
-prominent is {rCUDA}, which
-enables cluster configurations with fewer GPUs than nodes.  With this
-approach, GPU-equipped nodes act as GPGPU (general-purpose GPU), enabling a CUDA-sharing solution that potentially achieves
+Roughtly speaking these virtualization frameworks 
+enable cluster configurations with fewer GPUs than nodes.  The goal is that 
+GPU-equipped nodes act as GPGPU (general-purpose GPU) servers, yielding a CUDA-sharing solution that potentially achieves
 a higher overall GPU load in the system.
@@
 the point of view of HPC field and GPU usage, and to analyze and improve 
-them by adding flexibility with GPU virtualization.
+them by adding flexibility with GPU virtualization. In order to reach 
+this goal, we select rCUDA, a virtualization tool that is possibly the more complete 
+and up-to-date for NVIDIA GPUs.
 
-The rest of the paper is structured as follows: 
-in Section~\ref{sec:background} the technologies used are explained;
+The rest of the paper is structured as follows. 
+In Section~\ref{sec:background} the technologies we introduce used in this work;
 Section~\ref{sec:state} summarizes previous work in this field; 
 Section~\ref{sec:motivation} states the motivation to carry out this work;
-all the effort made using AWS is explained in Section~\ref{sec:workingaws};
-while the effort made in Openstack is described in Section~\ref{sec:gpgpuOS};
-finally Section~\ref{sec:conclusions} gives an overview of the advances made and~\ref{sec:future} propose the next steps
-should be take in this research.
+the effort to use AWS is explained in Section~\ref{sec:workingaws};
+while the work with Openstack is described in Section~\ref{sec:gpgpuOS};
+finally Section~\ref{sec:conclusions} summarizes the advances and Section~\ref{sec:future} proposes the next steps
+ of this research.
 

### 2015-10-23 paper.tex
revision 6.3

--- 2015-10-23 paper.tex
@@
 Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
-First of all, the Instance Launch Panel should be provided with a new field, 
-where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
-As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
+First of all, the Instance Launch Panel should be extended with a new field, where the user could assign an existing GPU pool, create a new one, or keep the instance without accelerators of this kind.
+When the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
 Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

### 2015-10-23 paper.tex
state-of-art revisado

--- 2015-10-23 paper.tex
@@
 to any NVIDIA GPU device present in a cluster from all compute
-nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
+nodes. The GPUs can be accessed and shared between nodes, and a single node can use all the graphic accelerators
 as if they were local.
-These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
-resource, space, and energy~\cite{energy14} requirements.
+These features focus in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
+resource, space, and energy requirements~\cite{energy14}.
 rCUDA is structured following a client-server distributed
-architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
+architecture: the client middleware runs in the same cluster node where the application demanding GPGPU
 acceleration services is executed, providing a transparent replacement for the
@@
 on top of a virtualization layer.
-With the aim to be updated with new GPU programming models, {rCUDA} also supports
+To deal with new GPU programming models, {rCUDA} has been recently extended to accommodate 
 directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.
@@
 technology, making accelerator-enabled clusters more flexible and
-energy efficient.
+energy-efficient.
 
@@
 \label{sec:aws}
-AWS~\cite{aws} is a public cloud computing provider
-and it is composed by several cloud services such as 
- cloud-based computation, storage and other functionality 
+AWS~\cite{aws} is a public cloud computing provider, 
+composed of several services, such as 
+ cloud-based computation, storage and other functionality, 
 that enables organizations and/or individuals to deploy
 services and applications on an on-demand basis. 
-These services replace IT infrastructure for most companies and provide agility and instant elasticity matching 
+These services replace local IT infrastructure for companies and provide agility and instant elasticity matching 
 perfectly with their software requirements.
 
-From the point of view of HPC, AWS offers high performance facilities including 
-instances equipped with GPUs and high performance network interconnection which 
-are common resources used in this research field.
+From the point of view of HPC, AWS offers high performance facilities via 
+instances equipped with GPUs and high performance network interconnection, which 
+are common resources leveraged in this research field.
 
@@
 control while empowering their users to provision resources through a web interface
-or a command-line. 
-It supports most available hypervisors and handles provisioning 
+or a command-line interface.  
+OpenStack supports most available hypervisors and handles provisioning 
 and life-cycle management of VMs.
 
-The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
+The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
 or software requirements and the ability to integrate with legacy systems and third party technologies.
 
-From the HPC perspective, OpenStack offers high performance computing (HPC) configurations using
-multiple supported hypervisors and different hardware architectures. However, despite to be able to
-specify some extra parameters for special resources like GPUs, choosing a host based on the existence
-of a GPU is currently unsupported in OpenStack~\cite{OpenStackGPU}.
+From the HPC perspective, OpenStack offers high performance configurations using
+multiple supported hypervisors and different hardware architectures. Despite in OpenStack is possible to 
+specify some extra parameters requesting special resources, like GPUs, choosing a host based on the existence
+of a GPU is currently unsupported~\cite{OpenStackGPU}.
 
-\section{\uppercase{State of Art}}
+\section{\uppercase{State-of-the-Art}}
 \label{sec:state}
-Our solutions is based on GPU virtualization and sharing resources in order
-to reach a fair balance between supply and demand. 
-Several attempts of achieving what we are pursing have been
-made, but none of them are so ambitious as our project.
+Our solutions to the defficiency exposed in the previous section is based on GPU virtualization, sharing resources in order
+to attain a fair balance between supply and demand. 
+While several efforst with the same goal have been made in the past, as exposed next, 
+none of them is so ambitious as ours.
 
-On
-the one hand the work done in~\cite{younge2013enabling} allows the VM managed
+The work in~\cite{younge2013enabling} allows the VM managed
 by the hypervisor Xen to access the GPUs in the physical
-node, with the implied limitations of restricting the nodes not
-to use more GPUs than the hosted in the machine and not
-to share idle GPUs to other machines. The solution presented 
-by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
+node, with the implied limitations that a node cannot 
+use more GPUs than those hosted in the machine, and idle GPUs can not be shared with other machines. The solution presented 
+by gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and makes
 them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor used, so
-does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
-This one continues without being integrated in a Cloud Computing Manager, though its main
+of virtualization strongly depends on the hypervisor and, so
+does its performance. Another similar solution is presented in gCloud~\cite{diab2013dynamic}. 
+iWhile this solution is not yet integrated in a Cloud Computing Manager, its main
 drawback is that the code of the applications must be modified
-in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
+in order to be run in the virtual-GPU environment. A runtime component to provide abstraction
 and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
-to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
-approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
-On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
-it is only focussed on compute intensive HPC applications.
+to isolate and share GPUs in a cluster for a set of applications. 
+The work introduced in~\cite{jungpgpu} is more mature, however,
+it is only focussed on compute-intensive HPC applications.
 
-The main idea of our proposal goes further and apart from
+Our proposal goes further and, in addition to 
 bringing solutions for all kind of HPC applications, it is aimed
-to boost the cluster flexibility in the use of GPUs.
+to boost flexibility in the use of GPUs.
 

### 2015-10-23 paper.tex
revision 6.4 (falta descripción infraestructura)

--- 2015-10-23 paper.tex
@@
 \subsection{Experimental Results}
-The tests were executed on 2 sets of nodes interconnected through a 1Gb Ethernet network.
+The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
 
 The first set, in charge of providing the cloud environment,
-was composed of 3 nodes, each one equipped with an Intel Xeon E7420 quadcore processor at
+consisted of 3 nodes, each equipped with an Intel Xeon E7420 quadcore processor at
 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
-To deploy an infrastructure as a cloud service, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
+To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
 A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
@@
 
-We have designed 6 different set-ups which can be divided
-into 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
-“Exclusive” mode will provide, at most, 4 accelerators. Whilst
-the number of available GPU in “Shared” mode will depend
-on the partition size. In this particular case, we have halved
-the memory, so we can use up to 8 GPUs. For each group, we
-have deployed virtual clusters of 1, 2 and 3 nodes where will
-run the processes of the application. The instances are based
-on the flavor “m1.medium” which determines VM of 2 cores
-and 4096 MB of memory.
+We have designed 6 different set-ups which can be divided into 2 groups: “exclusive GPUs” and “shared GPUs”. 
+The “exclusive” mode provides, at most, 4 accelerators. 
+Whilst the number of available GPU in “shared” mode will depend on the partition size. 
+In this particular case, we have halved the GPU memory, having as a result 8 partitions that can be addressed as independent GPUs.
+For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
+The instances are lanched with the Openstack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
 
-MCUDA-MEME~\cite{Liu2010} has been the application executed with
-different configurations in order to test the set-ups. MCUDA-
-MEME is an MPI application which is accelerated by GPUs.
-Besides, each of its processes must have a GPU to not abort 
-so, the number of GPUs will determine the number of processes that we can launch.
+MCUDA-MEME~\cite{Liu2010} was the application selected to test the set-ups. 
+Thus is an MPI application accelerated by GPUs, where each process must have access to a GPU.
+Therefore, the number of GPUs determines the maximum number of processes we can launch.
 
-The Figure~\ref{fig3} shows and compares the execution time of
-the application with different configurations over different setups. 
+Figure~\ref{fig3} compares the execution time of the application with different configurations over different setups. 
 We have used up to 3 nodes to spread the processes and launched up to 8 processes (only 4 in exclusive mode), one for each remote GPU.
-Firstly, we can state that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
-Secondly, the lowest execution time is obtained by both modes (exclusive and shared) when running their maximum number of processes with more than 1 node.
-It could mean that it is not worthy to continue giving resources to the application to scale, because the performance growth rate is barely increasing.
-Although, the time is lower sharing the GPUs the setup cannot take advantage of doubling the devices.
+We can first state that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
+In addition, the shortest execution time is obtained by both modes (exclusive and shared) when running their maximum number of processes with more than one node.
+This seems to imply that it is not worth to continue giving resources to the application to scale, because the performance growth rate is barely increasing.
+Although, the time is lower sharing the GPUs the setup cannot take advantage of an increase in the number of devices.
 
@@
   \includegraphics[width=\linewidth]{images/mcudameme-os.pdf}
-  \caption{Scalability results of executing MCUDAMEME with a different number of MPI processes}
+  \caption{Scalability results of MCUDAMEME with a different number of MPI processes.}
   \label{fig3}

### 2015-10-23 paper.tex
revision 6.5

@@
 \subsection{Discussion}
-Again, the networking issue is restricting the performance of our executions, but it has been discussed during virtually all the paper.
-However, it is not a new topic and in~\cite{tonithesis} is explained that improving the network infrastructure could make the difference.
+The network interconnect restricts the performance of our executions.
+This is not a new topic and in~\cite{tonithesis} is explained that improving the network infrastructure could make a big different for GPU virtualization.
 
-The most remarkable achievement is the wide range of possible configuration and the freedom that would have an administrator to set and reset a system to fit the 
+The most remarkable achievement is the wide range of possible configurations and the flexibility to set and reset a system to fit the 
 user requirements.
 
-The requests of GPU-capable hosts in a completely transparent way for the user can be fulfilled with an small investment in infrastructure and in maintenance.
-Energy can be saved, not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are not necessary.
+With this virtualization technology, the requests for GPU devices can be fulfilled with small investments in infrastructure and maintenance.
+Energy can be saved not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are idle.
 

### 2015-10-23 paper.tex
motivation revisado

--- 2015-10-23 paper.tex
@@
 
-Cloud computing, storage and other facilities are becoming 
-more popular during the last years thanks to the easy set up 
-process and the low required IT invest. The features provided 
-by the cloud services are useful for companies and individuals 
-which have access to the last technology without spending a 
-huge amount of money. Moreover, the systems offered by the cloud 
-companies matches with the ones required thanks to its flexibility.
+Cloud computing and storage have gained wide appeal  
+during the last years thanks to the easy set up 
+process and the low IT invest. The services offered in the cloud are useful for companies and individuals 
+that look for temporary access to the last technology without spending a 
+huge amount of money. In general, the systems provided by cloud 
+companies match with the those requested thanks to its flexibility.
 
-Two different kinds of cloud approaches are available:
-on the one hand, public facilities such as Amazon Web Services or 
-Microsoft Azure that allow users to chose a pre-configured system  
-and on the other hand, institutions/companies 
-are able to deploy their own cloud infrastructure by using software products 
-like OpenStack.
+Two different kinds of cloud approaches are currently available:
+on one hand, public facilities, such as AWS or 
+Microsoft Azure that allow users to chose a pre-configured system.   
+On the other hand, institutions/companies 
+are able to deploy their own cloud infrastructure by a software as OpenStack.
 
-In both cases, the power of current cloud facilities is based on the use 
-of Virtual Machines (VM) which  
-emulates a physical node with its own 
-CPUs, storage, network interconnection, OS and GPU. 
-However, to provide a VM with GPUs means
+In both cases, the power of current cloud facilities relies on the use 
+of VMs which  
+emulate a physical node with its own 
+CPUs, storage, network interconnection, GPU and operating system (OS). 
+However, providing a VM equipped with a GPU implies
 that the guest host must be equipped with at least one GPU, and
 once it has been assigned to an instance, no other VM can use
-it. Therefore, the number of VM that are allowed to use GPUs is 
+it. Therefore, the number of VMs that are allowed to use GPUs is 
 limited by the number of the installed devices. This lack of 
-flexibility is also noticed in the cluster configurations, where 
-depending on the users’ necessities, 
-two approaches to introduce GPGPU computation can be determined:
-the conservative, to have a lot of hosts with GPUs to be
-ensured that you are likely to satisfy all the requests; or a
-more daring approach, where you will count with some GPU-
-enabled nodes hoping not to arrive a large burst of requests
-so the cluster would run out of GPU resources.
+flexibility is also present in the cluster configurations, where 
+depending on the users' necessities, 
+two approaches to introduce GPGPU computation can be identified:
+the conservative option, features a lot of hosts with GPUs to 
+ensure that is likely to satisfy all requests; in a
+more daring approach, a reduced count of some GPU-enabled nodes is available, in the hope that a large burst of GPU requests 
+ does not exhaust the cluster's resources.
 
@@
 the desired flexibility because this software detaches the GPUs 
-from the nodes where they are installed allowing all nodes in a cluster
-not only use them as if they were local but also to share them 
-achieving a higher utilization rate and reducing the idle status time.
-Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
-environment should be able to use and share all the GPUs installed in 
-the physical nodes removing the current limitations.
+from the nodes where they are installed, allowing all nodes in a cluster
+not only to use them as if they were local, but also to share them 
+achieving a higher utilization rate and reducing the idle periods.
+Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
+environment can leverage and share all GPUs installed in 
+the physical nodes, removing past limitations.
 
-Our work have been focused on Amazon Web Service as example of public 
-solution and on OpenStack software acting as a private cloud environment. 
-In order to achieve it, these cloud facilities need to be provided with the {rCUDA} software and 
-in the case of OpenStack, a module for managing virtualized GPUs 
-requires to be implemented.
+Our work focuses on AWS as a public 
+solution, and on OpenStack software as a private cloud environment. 
+In order to achieve our goal, these cloud facilities need to be provided with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
+needs to be implemented.
 

### 2015-10-23 paper.tex

@@
 
-Thanks to the modular approach of the GPGPU Service, the Nova Project has not been modified and the tool could be easily ported to other Cloud Computing solutions.
+Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified and the tool can be easily ported to other Cloud Computing solutions.
 
 \subsection{Working Modes}
-The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
+The developed module allows users to create any of the scenarios displayed in Figure~\ref{fig2}. 
 The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
-We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
-Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
-A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
-So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
-As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.
+We refer to this as the ``mode'', with possible values being: ``exclusive'' or ``shared''. 
+Let us assume there are 4 GPU devices in the cluster (independently where they were hosted. 
+An example of those scenarios is shown in Figure~\ref{fig2}.
+There, in ``exclusive'' mode the instance monopolizes all the GPUs, while in the ``shared'' mode, the GPUs are partitioned. 
+As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
 

### 2015-10-23 paper.tex
revision 7

--- 2015-10-23 paper.tex
@@
 For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
-The instances are lanched with the Openstack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
+The instances are lanched with the OpenStack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
 
@@
 \label{sec:conclusions}
-We have seen a thoroughly study of the possibilities offered by AWS in everything to do with GPUs.
-Its limitations drove us to deploy our own private cloud to have more flexibility when working with these accelerators.
-So that, we have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
+We have presented a complete study of the possibilities offered by AWS when it comes to GPUs. 
+The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when working with these accelerators.
+For this purpose, we have introduced an extension of OpenStack which easily permits to create GPU-instances and manage the physical GPUs to better profit from them.
 
-Leaving aside the performance of the executions which we knew in advance that would be poor due to the interconnection network features, we have created new operation modes.
-They have brought interesting new ways of using GPUs in situations where the simple fact of having accessing a GPU is more important than having a powerful GPU to boost the performance.
+As we expected, due to the limited bandwidth of the interconnects used in the experimentation, the performance observed for the GPU virtualized scenarios in the test was quite low.
+On the other hand we have created new operation modes that open interesting new ways of using GPUs in situations where the having access to a GPU is more important than having a powerful GPU to boost the performance.
 

### 2015-10-23 paper.tex
revision 8

@@
 \label{sec:future}
-The first task that we should have keep in mind after this study is to upgrade the interconnection network to one more focused on HPC.
-So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
-The same thought, would guide us to try other Cloud vendors which more support for HPC.
+The first in the list of pending work is the upgrade of the interconnection network to one more prone on HPC.
+Porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
+The same thought, motivates us to try other Cloud vendors which better support for HPC.
 
-Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
-would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPUs.
+Looking for real situations where performance is less important than flexibility will drive us 
+to tools for easily deploying GPU-programming computer labs.
 
-Other interesting approach would be to make the system energy-aware. 
-Provided that a GPU could throw power information, a similar study based on the energy consumption could be done.
-
-Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
-With more scheduling policies the flexibility given by the GPGPU module for OpenStack would increase.
+Finally, an interesting future mode is to design new strategies in order to decide where a remote GPUs is created and assigned to a physical device
+In particular, innovate scheduling policies can enhance the flexibility given by the GPGPU module for OpenStack.
 

### 2015-10-23 paper.tex
revision 6.4

@@
 The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
+All the nodes were equipped with an Intel Xeon E7420 quadcore processor at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
 
 The first set, in charge of providing the cloud environment,
-consisted of 3 nodes, each equipped with an Intel Xeon E7420 quadcore processor at
-2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
-To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
-A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
-where the VMs are hosted, one network node that manages the logic
-virtual network for the VMs, and one or more compute nodes that 
-run the hypervisor and virtual machines.
+consisted of 3 nodes.
+To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisor.
+A fully-featured OpenStack deployment requires at least three nodes: a controller node manages the compute nodes 
+where the VMs are hosted, a network node that manages the logic virtual network for the VMs, and one or more compute nodes that 
+run the hypervisor and VMs.
 
-The second set, composed by 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
-Their system operates a Centos 6.6; and the GPUs use CUDA 6.5. 
-The GPU virtualization support is based on rCUDA v5.0.
+The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
+The OS was a Centos 6.6; and the GPUs used CUDA 6.5. 
+The GPU virtualization support was based on rCUDA v5.0.
 
@@
 
-The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
-TIN2011-23283 and~FEDER.
+This research was supported by Universitat Jaume I research project (P11B2013-21); and projects
+TIN2014-53495-R from MINECO and FEDER.
 

### 2015-10-23 paper.tex
mejorara la legibilidad de los escenarios aws
seccion 5 hasta 5.3 revisado

@@
 
-\section{\uppercase{Working with AWS}}
+\section{\uppercase{GPGPU share service with AWS}}
 \label{sec:workingaws}
-In this section, first of all, the current user-level features 
-provided by the Amazon Web Services are explained and analyzed 
+In this section, we briefly expose the current user-level features 
+provided by AWS and then analyze this service  
 from the point of view of the GPU usage. 
-Finally, the scenarios and experiments where virtualized GPUs are utilized are 
-described and first results are showed.
+Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of first results.
 
 \subsection{Current Features}
-This subsection explains the most common features checked during 
-the virtual machine creation step.
+This subsection identifies the most common features checked during 
+the VM's creation step.
 
@@
 
-An instance is a pre-configured virtual machine focused on an 
-specific target. Inside the large list of instances offered by AWS, 
-we can find: general purpose (T2, M4 and M3); computer science optimized 
+An instance is a pre-configured VM focused on an 
+specific target. Among the large list of instances offered by AWS, 
+we can find cases for general purpose (T2, M4 and M3); computer science optimized 
 (C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable instances (G2). Each kind of instance has its own purpose and its 
-own price. Moreover, each instance has different number of CPUs and network 
+GPU capable (G2). Each type of instance has its own purpose cost (price). 
+Moreover, each instance offers a different number of CPUs as well as network 
 interconnection which can be: low, medium, high or 10Gb.
 
-For this study, carried out in June 2015, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
-So, at that moment the instances used had the features presented in the Table~\ref{table:awsInstances}.
+For our study, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
+The instances available in that zone presented the features shown in the Table~\ref{table:awsInstances}.
 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{AWS available HPC instances.}
+\caption{Shown HPC instances available in US EAST (N. VIRGINIA) in June 2015.}
 \label{table:awsInstances}
@@
 
-The selection of these concrete types have the aim of using C3 family instances, which are not equipped with GPUs, as client whereas the instances of the G2 family are acting as servers.
+For the following experiments, we select C3 family instances, which are not equipped with GPUs, as clients; whereas instances of the G2 family will act as servers.
 
 \subsubsection{Networking}
-As it is shown in the Table~\ref{table:awsInstances}, each instance has a different 
-network and taking into account that the use of the bandwidth 
+Table~\ref{table:awsInstances} shows that each instance integrates a different 
+network. Thus, taking into account that bandwidth 
 is critical when GPU virtualization is applied, a simple test to verify the 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{IPERF results between instances.}
+\caption{IPERF results between selected instances.}
 \label{table:iperf}
@@
 
-The IPERF\footnote{\url{http://iperf.fr/}} tool has been executed between the instances showed in 
-the Table~\ref{table:iperf} in order to know the real bandwidth.
-Therefore, we can assure that the nomenclature ``High'' used in the 
-network column correspond to a 1~Gb interconnection network and the 10~Gb has 
+To evaluate the actual bandwidth, we executed 
+the IPERF\footnote{\url{http://iperf.fr/}} tool between the instances explained in 
+the Table~\ref{table:awsInstances} with the results shown in the Table~\ref{table:iperf}.
+From this experiment, we can assume that the nomenclature ``High'' corresponds to a 1~Gb interconnection network and the 10~Gb has 
 a real bandwidth of 7.5~Gb/s.
 Moreover, it looks like the instances equipped with a ``High'' interconnection network
-are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
-match perfectly. The gap between real and theoretical can be observed with 
-the 10~Gb interconnection which reaches up to 7.5~Gb/s.
+are constrained by software to 1~Gb/s since the theoretical and real bandwidth 
+match perfectly. The real gap between sustantiable and theoretical bandwidth can be observed with 
+the 10~Gb interconnection, which reaches up to 7.5~Gb/s.
 
 \subsubsection{GPUs}
-As it has been said before, an instance is based on a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
-Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}.
-As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
+An instance relies on a VM that runs on a real node with its own virtualized components. 
+Therefore it is possible that AWS uses a virtualization framework to offer GPU services to all the instances.
+Although, the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
+we need to verify that they are non-virtualized devices. For this proposal, we execute the NVIDIA SDK test {\tt bandwidthtest}. 
+As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an ad-hoc GPU. 
 % \begin{table}[htb]
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory in local GPU.}
 \label{table:bwt}
@@
 \subsection{Testbed Scenarios}
-In this section, the different client-server configurations 
-are explained. All the scenarios are based on the maximum number of instances 
+In this section, we introduce the different client-server configurations. 
+All scenarios are based on the maximum number of instances 
 that a user can freely select without submitting a formal request. 
-In particular, the maximum number of ``g2.2xlarge'' is 5 and 2 for the ``g2.8xlarge'' instances.
+In particular, the maximum number for ``g2.2xlarge'' is 5 and for ``g2.8xlarge'' it is 2.
 
-All the instances deployed operate the RHEL 7.1 64 bit system and the version 6.5 of CUDA.
+All the instances deployed operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
+We design three configuration scenarios for our tests:
 
@@
 \item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
-and used hardware configuration in real clusters, where each node is 
-populated with one GPU. Here, a simple node is able to use up to 5 GPUs 
+ hardware configuration in clusters, with each node  
+populated with one GPU. Here, a simple node can access to 5 GPUs 
 using the ``High'' network.
@@
 \item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
-with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
-a 10Gb network and the client is able to execute application using up to 8 GPUs.
+with 4 GPUs each one, and a GPU-less client. This scenario includes 
+a 10Gb network, and the client can execute the application using up to 8 GPUs.
 
-\item Scenario C (Figure~\ref{subfig:aws3}) joins both scenarios A and B. So a 
-single client, using a 1Gb network interconnection, has 13 GPUs that can be used as if they were locally installed.
+\item Scenario C (Figure~\ref{subfig:aws3}) combines scenarios A and B. A 
+single client, using a 1Gb network interconnection, can leverage 13 GPUs as if they were local.
 \end{itemize}
@@
 the {rCUDA} software needs to be installed in order to add 
-the required flexibility to the system. So, the {rCUDA} server is 
+the required flexibility to the system. The {rCUDA} server is 
 executed in the server nodes and the {rCUDA} libraries are installed 
@@
 
-Moreover, with the aim to know the expected network bandwidth using 
-a remote GPU, we have executed again the NVIDIA {\tt bandwidthtest} sample.
-The results observed in the Table~\ref{table:bwtrcuda} shows how the bandwidth achieved is limited 
+In order to evaluate the network bandwidth using 
+a remote GPU, we re-applied the NVIDIA {\tt bandwidthtest} sample.
+Table~\ref{table:bwtrcuda} shows how the bandwidth is limited 
 by the network.
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using {rCUDA}.}
+\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32 MB using pageable memory using {rCUDA}.}
 \label{table:bwtrcuda}
@@
 Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
-A & Host to Device & High& 127 MB/s \\ \hline
-A & Device to Host & High& 126 MB/s\\ \hline
-B & Host to Device & 10 Gb& 858 MB/s\\ \hline
-B & Device to Host & 10 Gb& 843 MB/s\\ \hline
+A & Host-to-Device & High& 127 MB/s \\ \hline
+A & Device-to-Host & High& 126 MB/s\\ \hline
+B & Host-to-Device & 10 Gb& 858 MB/s\\ \hline
+B & Device-to-Host & 10 Gb& 843 MB/s\\ \hline
 \end{tabular}\end{center}\end{table}
@@
 \subsection{Experimental results}
-In order to assess that GPU virtualization with AWS instances was working as expected, two applications were executed to analyze the results and their scalability.
+In this subsection we employ two GPU applications to assess the benefits of GPU virtualization.
 
-The first of them was {\tt MonteCarloMultiGPU} from NVIDIA SDK, an application purely computational. 
-It was launched with the default configuration ``scaling=weak'' which adjust the size of the problem depending on the number of accelerators.
-The Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and on the instances with local GPUs. 
-We have removed the results of the Scenario B for clarity ought to they are exactly the same that the obtained from Scenario C up to 8 GPUs. 
-In other words, we could state that the results until the 8th GPU would correspond to the Scenario B and thereafter to the Scenario C.
-Hence, in this particular case, rCUDA (using remote GPUs) is performing better than CUDA (with its local GPUs) due to rCUDA loads the libraries when its daemon starts.
-However, even with rCUDA we can see differences in the results between booth scenarios. We determined that the Scenario A could increase the throughput because
-unlike the other setups the GPUs are not sharing the PCI bus with other devices, as it happens in the 4-GPU instances.
+The first application is {\tt MonteCarloMultiGPU}, from the NVIDIA SDK, a code that is purely compute-based. 
+This was launched with the default configuration ``scaling=weak'', which adjusts the size of the problem depending on the number of accelerators.
+Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and using local GPUs. 
+For clarity, we have removed the results observed for Scenario B as they are exactly the same as those obtained from Scenario C with up to 8 GPUs. 
+In this particular case, rCUDA (using remote GPUs) outperforms CUDA (with its local GPUs) because the former loads the libraries when its daemon starts~\cite{tonithesis}.
+With rCUDA we can see differences in the results between booth scenarios. 
+Here, Scenario A can increase the throughput because the GPUs are not sharing the PCI bus with other devices due to each node only is equipped with one GPU.
+On the other hand, when 4-GPU instances (``g2.8xlarge'') are added (Scenario C), the PCI bus constrains the communication process reducing the scaling of performance.
 % \begin{figure}[!t]
@@
   \includegraphics[width=\linewidth]{images/mont.pdf}
-  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local and remote GPUs.}
+  \caption{NVIDIA SDK {\tt MonteCarloMultiGPU} execution using local (CUDA) and remote (Scenarios A and C) GPUs.}
   \label{fig:mont-opt}
@@
 
-The other application, LAMMPS\footnote{\url{http://lammps.sandia.gov}}, is a classical molecular dynamics simulator that can be applied at the atomic, meso, or continuum scale. 
-From the implementation perspective, this multi-process application employs at least one GPU to host its processes, but can benefit from the use of multiple GPUs.
-The Subfigure~\ref{subfig:lammps1} shows that running this application using remote GPUs does not offer any advantage in front of original CUDA. 
-Furthermore, for the execution on remote GPUs, a little difference between both networks can be appreciated, so, the ``High'' network performs worse than ``10 Gb'' network.
-Thus, if we compare the results of a LAMMPS execution with a larger problem (see Subfigure~\ref{subfig:lammps2}), although CUDA continues performing better, the interesting point is the execution time on remote GPUs.
-They are almost the same even with different networks what means that the transfers are making the interconnection network the bottleneck, and enlarging the problem size
+The second application, LAMMPS\footnote{\url{http://lammps.sandia.gov}}, 
+is a classical molecular dynamics simulator that can be applied at the atomic, 
+meso, or continuum scale. 
+From the implementation perspective, this multi-process application employs 
+at least one GPU to host its processes, but can benefit from the presence of multiple GPUs.
+Figure~\ref{subfig:lammps1} shows that, for this application, the use of remote GPUs 
+does not offer any advantage over the original CUDA. 
+Furthermore, for the execution on remote GPUs, the difference between both networks is small, although, the ``High'' network performs worse than ``10 Gb'' network.
+For execution of LAMMPS with a larger problem (see Figure~\ref{subfig:lammps2}), CUDA still performs better, but the interesting point is the execution time on remote GPUs. 
+They are almost the same even with different networks which indicates that the transfers turn the interconnection network into a bottleneck, and enlarging the problem size
 pays off the drawback of a slower network.
@@
 The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
-All the nodes were equipped with an Intel Xeon E7420 quadcore processor at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
 
 The first set, in charge of providing the cloud environment,
-consisted of 3 nodes.
-To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisor.
-A fully-featured OpenStack deployment requires at least three nodes: a controller node manages the compute nodes 
-where the VMs are hosted, a network node that manages the logic virtual network for the VMs, and one or more compute nodes that 
-run the hypervisor and VMs.
+consisted of 3 nodes, each equipped with an Intel Xeon E7420 quadcore processor at
+2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
+To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
+A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
+where the VMs are hosted, one network node that manages the logic
+virtual network for the VMs, and one or more compute nodes that 
+run the hypervisor and virtual machines.
 
-The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
-The OS was a Centos 6.6; and the GPUs used CUDA 6.5. 
-The GPU virtualization support was based on rCUDA v5.0.
+The second set, composed by 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
+Their system operates a Centos 6.6; and the GPUs use CUDA 6.5. 
+The GPU virtualization support is based on rCUDA v5.0.
 
@@
 For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
-The instances are lanched with the OpenStack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
+The instances are lanched with the Openstack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
 
@@
 \label{sec:conclusions}
-We have presented a complete study of the possibilities offered by AWS when it comes to GPUs. 
-The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when working with these accelerators.
-For this purpose, we have introduced an extension of OpenStack which easily permits to create GPU-instances and manage the physical GPUs to better profit from them.
+We have seen a thoroughly study of the possibilities offered by AWS in everything to do with GPUs.
+Its limitations drove us to deploy our own private cloud to have more flexibility when working with these accelerators.
+So that, we have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
 
-As we expected, due to the limited bandwidth of the interconnects used in the experimentation, the performance observed for the GPU virtualized scenarios in the test was quite low.
-On the other hand we have created new operation modes that open interesting new ways of using GPUs in situations where the having access to a GPU is more important than having a powerful GPU to boost the performance.
+Leaving aside the performance of the executions which we knew in advance that would be poor due to the interconnection network features, we have created new operation modes.
+They have brought interesting new ways of using GPUs in situations where the simple fact of having accessing a GPU is more important than having a powerful GPU to boost the performance.
 
@@
 \label{sec:future}
-The first in the list of pending work is the upgrade of the interconnection network to one more prone on HPC.
-Porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
-The same thought, motivates us to try other Cloud vendors which better support for HPC.
+The first task that we should have keep in mind after this study is to upgrade the interconnection network to one more focused on HPC.
+So that, porting the setup and the tests to an infrastructure with an Infiniband network would shed light on the viability of this kind of solutions for HPC.
+The same thought, would guide us to try other Cloud vendors which more support for HPC.
 
-Looking for real situations where performance is less important than flexibility will drive us 
-to tools for easily deploying GPU-programming computer labs.
+Looking for real situations where the performance is not so important as the flexibility to provide more accelerators, 
+would drive us to take the first steps through tools for easily deploying programming computer labs with the necessity for GPUs.
 
-Finally, an interesting future mode is to design new strategies in order to decide where a remote GPUs is created and assigned to a physical device
-In particular, innovate scheduling policies can enhance the flexibility given by the GPGPU module for OpenStack.
+Other interesting approach would be to make the system energy-aware. 
+Provided that a GPU could throw power information, a similar study based on the energy consumption could be done.
+
+Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
+With more scheduling policies the flexibility given by the GPGPU module for OpenStack would increase.
 
@@
 
-This research was supported by Universitat Jaume I research project (P11B2013-21); and projects
-TIN2014-53495-R from MINECO and FEDER.
+The researchers were supported by Universitat Jaume I research project (P11B2013-21), project
+TIN2011-23283 and~FEDER.
 

### 2015-10-23 paper.tex

@@
 \subfigure[5 remote GPUs over 1GbE.]{%
-\includegraphics[width=0.31\linewidth]{images/aws3.png}
+\includegraphics[width=0.315\linewidth]{images/aws3.pdf}
 \label{subfig:aws1}}
@@
 \subfigure[8 remote GPUs over 10GbE.]{%
-\includegraphics[width=0.31\linewidth]{images/aws2.png}
+\includegraphics[width=0.315\linewidth]{images/aws2.pdf}
 \label{subfig:aws2}}
 \subfigure[13 remote GPUs over 1GbE.]{%
-\includegraphics[width=0.31\linewidth]{images/aws1.png}
+\includegraphics[width=0.315\linewidth]{images/aws1.pdf}
 \label{subfig:aws3}}

### 2015-10-23 paper.tex
revisión 5.4

--- 2015-10-23 paper.tex
@@
 \subfigure[8 remote GPUs over 10GbE.]{%
-\includegraphics[width=0.31\linewidth]{images/aws2.png}
+\includegraphics[width=0.315\linewidth]{images/aws2.pdf}
 \label{subfig:aws2}}
 \subfigure[13 remote GPUs over 1GbE.]{%
-\includegraphics[width=0.31\linewidth]{images/aws1.png}
+\includegraphics[width=0.315\linewidth]{images/aws1.pdf}
 \label{subfig:aws3}}
-\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
+\caption{Configurations where an AWS instance, acting as a client, employs several remote GPUs.}
 \label{fig:aws}
@@
 \subsection{Discussion}
-Putting all together,  it is crystal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
-while running intensive parallel applications. 
-Hence, AWS's target is more oriented to offer the service than to provide performance.
-Also with the aim of plainly offering the service of GPGPU, AWS lacks in flexibility. 
-AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
+The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
+because neither the network nor the accelerators are powerful enough to deliver high performance
+when running compute-intensive parallel applications. 
+In other words, AWS is more oriented to offer service than to provide performance.
+Also, AWS aims to offer a GPGPU service but fails in offering flexibility.
+Concretely, AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
 
-Coming back to the Table~\ref{table:awsInstances} can be seen how not only the resources of the ``g2.8xlarge'' has been quadrupled, but also the cost per hour.
-So, in the case of having other necessities (talking in terms of the instances type), with remotes GPUs we could attach an accelerator to any kind of instance.
-Furthermore, reducing the budget in VM could be possible by customizing the resources of the available instances.
-If needed, we could work on an instance of 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs. 
-Avoiding to pay the double for features that we are not willing to use in ``g2.8xlarge''.
+Table~\ref{table:awsInstances} shows that the resources of the ``g2.8xlarge'' are quadrupled, but so is the cost per hour.
+Therefore, in the case of having other necessities (in terms of the instances type), with GPU virtualization technology we could attach an accelerator to any kind of instance.
+Furthermore, reducing the budget destinated to cloud services could be possible by customizing the resources of the available instances.
+For example, we can work on an instance with 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs, avoiding to pay the double for features that we are not willing to use in ``g2.8xlarge''.
 
-The last conclusion we will draw is related to GPU-shareability. 
-As we deduced, AWS has GPU-capable physical machines reserved which could be idle until a GPU-instance request were made.
-To have such an expensive accelerators in a standby state is counter-productive. So, it would make more sense if we had less devices, accessible from any machine, though.
+The last conclusion is related to GPU-shareability. 
+As we deduced, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
+To maintain an expensive accelerator in a standby state is counter-productive. 
+It makes more sense to dedicate less devices, accessible from any machine.
 

### 2015-10-23 paper.tex

--- 2015-10-23 paper.tex
@@
 \label{subfig:aws3}}
-\caption{Different configurations where an AWS instance acting as a client is using several remote GPUs.}
+\caption{Configurations where an AWS instance, acting as a client, employs several remote GPUs.}
 \label{fig:aws}

### 2015-10-23 paper.tex
seccion 5 hasta 5.3 revisado

--- 2015-10-23 paper.tex
@@
 For execution of LAMMPS with a larger problem (see Figure~\ref{subfig:lammps2}), CUDA still performs better, but the interesting point is the execution time on remote GPUs. 
-They are almost the same even with different networks which indicates that the transfers turn the interconnection network into a bottleneck, and enlarging the problem size
-pays off the drawback of a slower network.
+They are almost the same even with different networks which indicates that the transfers turn the interconnection network into a bottleneck. For this type of application's 
+patter, enlarging the problem size pays off the drawback of a slower network.
 

### 2015-10-26 paper.tex
abstract re-revisado

@@
 
-\keywords{Cloud Computing, GPU Virtualization, AWS, Resource Management}
+\keywords{Cloud Computing, GPU Virtualization, Amazon Web Services (AWS), OpenStack, Resource Management}
 
-\abstract{The use of accelerators, such as graphics processing units (GPUs), to reduce the execution time of compute-intensive applications has become more popular during the past few years. 
-These devices increment the computational power of a node thanks to its parallel architecture.
-This trend has led cloud vendors, as Amazon Web Services, or middlewares such as OpenStack to add to their current facilities   
-instances of virtual machines including GPUs. 
-To fullfil these needs, the guest hosts must be equipped with GPUs which will be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
-The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and applications' demand of accelerators. 
-Hence, we propose to decouple real GPUs from the nodes by using the virtualization technology rCUDA. 
+\abstract{The use of accelerators, such as graphics processing units (GPUs), to reduce the execution time of compute-intensive applications has become popular during the past few years. 
+These devices increment the computational power of a node thanks to their parallel architecture.
+This trend has led cloud service providers as Amazon or middlewares such as OpenStack to add virtual machines (VMs) including GPUs to their facilities   
+instances. 
+To fullfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and the applications' demand of accelerators. 
+Concretely, we propose to decouple real GPUs from the nodes by using the virtualization technology rCUDA. 
 With this software configuration, GPUs can be accessed from any VM avoiding the need  
-of having a physical GPUs in the guest host. 
-Moreover, we study the viability of this approach by using a public cloud 
-service configuration and develop a module for OpenStack middleware in order to 
+of placing a physical GPUs in each guest host. 
+Moreover, we study the viability of this approach using a public cloud 
+service configuration, and we develop a module for OpenStack in order to 
 add support for the virtualized devices and the logic to manage them. 
-The results demonstrate this is a viable configuration which adds flexibility to the current and well-known 
+The results demonstrate this is a viable configuration which adds flexibility to current and well-known 
 cloud solutions. 

### 2015-10-26 paper.tex
intro re-revisado

@@
 \label{sec:introduction}
-Nowadays, many cloud vendors have started offering virtual machines (VMs) with GPUs in order to provide
-GPGPU computation services to users. A few relevant examples include 
+Nowadays, many cloud vendors have started offering virtual machines (VMs) with graphics processing units (GPUs) in order to provide
+GPGPU (general-purpose GPU) computation services. A few relevant examples include 
 Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
@@
 Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
-In the public scope, one of the most popular cloud vendors is AWS, that offers a wide range of preconfigured instances ready to be launched.
+In the public scope, one of the most popular cloud vendors is AWS, which offers a wide range of preconfigured instances ready to be launched.
 Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
@@
 Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
-is not straightforward as is in a physical servers resulting into a low utilization rate.
+is as quite more complex than in physical servers, resulting in a low utilization rate.
 On one hand, instances in public clouds are not easily customizable. 
@@
 the number of options is reduced.
-Moreover, neither vendors nor tools are offer GPGPU wasting the opportunity of providing flexibility and boosty the utilization of GPU resources in both private or public clouds.
+Moreover, neither vendors nor tools offer GPGPU services, missing the opportunity of providing flexibility and boost the utilization of GPU resources in both private or public clouds.
 
-These devices were adopted with the release of the CUDA programming model for
-NVIDIA GPUs~\cite{cuda65}, which includes both high- and low-level application programming interfaces (APIs)
-to allow the use of NVIDIA devices as general-purpose
-accelerators.
-Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
-to offer a cross-vendor solution.
-
-The current approach to assemble a cluster with accelerators consist in installing 
-one or more of these devices in each compute node.
-However, these configurations often present a low
-utilization rate of the computational resources
- because of mismatches between the application's type of parallelism
-and the accelerators' architecture.
-
-Remote virtualization has been recently proposed to deal with the low-usage problem 
-(e.g. {rCUDA}~\cite{tonithesis},
-DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).
+Remote virtualization has been recently proposed to deal with the low-usage problem. Some relevant examples include {rCUDA}~\cite{tonithesis},
+DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}.
 
@@
 enable cluster configurations with fewer GPUs than nodes.  The goal is that 
-GPU-equipped nodes act as GPGPU (general-purpose GPU) servers, yielding a CUDA-sharing solution that potentially achieves
-a higher overall GPU load in the system.
+GPU-equipped nodes act as GPGPU servers, yielding a GPU-sharing solution that potentially achieves
+a higher overall utilization of the accelerators in the system.
 
-The main goals of this work are to study current cloud solutions from 
-the point of view of HPC field and GPU usage, and to analyze and improve 
+The main goals of this work are to study current cloud solutions in an HPC GPU-enabled scenario, and to analyze and improve 
 them by adding flexibility with GPU virtualization. In order to reach 
@@
 The rest of the paper is structured as follows. 
-In Section~\ref{sec:background} the technologies we introduce used in this work;
-Section~\ref{sec:state} summarizes previous work in this field; 
-Section~\ref{sec:motivation} states the motivation to carry out this work;
+In Section~\ref{sec:background} we introduce the technologies used in this work;
+Section~\ref{sec:state} summarizes related work; 
+Section~\ref{sec:motivation} states the motivation to address this task;
 the effort to use AWS is explained in Section~\ref{sec:workingaws};
 while the work with Openstack is described in Section~\ref{sec:gpgpuOS};
-finally Section~\ref{sec:conclusions} summarizes the advances and Section~\ref{sec:future} proposes the next steps
+finally, Section~\ref{sec:conclusions} summarizes the advances and Section~\ref{sec:future} outlines the next steps
  of this research.

### 2015-10-26 paper.tex
background re-revisado

@@
 to any NVIDIA GPU device present in a cluster from all compute
-nodes. The GPUs can be accessed and shared between nodes, and a single node can use all the graphic accelerators
+nodes. The GPUs can also be shared among nodes, and a single node can use all the graphic accelerators
 as if they were local.
-These features focus in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
+These features aim to attain higher accelerator utilization rates in the overall system while simultaneously reducing
 resource, space, and energy requirements~\cite{energy14}.
@@
 native CUDA libraries. Furthermore, the server middleware is executed in the
-cluster nodes from which the actual GPUs provide the requested GPGPU service.
+cluster nodes from which the actual GPUs provide the requested GPGPU services.
 To support a concurrent scenario where GPUs are shared between
@@
 CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
-APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
-Therefore, applications are not aware of the fact that they are being executed
+APIs (Application Programming Interface) as well as other frequently used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
+With this middleware, applications are not aware that they are executed
 on top of a virtualization layer.
 To deal with new GPU programming models, {rCUDA} has been recently extended to accommodate 
-directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.
+directive-based models such as OmpSs~\cite{repara15} and OpenACC~\cite{cluster15}.
 
 The integration of remote GPGPU virtualization with global
-resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
+resource schedulers such as SLURM~\cite{sbacpad14} completes this appealing 
 technology, making accelerator-enabled clusters more flexible and
@@
 that enables organizations and/or individuals to deploy
-services and applications on an on-demand basis. 
-These services replace local IT infrastructure for companies and provide agility and instant elasticity matching 
-perfectly with their software requirements.
+services and applications on demand. 
+These services replace company-owned local IT infrastructure and provide agility and instant elasticity matching 
+perfectly with enterprise software requirements.
 
 From the point of view of HPC, AWS offers high performance facilities via 
-instances equipped with GPUs and high performance network interconnection, which 
-are common resources leveraged in this research field.
+instances equipped with GPUs and high performance network interconnection.
 
@@
 
-OpenStack~\cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
-It is able to control large pools of compute, storage, and networking resources throughout a 
+OpenStack~\cite{OpenStack} is a cloud operating system (OS) that provides Infrastructure as a Service (IaaS). 
+OpenStack controls large pools of compute, storage, and networking resources throughout a 
 datacenter. All these resources are managed through a dashboard or an API that gives administrators 
@@
 or a command-line interface.  
-OpenStack supports most available hypervisors and handles provisioning 
+OpenStack supports most recent hypervisors and handles provisioning 
 and life-cycle management of VMs.
@@
 The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
-or software requirements and the ability to integrate with legacy systems and third party technologies.
+or software requirements, and the ability to integrate with legacy systems and third party technologies.
 
-From the HPC perspective, OpenStack offers high performance configurations using
-multiple supported hypervisors and different hardware architectures. Despite in OpenStack is possible to 
-specify some extra parameters requesting special resources, like GPUs, choosing a host based on the existence
+From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. Despite in OpenStack it is possible to 
+request special resources, such as GPUs, choosing a host based on the existence
 of a GPU is currently unsupported~\cite{OpenStackGPU}.

### 2015-10-26 paper.tex
revisión 6

--- 2015-10-26 paper.tex
@@
 
-\keywords{Cloud Computing, GPU Virtualization, Amazon Web Services (AWS), OpenStack, Resource Management}
+\keywords{Cloud Computing, GPU Virtualization, AWS, Resource Management}
 
-\abstract{The use of accelerators, such as graphics processing units (GPUs), to reduce the execution time of compute-intensive applications has become popular during the past few years. 
-These devices increment the computational power of a node thanks to their parallel architecture.
-This trend has led cloud service providers as Amazon or middlewares such as OpenStack to add virtual machines (VMs) including GPUs to their facilities   
-instances. 
-To fullfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
-The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and the applications' demand of accelerators. 
-Concretely, we propose to decouple real GPUs from the nodes by using the virtualization technology rCUDA. 
+\abstract{The use of accelerators, such as graphics processing units (GPUs), to reduce the execution time of compute-intensive applications has become more popular during the past few years. 
+These devices increment the computational power of a node thanks to its parallel architecture.
+This trend has led cloud vendors, as Amazon Web Services, or middlewares such as OpenStack to add to their current facilities   
+instances of virtual machines including GPUs. 
+To fullfil these needs, the guest hosts must be equipped with GPUs which will be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
+The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and applications' demand of accelerators. 
+Hence, we propose to decouple real GPUs from the nodes by using the virtualization technology rCUDA. 
 With this software configuration, GPUs can be accessed from any VM avoiding the need  
-of placing a physical GPUs in each guest host. 
-Moreover, we study the viability of this approach using a public cloud 
-service configuration, and we develop a module for OpenStack in order to 
+of having a physical GPUs in the guest host. 
+Moreover, we study the viability of this approach by using a public cloud 
+service configuration and develop a module for OpenStack middleware in order to 
 add support for the virtualized devices and the logic to manage them. 
-The results demonstrate this is a viable configuration which adds flexibility to current and well-known 
+The results demonstrate this is a viable configuration which adds flexibility to the current and well-known 
 cloud solutions. 
@@
 \label{sec:gpgpuOS}
-In this section we review the work related with OpenStack, from the development of an external module to handle remote GPUs to the physical setup and its testing.
+In this section we describe our work on OpenStack, from the development of an external module to handle remote GPUs to the physical setup and its validation.
 
@@
 The main idea is to evolve from the original OpenStack architecture (see Figure~\ref{subfig:os-orig})
- to a solution where a new shared service, responsible for all about GPU accelerators, becomes integrated into the architecture (see Figure~\ref{subfig:os-mod}).
+ to a solution where a new shared service, responsible for the GPU accelerators, becomes integrated into the architecture (see Figure~\ref{subfig:os-mod}).
 This new service brings more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
@@
 which splits the HTTP query in order to make use of both the GPGPU API for GPU-related operations, and the Nova API for the rest of the computations. 
-The new GPGPU Service allows to access GPUs in a VM, but also to create ``GPU-pools'', a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more VMs.
+The new GPGPU Service grants access to GPUs in a VM, but also allows the creation of ``GPU-pools'', consisting of a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more VMs.
 
@@
 
-Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified and the tool can be easily ported to other Cloud Computing solutions.
+Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified, and the tool can be easily ported to other Cloud Computing solutions.
 
 \subsection{Working Modes}
-The developed module allows users to create any of the scenarios displayed in Figure~\ref{fig2}. 
+The developed module allows users to set up any of the scenarios displayed in Figure~\ref{fig2}. 
 The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
 We refer to this as the ``mode'', with possible values being: ``exclusive'' or ``shared''. 
-Let us assume there are 4 GPU devices in the cluster (independently where they were hosted. 
-An example of those scenarios is shown in Figure~\ref{fig2}.
-There, in ``exclusive'' mode the instance monopolizes all the GPUs, while in the ``shared'' mode, the GPUs are partitioned. 
-As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
+Let us assume there are 4 GPU devices in the cluster (independently of where they were hosted). 
+An example of these scenarios is shown in Figure~\ref{fig2}.
+There, in the ``exclusive'' mode the instance monopolizes all the GPUs; while in the ``shared'' mode, the GPUs are partitioned. 
+As a result of sharing the GPU memory, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
 
 Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
-This behaviour is refereed as ``scope'' and it determines that a group of instances is logically connected to a pool of GPUs.
-Working with ``public scope'' (bottom row of Figure~\ref{fig2}) implies that the GPUs of a pool can be used simultaneously by all the instances linked to it.
+This behaviour is refereed as ``scope'', and it determines that a group of instances is logically connected to a pool of GPUs.
+Working with the ``public'' scope (bottom row of Figure~\ref{fig2}) implies that the GPUs of a pool can be used simultaneously by all the instances linked to it.
 Again, the GPU pool can be composed of ``exclusive'' or ``shared'' GPUs.
@@
 \subsection{User Interface}
-Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
+In order to deal with the new features, several modifications have been planned in the OpenStack Dashboard, they have not been implemented yet, though.
 First of all, the Instance Launch Panel should be extended with a new field, where the user could assign an existing GPU pool, create a new one, or keep the instance without accelerators of this kind.
@@
 The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
-All the nodes were equipped with an Intel Xeon E7420 quadcore processor at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
+All the nodes were equipped with an Intel Xeon E7420 quadcore processor, at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
 
@@
 consisted of 3 nodes.
-To deploy an infrastructure as a cloud service, we used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisor.
-A fully-featured OpenStack deployment requires at least three nodes: a controller node manages the compute nodes 
-where the VMs are hosted, a network node that manages the logic virtual network for the VMs, and one or more compute nodes that 
-run the hypervisor and VMs.
+To deploy an IaaS, we used OpenStack Icehouse version, and QEMU/KVM 0.12.1 as the hypervisor.
+A fully-featured OpenStack deployment requires at least three nodes: a controller manages the compute nodes 
+where the VMs are hosted; a network node manages the logic virtual network for the VMs, and one or more compute nodes run the hypervisor and VMs.
 
-The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each one. 
+The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each. 
 The OS was a Centos 6.6; and the GPUs used CUDA 6.5. 
@@
 
-We have designed 6 different set-ups which can be divided into 2 groups: “exclusive GPUs” and “shared GPUs”. 
-The “exclusive” mode provides, at most, 4 accelerators. 
-Whilst the number of available GPU in “shared” mode will depend on the partition size. 
-In this particular case, we have halved the GPU memory, having as a result 8 partitions that can be addressed as independent GPUs.
-For each group, we have deployed virtual clusters of 1, 2 and 3 nodes where the application processes will be executed. 
-The instances are lanched with the OpenStack predifined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4096 MB of memory.
+We have designed 6 different set-ups which can be divided into 2 groups: exclusive and shared GPUs. 
+The exclusive mode provides, at most, 4 accelerators. 
+The number of available GPU in “shared” mode will depend on the partition size. 
+In this particular case, we halved the GPU memory, resulting 8 partitions that can be addressed as independent GPUs.
+For each group, we deployed virtual clusters of 1, 2 and 3 nodes, where the application processes were executed. 
+The instances were launched with the OpenStack predefined flavor “m1.medium”, which determines a configuration of VMs consisting of 2 cores and 4~GB of memory.
 
 MCUDA-MEME~\cite{Liu2010} was the application selected to test the set-ups. 
-Thus is an MPI application accelerated by GPUs, where each process must have access to a GPU.
+Thus is an MPI software, where each process must have access to a GPU.
 Therefore, the number of GPUs determines the maximum number of processes we can launch.
@@
 Figure~\ref{fig3} compares the execution time of the application with different configurations over different setups. 
-We have used up to 3 nodes to spread the processes and launched up to 8 processes (only 4 in exclusive mode), one for each remote GPU.
-We can first state that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
+We used up to 3 nodes to spread the processes and launched up to 8 processes (only 4 in exclusive mode), one per remote GPU.
+We can first observe that the performance is higher with more than one node, because the traffic network is distributed among the nodes.
 In addition, the shortest execution time is obtained by both modes (exclusive and shared) when running their maximum number of processes with more than one node.
-This seems to imply that it is not worth to continue giving resources to the application to scale, because the performance growth rate is barely increasing.
-Although, the time is lower sharing the GPUs the setup cannot take advantage of an increase in the number of devices.
+This seems to imply that it is not worth to scale (increase) the number of resources, because the performance growth rate is barely increasing.
+Although, the time is lower when the GPUs are shared, the setup cannot take advantage of an increase in the number of devices.
 
@@
 The network interconnect restricts the performance of our executions.
-This is not a new topic and in~\cite{tonithesis} is explained that improving the network infrastructure could make a big different for GPU virtualization.
+The analysis in~\cite{tonithesis} reveals that improving the network infrastructure can make a big different for GPU virtualization.
 
-The most remarkable achievement is the wide range of possible configurations and the flexibility to set and reset a system to fit the 
-user requirements.
+The most remarkable achievement is the wide range of possible configurations and the flexibility to adapt a system to fit the user requirements.
 
-With this virtualization technology, the requests for GPU devices can be fulfilled with small investments in infrastructure and maintenance.
-Energy can be saved not only thanks to the remote access and the ability to emulate several GPUs from a real one, but also  by grouping the accelerators in a single machine (when possible) or turning down nodes when their GPUs are idle.
+In addition, with this virtualization technology, the requests for GPU devices can be fulfilled with small investment in infrastructure and maintenance.
+Energy can be saved not only thanks to the remote access and the ability to emulate several GPUs using only a few real ones, but also by consolidating the accelerators in a single machine (when possible), or turning down nodes when their GPUs are idle.
 

### 2015-10-26 paper.tex
revisión 7

@@
 We have presented a complete study of the possibilities offered by AWS when it comes to GPUs. 
-The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when working with these accelerators.
-For this purpose, we have introduced an extension of OpenStack which easily permits to create GPU-instances and manage the physical GPUs to better profit from them.
+The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when dealing with these accelerators.
+For this purpose, we have introduced an extension of OpenStack which can be easily exploited to create GPU-instances as well as manage the physical GPUs to better profit from them.
 
-As we expected, due to the limited bandwidth of the interconnects used in the experimentation, the performance observed for the GPU virtualized scenarios in the test was quite low.
-On the other hand we have created new operation modes that open interesting new ways of using GPUs in situations where the having access to a GPU is more important than having a powerful GPU to boost the performance.
+As we expected, due to the limited bandwidth of the interconnects used in the experimentation, the performances observed for the GPU virtualized scenarios in the tests were quite low.
+On the other hand, we have created new operation modes that open interesting new ways to leverage GPUs in situations where having access to a GPU is more important than having a powerful GPU to boost the performance.
 
@@
 \label{sec:future}
-The first in the list of pending work is the upgrade of the interconnection network to one more prone on HPC.
-Porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
-The same thought, motivates us to try other Cloud vendors which better support for HPC.
+The first item in the list of pending work is an upgrade of the network to an interconnect that is more prone to HPC.
+In particular, porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
+Similar reasons, motivate us to try other Cloud vendors which better support for HPC.
 
-Looking for real situations where performance is less important than flexibility will drive us 
-to tools for easily deploying GPU-programming computer labs.
+Looking for real situations where performance is less important than flexibility will drive us to explore alternative tools to easily deploy GPU-programming computer labs.
 
-Finally, an interesting future mode is to design new strategies in order to decide where a remote GPUs is created and assigned to a physical device
-In particular, innovate scheduling policies can enhance the flexibility given by the GPGPU module for OpenStack.
+Finally, an interesting future work is to design new strategies in order to decide where a remote GPUs is created and assigned to a physical device
+Concretely, to innovate scheduling policies can enhance the flexibility offered by the GPGPU module for OpenStack.
 

### 2015-10-26 paper.tex
motivation re-revisado

--- 2015-10-26 paper.tex
@@
 \label{sec:state}
-Our solutions to the defficiency exposed in the previous section is based on GPU virtualization, sharing resources in order
+Our solutions to the defficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
 to attain a fair balance between supply and demand. 
-While several efforst with the same goal have been made in the past, as exposed next, 
-none of them is so ambitious as ours.
+While several efforts with the same goal have been initiated in the past, as exposed next, 
+none of them is as ambitious as ours.
 
 The work in~\cite{younge2013enabling} allows the VM managed
-by the hypervisor Xen to access the GPUs in the physical
-node, with the implied limitations that a node cannot 
-use more GPUs than those hosted in the machine, and idle GPUs can not be shared with other machines. The solution presented 
-by gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and makes
+by the Xen hypervisor to access the GPUs in a physical
+node, but with this solution a node cannot 
+use more GPUs than those hosted in the machine, and an idle GPU can not be shared with other machines. The solution presented 
+by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
 them accessible for any VM in the cluster. However, this kind
-of virtualization strongly depends on the hypervisor and, so
-does its performance. Another similar solution is presented in gCloud~\cite{diab2013dynamic}. 
-iWhile this solution is not yet integrated in a Cloud Computing Manager, its main
-drawback is that the code of the applications must be modified
-in order to be run in the virtual-GPU environment. A runtime component to provide abstraction
+of virtualization strongly depends on the hypervisor, and so
+does its performance. Similar solution is presented in gCloud~\cite{diab2013dynamic}. 
+While this solution is not yet integrated in a Cloud Computing Manager, its main
+drawback is that the application's code must be modified
+in order to run in the virtual-GPU environment. A runtime component to provide abstraction
 and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
 to isolate and share GPUs in a cluster for a set of applications. 
-The work introduced in~\cite{jungpgpu} is more mature, however,
+The work introduced in~\cite{jungpgpu} is more mature; however,
 it is only focussed on compute-intensive HPC applications.
@@
 Our proposal goes further and, in addition to 
-bringing solutions for all kind of HPC applications, it is aimed
+bringing solutions for all kind of HPC applications, aims 
 to boost flexibility in the use of GPUs.
@@
 Cloud computing and storage have gained wide appeal  
-during the last years thanks to the easy set up 
-process and the low IT invest. The services offered in the cloud are useful for companies and individuals 
-that look for temporary access to the last technology without spending a 
+during the last years thanks to a relatively easy configuration  
+process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
+that look for temporary access to state-of-the-art technology without spending a 
 huge amount of money. In general, the systems provided by cloud 
-companies match with the those requested thanks to its flexibility.
+companies match those requested by the clients thanks to its flexibility.
 
-Two different kinds of cloud approaches are currently available:
+Two different types of cloud approaches are currently available:
 on one hand, public facilities, such as AWS or 
-Microsoft Azure that allow users to chose a pre-configured system.   
+Microsoft Azure, that allow users to chose a pre-configured system.   
 On the other hand, institutions/companies 
-are able to deploy their own cloud infrastructure by a software as OpenStack.
+are able to deploy their own cloud infrastructure using a software as OpenStack.
 
@@
 emulate a physical node with its own 
-CPUs, storage, network interconnection, GPU and operating system (OS). 
+CPUs, storage, network interconnection, GPU and OS. 
 However, providing a VM equipped with a GPU implies
-that the guest host must be equipped with at least one GPU, and
+that the guest host must be equipped with at least one real GPU, and
 once it has been assigned to an instance, no other VM can use
-it. Therefore, the number of VMs that are allowed to use GPUs is 
+it. Therefore, the number of VMs that can access the GPUs is 
 limited by the number of the installed devices. This lack of 
-flexibility is also present in the cluster configurations, where 
+flexibility is also present in cluster configurations, where 
 depending on the users' necessities, 
-two approaches to introduce GPGPU computation can be identified:
-the conservative option, features a lot of hosts with GPUs to 
-ensure that is likely to satisfy all requests; in a
-more daring approach, a reduced count of some GPU-enabled nodes is available, in the hope that a large burst of GPU requests 
+two approaches to enable GPGPU computation can be identified:
+the conservative option integrates many hosts with GPUs to 
+ensure that it is likely to satisfy all requests; in a
+more daring approach, a reduced count of some GPU-enabled nodes is configured, in the hope that a large burst of GPU requests 
  does not exhaust the cluster's resources.
 
-The use of a virtualization technology such as {rCUDA} adds 
+The use of a virtualization technology, such as {rCUDA}, adds 
 the desired flexibility because this software detaches the GPUs 
 from the nodes where they are installed, allowing all nodes in a cluster
-not only to use them as if they were local, but also to share them 
+to use them as if they were local, as well as to share them, 
 achieving a higher utilization rate and reducing the idle periods.
 Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
-environment can leverage and share all GPUs installed in 
+environment can leverage and access all GPUs installed in 
 the physical nodes, removing past limitations.
 
-Our work focuses on AWS as a public 
-solution, and on OpenStack software as a private cloud environment. 
-In order to achieve our goal, these cloud facilities need to be provided with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
+Our work selects AWS as a public 
+solution, and OpenStack as a private cloud environment. 
+In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
 needs to be implemented.

### 2015-10-26 paper.tex
hasta el 5.1 re-revisado

@@
 In this section, we briefly expose the current user-level features 
-provided by AWS and then analyze this service  
-from the point of view of the GPU usage. 
-Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of first results.
+provided by AWS, and then analyze this service  
+from the point of view of GPU usage. 
+Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of initial results.
 
 \subsection{Current Features}
-This subsection identifies the most common features checked during 
+This subsection identifies several common features checked during 
 the VM's creation step.
@@
 specific target. Among the large list of instances offered by AWS, 
-we can find cases for general purpose (T2, M4 and M3); computer science optimized 
-(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
-GPU capable (G2). Each type of instance has its own purpose cost (price). 
-Moreover, each instance offers a different number of CPUs as well as network 
-interconnection which can be: low, medium, high or 10Gb.
+we can find specialized versions for general-purpose (T2, M4 and M3); computer science (C4 and C3); memory (R3); storage (I2 and D2) and 
+GPU capable (G2). Each type of instance has its own purpose and cost (price). 
+Moreover, each type offers a different number of CPUs as well as network 
+interconnection, which can be: low, medium, high or 10Gb.
 
 For our study, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
-The instances available in that zone presented the features shown in the Table~\ref{table:awsInstances}.
+The instances available in that case present the features reported in Table~\ref{table:awsInstances}.
 
@@
 
-For the following experiments, we select C3 family instances, which are not equipped with GPUs, as clients; whereas instances of the G2 family will act as servers.
+For the following experiments, we select C3 family instances, which are not equipped with GPUs, as clients; whereas instances of the G2 family will act as GPU-enabled servers.
 
@@
 Table~\ref{table:awsInstances} shows that each instance integrates a different 
-network. Thus, taking into account that bandwidth 
-is critical when GPU virtualization is applied, a simple test to verify the 
-real network bandwidth is needed.
+network. As the bandwidth 
+is critical when GPU virtualization is applied, we first perform a simple test to verify the 
+real network bandwidth.
 
@@
 To evaluate the actual bandwidth, we executed 
-the IPERF\footnote{\url{http://iperf.fr/}} tool between the instances explained in 
-the Table~\ref{table:awsInstances} with the results shown in the Table~\ref{table:iperf}.
-From this experiment, we can assume that the nomenclature ``High'' corresponds to a 1~Gb interconnection network and the 10~Gb has 
+the IPERF\footnote{\url{http://iperf.fr/}} tool between the instances described in 
+Table~\ref{table:awsInstances}, with the results shown in Table~\ref{table:iperf}.
+From this experiment, we can derive that network ``High'' corresponds to a 1~Gb interconnect while ``10~Gb'' has 
 a real bandwidth of 7.5~Gb/s.
-Moreover, it looks like the instances equipped with a ``High'' interconnection network
-are constrained by software to 1~Gb/s since the theoretical and real bandwidth 
-match perfectly. The real gap between sustantiable and theoretical bandwidth can be observed with 
+Moreover, it seems that the bandwidth of the instances equipped with a ``High'' interconnection network
+is constrained by software to 1~Gb/s since the theoretical and real bandwidth 
+match perfectly. The real gap between sustained and theoretical bandwidth can be observed with 
 the 10~Gb interconnection, which reaches up to 7.5~Gb/s.
@@
 An instance relies on a VM that runs on a real node with its own virtualized components. 
-Therefore it is possible that AWS uses a virtualization framework to offer GPU services to all the instances.
-Although, the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
-we need to verify that they are non-virtualized devices. For this proposal, we execute the NVIDIA SDK test {\tt bandwidthtest}. 
-As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an ad-hoc GPU. 
+Therefore AWS can leverage a virtualization framework to offer GPU services to all the instances.
+Although the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
+we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
+As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an actual GPU. 
 % \begin{table}[htb]
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory in local GPU.}
+\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a local GPU.}
 \label{table:bwt}
@@
 that a user can freely select without submitting a formal request. 
-In particular, the maximum number for ``g2.2xlarge'' is 5 and for ``g2.8xlarge'' it is 2.
+In particular, the maximum number for ``g2.2xlarge'' is 5; for ``g2.8xlarge'' it is 2.
 
-All the instances deployed operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
+All instances operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
 We design three configuration scenarios for our tests:
@@
 \begin{itemize}
-\item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
- hardware configuration in clusters, with each node  
-populated with one GPU. Here, a simple node can access to 5 GPUs 
+\item Scenario A (Figure~\ref{subfig:aws1}) shows a common 
+configuration in GPU-accelerated clusters, with each node  
+populated with a single GPU. Here, a node can access 5 GPUs 
 using the ``High'' network.
 
-\item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
-with 4 GPUs each one, and a GPU-less client. This scenario includes 
+\item Scenario B (Figure~\ref{subfig:aws2}) is composed of 2 server nodes, equipped 
+with 4 GPUs each, and a GPU-less client. This scenario includes 
 a 10Gb network, and the client can execute the application using up to 8 GPUs.
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32 MB using pageable memory using {rCUDA}.}
+\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
 \label{table:bwtrcuda}

### 2015-10-26 paper.tex
revisión 5 sergio

--- 2015-10-26 paper.tex
@@
 Once the scenarios are configured from the point of view of hardware, 
-the {rCUDA} software needs to be installed in order to add 
+the {rCUDA} middleware needs to be installed in order to add 
 the required flexibility to the system. The {rCUDA} server is 
-executed in the server nodes and the {rCUDA} libraries are installed 
-in the node that is acting as a client.
+executed in the server nodes and the {rCUDA} libraries are invoked from the node that acts as client.
 
-In order to evaluate the network bandwidth using 
-a remote GPU, we re-applied the NVIDIA {\tt bandwidthtest} sample.
-Table~\ref{table:bwtrcuda} shows how the bandwidth is limited 
-by the network.
+In order to evaluate the network bandwidth using a remote GPU, we re-applied NVIDIA SDK {\tt bandwidthtest}.
+Table~\ref{table:bwtrcuda} exposes that the bandwidth is limited by the network.
 
@@
 
-The first application is {\tt MonteCarloMultiGPU}, from the NVIDIA SDK, a code that is purely compute-based. 
-This was launched with the default configuration ``scaling=weak'', which adjusts the size of the problem depending on the number of accelerators.
-Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios of Figure~\ref{fig:aws} and using local GPUs. 
+The first application is {\tt MonteCarloMultiGPU}, from the NVIDIA SDK, a code that is compute bound (its execution barely involves memory operations). 
+This was launched with the default configuration, ``scaling=weak'', which adjusts the size of the problem depending on the number of accelerators.
+Figure~\ref{fig:mont-opt} depicts the options per second calculated by the application running on the scenarios in Figure~\ref{fig:aws} as well as using local GPUs. 
 For clarity, we have removed the results observed for Scenario B as they are exactly the same as those obtained from Scenario C with up to 8 GPUs. 
-In this particular case, rCUDA (using remote GPUs) outperforms CUDA (with its local GPUs) because the former loads the libraries when its daemon starts~\cite{tonithesis}.
-With rCUDA we can see differences in the results between booth scenarios. 
-Here, Scenario A can increase the throughput because the GPUs are not sharing the PCI bus with other devices due to each node only is equipped with one GPU.
-On the other hand, when 4-GPU instances (``g2.8xlarge'') are added (Scenario C), the PCI bus constrains the communication process reducing the scaling of performance.
-% \begin{figure}[!t]
-%   \centering
-%   \includegraphics[width=\linewidth]{images/aws-mont2.png}
-%   \caption{Montecarlo execution time}
-%   \label{fig:mont-exe}
-% \end{figure}
+In this particular case, rCUDA (remote GPUs) outperforms CUDA (local GPUs) because the former loads the libraries when the daemon is started~\cite{tonithesis}.
+With rCUDA we can observe differences in the results between both scenarios. 
+Here, Scenario A can increase the throughput because the GPUs do not share the PCI bus with other devices as each node only is equipped with one GPU.
+On the other hand, when the 4-GPU instances (``g2.8xlarge'') are added (Scenario C), the PCI bus constrains the communication bandwidth, hurting the scalability.
+
 \begin{figure}[htb]
@@
 does not offer any advantage over the original CUDA. 
-Furthermore, for the execution on remote GPUs, the difference between both networks is small, although, the ``High'' network performs worse than ``10 Gb'' network.
-For execution of LAMMPS with a larger problem (see Figure~\ref{subfig:lammps2}), CUDA still performs better, but the interesting point is the execution time on remote GPUs. 
-They are almost the same even with different networks which indicates that the transfers turn the interconnection network into a bottleneck. For this type of application's 
-patter, enlarging the problem size pays off the drawback of a slower network.
+Furthermore, for the execution on remote GPUs, the difference between both networks is small, although, the results observed with the ``High'' network are worse than those obtained with the ``10 Gb'' network.
+In execution of LAMMPS on a larger problem (see Figure~\ref{subfig:lammps2}), CUDA still performs better, but the interesting point is the execution time when using remote GPUs. 
+These are almost the same even with different networks, which indicates that the transfers turn the interconnection network into a bottleneck. 
+For this type of application, enlarging the problem size compensates the negative effect of a slower network.
 
@@
 when running compute-intensive parallel applications. 
-In other words, AWS is more oriented to offer service than to provide performance.
-Also, AWS aims to offer a GPGPU service but fails in offering flexibility.
-Concretely, AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
+In other words, AWS is more oriented toward offering a general-purpose service than to provide high performance.
+Also, AWS fails in offering flexibility as it forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
 
 Table~\ref{table:awsInstances} shows that the resources of the ``g2.8xlarge'' are quadrupled, but so is the cost per hour.
-Therefore, in the case of having other necessities (in terms of the instances type), with GPU virtualization technology we could attach an accelerator to any kind of instance.
-Furthermore, reducing the budget destinated to cloud services could be possible by customizing the resources of the available instances.
-For example, we can work on an instance with 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs, avoiding to pay the double for features that we are not willing to use in ``g2.8xlarge''.
+Therefore, in the case of having other necessities (in terms of the instances type), using GPU virtualization technology we could in principle attach an accelerator to any type of instance.
+Furthermore, reducing the budget spent in cloud services is possible by customizing the resources of the available instances.
+For example, we can work on an instance with 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs, avoiding to pay the double for features that we do not need in ``g2.8xlarge''.
 
@@
 As we deduced, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
-To maintain an expensive accelerator in a standby state is counter-productive. 
+Maintaining an expensive accelerator in a standby state is counter-productive. 
 It makes more sense to dedicate less devices, accessible from any machine.

### 2015-10-26 paper.tex
ispell pasado

--- 2015-10-26 paper.tex
@@
 instances. 
-To fullfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
+To fulfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
 The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and the applications' demand of accelerators. 
@@
 In the public scope, one of the most popular cloud vendors is AWS, which offers a wide range of preconfigured instances ready to be launched.
-Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
-Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
+Alternatively, owning the proper infrastructure, a private cloud can be deployed using a specific middleware such as
+OpenStack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
 
@@
 
-Roughtly speaking these virtualization frameworks 
+Roughly speaking these virtualization frameworks 
 enable cluster configurations with fewer GPUs than nodes.  The goal is that 
@@
 \label{sec:state}
-Our solutions to the defficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
+Our solutions to the deficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
 to attain a fair balance between supply and demand. 
@@
 we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
-As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an actual GPU. 
+As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, revealing that the accelerator is an actual GPU. 
 % \begin{table}[htb]
@@
 
-Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified and the tool can be easily ported to other Cloud Computing solutions.
+Thanks to the modular approach of the GPGPU Service, the Nova Project does not need to be modified, and the tool can be easily ported to other Cloud Computing solutions.
 
 \subsection{Working Modes}
-The developed module allows users to create any of the scenarios displayed in Figure~\ref{fig2}. 
+The developed module allows users to set up any of the scenarios displayed in Figure~\ref{fig2}. 
 The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
 We refer to this as the ``mode'', with possible values being: ``exclusive'' or ``shared''. 
-Let us assume there are 4 GPU devices in the cluster (independently where they were hosted. 
-An example of those scenarios is shown in Figure~\ref{fig2}.
-There, in ``exclusive'' mode the instance monopolizes all the GPUs, while in the ``shared'' mode, the GPUs are partitioned. 
-As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
+Let us assume there are 4 GPU devices in the cluster (independently of where they were hosted). 
+An example of these scenarios is shown in Figure~\ref{fig2}.
+There, in the ``exclusive'' mode the instance monopolizes all the GPUs; while in the ``shared'' mode, the GPUs are partitioned. 
+As a result of sharing the GPU memory, the instance will be able to work with up to 8 GPUs, provided that each partition can be addressed as an independent GPU.
 
 Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
-This behaviour is refereed as ``scope'' and it determines that a group of instances is logically connected to a pool of GPUs.
-Working with ``public scope'' (bottom row of Figure~\ref{fig2}) implies that the GPUs of a pool can be used simultaneously by all the instances linked to it.
+This behavior is refereed as ``scope'', and it determines that a group of instances is logically connected to a pool of GPUs.
+Working with the ``public'' scope (bottom row of Figure~\ref{fig2}) implies that the GPUs of a pool can be used simultaneously by all the instances linked to it.
 Again, the GPU pool can be composed of ``exclusive'' or ``shared'' GPUs.

### 2015-10-26 paper.tex
spell sergio

@@
 instances. 
-To fulfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
+To fullfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
 The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and the applications' demand of accelerators. 
@@
 In the public scope, one of the most popular cloud vendors is AWS, which offers a wide range of preconfigured instances ready to be launched.
-Alternatively, owning the proper infrastructure, a private cloud can be deployed using a specific middleware such as
-OpenStack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
+Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
+Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.
 
@@
 
-Roughly speaking these virtualization frameworks 
+Roughtly speaking these virtualization frameworks 
 enable cluster configurations with fewer GPUs than nodes.  The goal is that 
@@
 \label{sec:state}
-Our solutions to the deficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
+Our solutions to the defficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
 to attain a fair balance between supply and demand. 
@@
 The work introduced in~\cite{jungpgpu} is more mature; however,
-it is only focussed on compute-intensive HPC applications.
+it is only focused on compute-intensive HPC applications.
 

### 2015-10-26 paper.tex
máquina

--- 2015-10-26 paper.tex
@@
 instances. 
-To fulfil these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
+To fulfill these needs, the guest hosts must be equipped with GPUs which, unfortunately, will be barely utilized if a non GPU-enabled VM is running in the host.
 The solution presented in this work is based on GPU virtualization and shareability in order to reach an equilibrium between service supply and the applications' demand of accelerators. 
@@
 Table~\ref{table:awsInstances} shows that the resources of the ``g2.8xlarge'' are quadrupled, but so is the cost per hour.
-Therefore, in the case of having other necessities (in terms of the instances type), using GPU virtualization technology we could in principle attach an accelerator to any type of instance.
+Therefore, in the case of having other necessities (instances type), using GPU virtualization technology we could in principle attach an accelerator to any type of instance.
 Furthermore, reducing the budget spent in cloud services is possible by customizing the resources of the available instances.
@@
 
-The last conclusion is related to GPU-shareability. 
-As we deduced, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
+In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
 Maintaining an expensive accelerator in a standby state is counter-productive. 
@@
 First of all, the Instance Launch Panel should be extended with a new field, where the user could assign an existing GPU pool, create a new one, or keep the instance without accelerators of this kind.
-When the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
+When the option ``New GPU Pool'' is chosen, fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
 Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).
@@
 
-The first set, in charge of providing the cloud environment,
-consisted of 3 nodes.
+The first set, in charge of providing the cloud environment, consisted of 3 nodes.
 To deploy an IaaS, we used OpenStack Icehouse version, and QEMU/KVM 0.12.1 as the hypervisor.
@@
 The second set, composed of 4 nodes, were auxiliary servers with a Tesla C1060 GPU each. 
-The OS was a Centos 6.6; and the GPUs used CUDA 6.5. 
-The GPU virtualization support was based on rCUDA v5.0.
+The OS was a Centos 6.6; the GPUs used CUDA 6.5; and rCUDA v5.0 as GPU virtualization framework.
 
@@
 The constraints imposed by this service motivated us to deploy our own private cloud, in order to gain flexibility when dealing with these accelerators.
-For this purpose, we have introduced an extension of OpenStack which can be easily exploited to create GPU-instances as well as manage the physical GPUs to better profit from them.
+For this purpose, we have introduced an extension of OpenStack which can be easily exploited to create GPU-instances as well as manage the physical GPUs to better profit.
 
@@
 
-This research was supported by Universitat Jaume I research project (P11B2013-21); and projects
+This research was supported by Universitat Jaume~I research project (P11B2013-21); and project
 TIN2014-53495-R from MINECO and FEDER.

### 2016-02-02 paper.tex
reviews

@@
 \email{\{siserte, fclement, adcastel, mayo, quintana\}@uji.es}}
+
+%\author{Ommitted for blind review}
 
@@
 
-Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
-is as quite more complex than in physical servers, resulting in a low utilization rate.
+Unfortunately, sharing GPU resources among multiple VMs in cloud environments is more complex than in physical servers.
 On one hand, instances in public clouds are not easily customizable. 
@@
 the number of options is reduced.
-Moreover, neither vendors nor tools offer GPGPU services, missing the opportunity of providing flexibility and boost the utilization of GPU resources in both private or public clouds.
+As a result, neither vendors nor tools offer GPGPU services.
 
@@
 The main goals of this work are to study current cloud solutions in an HPC GPU-enabled scenario, and to analyze and improve 
-them by adding flexibility with GPU virtualization. In order to reach 
-this goal, we select rCUDA, a virtualization tool that is possibly the more complete 
-and up-to-date for NVIDIA GPUs.
+them by adding flexibility with GPU virtualization. 
+In order to reach this goal, we select rCUDA, a virtualization tool that is possibly the more complete and up-to-date for NVIDIA GPUs.
 
@@
 
-From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. Despite in OpenStack it is possible to 
-request special resources, such as GPUs, choosing a host based on the existence
-of a GPU is currently unsupported~\cite{OpenStackGPU}.
+From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. 
+Even though in OpenStack it is possible to work with GPUs, the Nova project is not currently aware of them~\cite{OpenStackGPU}.
 
@@
 node, but with this solution a node cannot 
-use more GPUs than those hosted in the machine, and an idle GPU can not be shared with other machines. The solution presented 
-by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
+use more GPUs than those hosted in itself, and an idle GPU can not be shared with other nodes. 
+The solution presented by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
 them accessible for any VM in the cluster. However, this kind
@@
 
-Our proposal goes further and, in addition to 
-bringing solutions for all kind of HPC applications, aims 
-to boost flexibility in the use of GPUs.
+Our proposal goes further, not only bringing solutions for all kind of HPC applications, 
+but also aiming to boost flexibility in the use of GPUs.
 
@@
 it. Therefore, the number of VMs that can access the GPUs is 
-limited by the number of the installed devices. This lack of 
+limited by the number of the installed devices. 
+This lack of 
 flexibility is also present in cluster configurations, where 
 depending on the users' necessities, 
-two approaches to enable GPGPU computation can be identified:
-the conservative option integrates many hosts with GPUs to 
-ensure that it is likely to satisfy all requests; in a
-more daring approach, a reduced count of some GPU-enabled nodes is configured, in the hope that a large burst of GPU requests 
- does not exhaust the cluster's resources.
+two approaches to enable GPGPU computation can be identified.
+On the one hand, the conservative option, that integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
+On the other hand, a more daring approach, a reduced count of GPU-enabled nodes is configured, based on an euristhic value calculated from previous experience.
 
@@
 Name & vCPUs & Memory & Network & GPUs & Price\\ \hline \hline
-c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.21\\ \hline
+c3.2xlarge & 8 & 15 GiB & High & 0 & \$ 0.42\\ \hline
 c3.8xlarge & 32 & 60 GiB & 10 Gb & 0 & \$ 1.68 \\ \hline
@@
 we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
-As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, revealing that the accelerator is an actual GPU. 
+As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. 
 % \begin{table}[htb]
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a local GPU.}
+\caption{Results of {\tt bandwidthtest} transferring 32MB using pageable memory in a local GPU.}
 \label{table:bwt}
@@
 \subsection{Testbed Scenarios}
-In this section, we introduce the different client-server configurations. 
+In this section, we introduce different client-server configurations. 
 All scenarios are based on the maximum number of instances 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
+\caption{Results of {\tt bandwidthtest} transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
 \label{table:bwtrcuda}
@@
 because neither the network nor the accelerators are powerful enough to deliver high performance
-when running compute-intensive parallel applications. 
+when running compute-intensive parallel applications. As~\cite{tonithesis} demonstrates, network and device types are critical in the performance.
 In other words, AWS is more oriented toward offering a general-purpose service than to provide high performance.
@@
 
-In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be idle until a GPU-instance request is made.
-Maintaining an expensive accelerator in a standby state is counter-productive. 
-It makes more sense to dedicate less devices, accessible from any machine.
+In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be waiting for a GPU-instance request.
+Investing in expensive accelerators to keep them in a standby state is counter-productive. 
+It makes more sense to dedicate less devices, accessible from any machine, with a higher utilization rate.
 

### 2016-02-02 paper.tex
rcuda section reduced

@@
 cluster nodes from which the actual GPUs provide the requested GPGPU services.
-To support a concurrent scenario where GPUs are shared between
-processes\slash nodes, {rCUDA} manages separate device contexts for
-each client application.
+%To support a concurrent scenario where GPUs are shared between
+%processes\slash nodes, {rCUDA} manages separate device contexts for
+%each client application.
 
 The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
-CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
-APIs (Application Programming Interface) as well as other frequently used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
+CUDA 6.5 release~\cite{cuda65}.%, including the runtime and driver
+%APIs (Application Programming Interface) as well as other frequently used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
 With this middleware, applications are not aware that they are executed

### 2016-02-04 paper.tex
blue reviews

--- 2016-02-02 paper.tex
@@
 \usepackage{flushend}
+
+\usepackage{color}
 
@@
 
+\textcolor{blue}{
 Unfortunately, sharing GPU resources among multiple VMs in cloud environments is more complex than in physical servers.
@@
 As a result, neither vendors nor tools offer GPGPU services.
-
+}
 Remote virtualization has been recently proposed to deal with the low-usage problem. Some relevant examples include {rCUDA}~\cite{tonithesis},
@@
 
+\textcolor{blue}{
 The main goals of this work are to study current cloud solutions in an HPC GPU-enabled scenario, and to analyze and improve 
@@
 In order to reach this goal, we select rCUDA, a virtualization tool that is possibly the more complete and up-to-date for NVIDIA GPUs.
+}
 
@@
 or software requirements, and the ability to integrate with legacy systems and third party technologies. 
-
+\textcolor{blue}{
 From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. 
 Even though in OpenStack it is possible to work with GPUs, the Nova project is not currently aware of them~\cite{OpenStackGPU}. 
+}
 
@@
 none of them is as ambitious as ours. 
-
+\textcolor{blue}{
 The work in~\cite{younge2013enabling} allows the VM managed
@@
 node, but with this solution a node cannot 
-use more GPUs than those hosted in itself, and an idle GPU can not be shared with other nodes. 
+use more GPUs than those hosted in itself, and an idle GPU can not be shared with other nodes. }
 The solution presented by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
@@
 once it has been assigned to an instance, no other VM can use
-it. Therefore, the number of VMs that can access the GPUs is 
+it. 
+\textcolor{blue}{Therefore, the number of VMs that can access the GPUs is 
 limited by the number of the installed devices. 
@@
 On the other hand, a more daring approach, a reduced count of GPU-enabled nodes is configured, based on an euristhic value calculated from previous experience.
+}
 
@@
 we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
-As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. 
+\textcolor{blue}{As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. }
 % \begin{table}[htb]
@@
 \subsection{Discussion}
-The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
+\textcolor{blue}{The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
 because neither the network nor the accelerators are powerful enough to deliver high performance
@@
 Also, AWS fails in offering flexibility as it forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
-
+}
 Table~\ref{table:awsInstances} shows that the resources of the ``g2.8xlarge'' are quadrupled, but so is the cost per hour.
@@
 For example, we can work on an instance with 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs, avoiding to pay the double for features that we do not need in ``g2.8xlarge''.
-
+\textcolor{blue}{
 In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be waiting for a GPU-instance request.
@@
 It makes more sense to dedicate less devices, accessible from any machine, with a higher utilization rate.
-
+}
 \section{\uppercase{GPGPU Shared Service in OpenStack}}

### 2016-02-04 paper.tex
agradecimientos de la UPV

@@
 The authors would like to thank the IT members of the department Gustavo Edo and Vicente Roca for their help.
-
 This research was supported by Universitat Jaume~I research project (P11B2013-21); and project
 TIN2014-53495-R from MINECO and FEDER.
-
+The initial version of rCUDA was jointly developed by Universitat Polit\`ecnica de 
+Val\`encia (UPV) and Universitat Jaume I de Castell\'on (UJI) until year 2010. 
+This initial development was later split into two branches. Part of the 
+UPV version was used in this paper. The development of the UPV branch 
+was supported by Generalitat Valenciana under Grants PROMETEO 2008/060 
+and Prometeo II 2013/009. 
 \bibliographystyle{apalike}

### 2016-02-09 paper.tex
cambios de enrique añadidos

@@
 
-\textcolor{blue}{
+\textcolor{black}{
 Unfortunately, sharing GPU resources among multiple VMs in cloud environments is more complex than in physical servers.
 On one hand, instances in public clouds are not easily customizable. 
-On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
+On the other, although the instances in a private cloud can be customized in many aspects, when referring to GPUs
 the number of options is reduced.
@@
 
-\textcolor{blue}{
+\textcolor{black}{
 The main goals of this work are to study current cloud solutions in an HPC GPU-enabled scenario, and to analyze and improve 
-them by adding flexibility with GPU virtualization. 
+them by adding flexibility via GPU virtualization. 
 In order to reach this goal, we select rCUDA, a virtualization tool that is possibly the more complete and up-to-date for NVIDIA GPUs.
@@
 or software requirements, and the ability to integrate with legacy systems and third party technologies. 
-\textcolor{blue}{
-From the HPC perspective, OpenStack offers high performance configurations with different hardware architectures. 
-Even though in OpenStack it is possible to work with GPUs, the Nova project is not currently aware of them~\cite{OpenStackGPU}. 
+\textcolor{black}{
+From the HPC perspective, OpenStack offers high performance virtual machine configurations with different hardware architectures. 
+Even though in OpenStack it is possible to work with GPUs, the Nova project does not support this architecture yet~\cite{OpenStackGPU}. 
 }
@@
 none of them is as ambitious as ours. 
-\textcolor{blue}{
+\textcolor{black}{
 The work in~\cite{younge2013enabling} allows the VM managed
@@
 node, but with this solution a node cannot 
-use more GPUs than those hosted in itself, and an idle GPU can not be shared with other nodes. }
+use more GPUs than those locally hosted, and an idle GPU cannot be shared with other nodes. }
 The solution presented by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
@@
 it. 
-\textcolor{blue}{Therefore, the number of VMs that can access the GPUs is 
-limited by the number of the installed devices. 
+\textcolor{black}{Therefore, the number of VMs that can access the GPUs is 
+limited by the number of local devices. 
 This lack of 
@@
 depending on the users' necessities, 
-two approaches to enable GPGPU computation can be identified.
-On the one hand, the conservative option, that integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
-On the other hand, a more daring approach, a reduced count of GPU-enabled nodes is configured, based on an euristhic value calculated from previous experience.
+two approaches can be identified to enable GPGPU computation.
+On the one hand, the conservative option integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
+On the other hand, a more daring approach employs a reduced count of GPU-enabled nodes, based on an euristhic value calculated from past experience.
 }
@@
 we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
-\textcolor{blue}{As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. }
+\textcolor{black}{As shown in Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwidth, which suggests that the accelerator is an actual GPU. }
 % \begin{table}[htb]
@@
 \subsection{Discussion}
-\textcolor{blue}{The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
+\textcolor{black}{The previous experiments reveal that the AWS GPU-instances are not appropriate for HPC
 because neither the network nor the accelerators are powerful enough to deliver high performance
-when running compute-intensive parallel applications. As~\cite{tonithesis} demonstrates, network and device types are critical in the performance.
+when running compute-intensive parallel applications. As~\cite{tonithesis} demonstrates, network and device types are critical factors to performance.
 In other words, AWS is more oriented toward offering a general-purpose service than to provide high performance.
-Also, AWS fails in offering flexibility as it forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.
-}
+Also, AWS fails in offering flexibility as it enforces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.}
 Table~\ref{table:awsInstances} shows that the resources of the ``g2.8xlarge'' are quadrupled, but so is the cost per hour.
@@
 For example, we can work on an instance with 2 GPUs for only \$ 1.3 by launching 2 ``g2.2xlarge'' and using remote GPUs, avoiding to pay the double for features that we do not need in ``g2.8xlarge''.
-\textcolor{blue}{
+\textcolor{black}{
 In terms of GPU-shareability, AWS reserves GPU-capable physical machines which will be waiting for a GPU-instance request.
 Investing in expensive accelerators to keep them in a standby state is counter-productive. 
-It makes more sense to dedicate less devices, accessible from any machine, with a higher utilization rate.
+It makes more sense to dedicate less devices, accessible from any machine, resulting in a higher utilization rate.
 }

### 2016-02-09 paper.tex
motivation removed

@@
 but also aiming to boost flexibility in the use of GPUs.
-
-\section{\uppercase{Motivation}}
-\label{sec:motivation}
-
-Cloud computing and storage have gained wide appeal  
-during the last years thanks to a relatively easy configuration  
-process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
-that look for temporary access to state-of-the-art technology without spending a 
-huge amount of money. In general, the systems provided by cloud 
-companies match those requested by the clients thanks to its flexibility.
-
-Two different types of cloud approaches are currently available:
-on one hand, public facilities, such as AWS or 
-Microsoft Azure, that allow users to chose a pre-configured system.   
-On the other hand, institutions/companies 
-are able to deploy their own cloud infrastructure using a software as OpenStack.
-
-In both cases, the power of current cloud facilities relies on the use 
-of VMs which  
-emulate a physical node with its own 
-CPUs, storage, network interconnection, GPU and OS. 
-However, providing a VM equipped with a GPU implies
-that the guest host must be equipped with at least one real GPU, and
-once it has been assigned to an instance, no other VM can use
-it. 
-\textcolor{black}{Therefore, the number of VMs that can access the GPUs is 
-limited by the number of local devices. 
-This lack of 
-flexibility is also present in cluster configurations, where 
-depending on the users' necessities, 
-two approaches can be identified to enable GPGPU computation.
-On the one hand, the conservative option integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
-On the other hand, a more daring approach employs a reduced count of GPU-enabled nodes, based on an euristhic value calculated from past experience.
-}
-
-The use of a virtualization technology, such as {rCUDA}, adds 
-the desired flexibility because this software detaches the GPUs 
-from the nodes where they are installed, allowing all nodes in a cluster
-to use them as if they were local, as well as to share them, 
-achieving a higher utilization rate and reducing the idle periods.
-Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
-environment can leverage and access all GPUs installed in 
-the physical nodes, removing past limitations.
-
-Our work selects AWS as a public 
-solution, and OpenStack as a private cloud environment. 
-In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
-needs to be implemented.
+%
+%\section{\uppercase{Motivation}}
+%\label{sec:motivation}
+%
+%Cloud computing and storage have gained wide appeal  
+%during the last years thanks to a relatively easy configuration  
+%process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
+%that look for temporary access to state-of-the-art technology without spending a 
+%huge amount of money. In general, the systems provided by cloud 
+%companies match those requested by the clients thanks to its flexibility.
+%
+%Two different types of cloud approaches are currently available:
+%on one hand, public facilities, such as AWS or 
+%Microsoft Azure, that allow users to chose a pre-configured system.   
+%On the other hand, institutions/companies 
+%are able to deploy their own cloud infrastructure using a software as OpenStack.  
+%
+%In both cases, the power of current cloud facilities relies on the use 
+%of VMs which  
+%emulate a physical node with its own 
+%CPUs, storage, network interconnection, GPU and OS. 
+%However, providing a VM equipped with a GPU implies
+%that the guest host must be equipped with at least one real GPU, and
+%once it has been assigned to an instance, no other VM can use
+%it. 
+%\textcolor{black}{Therefore, the number of VMs that can access the GPUs is 
+%limited by the number of local devices. 
+%This lack of 
+%flexibility is also present in cluster configurations, where 
+%depending on the users' necessities, 
+%two approaches can be identified to enable GPGPU computation.
+%On the one hand, the conservative option integrates many hosts with GPUs to ensure that it is likely to satisfy all requests.
+%On the other hand, a more daring approach employs a reduced count of GPU-enabled nodes, based on an euristhic value calculated from past experience.
+%}
+%
+%The use of a virtualization technology, such as {rCUDA}, adds 
+%the desired flexibility because this software detaches the GPUs 
+%from the nodes where they are installed, allowing all nodes in a cluster
+%to use them as if they were local, as well as to share them, 
+%achieving a higher utilization rate and reducing the idle periods.
+%Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
+%environment can leverage and access all GPUs installed in 
+%the physical nodes, removing past limitations. 
+%
+%Our work selects AWS as a public 
+%solution, and OpenStack as a private cloud environment. 
+%In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
+%needs to be implemented.
 

### 2016-02-09 paper.tex
8 paginas

@@
 \usepackage{url}% http://ctan.org/pkg/url
-\usepackage{flushend}
+%\usepackage{flushend}
 
@@
 Remote virtualization has been recently proposed to deal with the low-usage problem. Some relevant examples include {rCUDA}~\cite{tonithesis},
-DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}.
-
+DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, 
+vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}.
 Roughly speaking these virtualization frameworks 
@@
 Section~\ref{sec:state} summarizes related work; 
-Section~\ref{sec:motivation} states the motivation to address this task;
+%Section~\ref{sec:motivation} states the motivation to address this task;
 the effort to use AWS is explained in Section~\ref{sec:workingaws};
@@
 as if they were local.
-These features aim to attain higher accelerator utilization rates in the overall system while simultaneously reducing
-resource, space, and energy requirements~\cite{energy14}.
+%These features aim to attain higher accelerator utilization rates in the overall system while simultaneously reducing
+%resource, space, and energy requirements~\cite{energy14}.
 rCUDA is structured following a client-server distributed
-architecture: the client middleware runs in the same cluster node where the application demanding GPGPU
-acceleration services is executed, providing a transparent replacement for the
-native CUDA libraries. Furthermore, the server middleware is executed in the
-cluster nodes from which the actual GPUs provide the requested GPGPU services.
+architecture and its %: the client middleware runs in the same cluster node where the application demanding GPGPU
+%acceleration services is executed, providing a transparent replacement for the
+%native CUDA libraries. Furthermore, the server middleware is executed in the
+%cluster nodes from which the actual GPUs provide the requested GPGPU services.
 %To support a concurrent scenario where GPUs are shared between
@@
 %each client application.
-
-The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
+%The {rCUDA} 5.0 
+client exposes the same interface as the regular NVIDIA
 CUDA 6.5 release~\cite{cuda65}.%, including the runtime and driver
@@
 directive-based models such as OmpSs~\cite{repara15} and OpenACC~\cite{cluster15}.
-
 The integration of remote GPGPU virtualization with global
@@
 technology, making accelerator-enabled clusters more flexible and
-energy-efficient.
+energy-efficient~\cite{energy14}.
 
@@
 and life-cycle management of VMs.
-
  The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
 or software requirements, and the ability to integrate with legacy systems and third party technologies.
+
 \textcolor{black}{
 From the HPC perspective, OpenStack offers high performance virtual machine configurations with different hardware architectures. 
-Even though in OpenStack it is possible to work with GPUs, the Nova project does not support this architecture yet~\cite{OpenStackGPU}. 
+Even though in OpenStack it is possible to work with GPUs, the Nova project does not support this architecture yet.%~\cite{OpenStackGPU}. 
 }
@@
 \label{sec:workingaws}
-In this section, we briefly expose the current user-level features 
-provided by AWS, and then analyze this service  
-from the point of view of GPU usage. 
-Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of initial results.
+%In this section, we briefly expose the current user-level features 
+%provided by AWS, and then analyze this service  
+%from the point of view of GPU usage. 
+%Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of initial results.
 
 \subsection{Current Features}
-This subsection identifies several common features checked during 
-the VM's creation step.
+%This subsection identifies several common features checked during 
+%the VM's creation step.
 
@@
 interconnection, which can be: low, medium, high or 10Gb.
-
 For our study, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
@@
 \subsection{Testbed Scenarios}
-In this section, we introduce different client-server configurations. 
+%In this section, we introduce different client-server configurations. 
 All scenarios are based on the maximum number of instances 
@@
 In particular, the maximum number for ``g2.2xlarge'' is 5; for ``g2.8xlarge'' it is 2.
-
-All instances operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
+Ant the instances operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
 We design three configuration scenarios for our tests:
@@
 \subfigure[5 remote GPUs over 1GbE.]{%
-\includegraphics[width=0.315\linewidth]{images/aws3.pdf}
+\includegraphics[width=0.3\linewidth]{images/aws3.pdf}
 \label{subfig:aws1}}
@@
 \subfigure[8 remote GPUs over 10GbE.]{%
-\includegraphics[width=0.315\linewidth]{images/aws2.pdf}
+\includegraphics[width=0.3\linewidth]{images/aws2.pdf}
 \label{subfig:aws2}}
 \subfigure[13 remote GPUs over 1GbE.]{%
-\includegraphics[width=0.315\linewidth]{images/aws1.pdf}
+\includegraphics[width=0.3\linewidth]{images/aws1.pdf}
 \label{subfig:aws3}}
@@
 \subsection{Experimental results}
-In this subsection we employ two GPU applications to assess the benefits of GPU virtualization.
+%In this subsection we employ two GPU applications to assess the benefits of GPU virtualization.
 
@@
 \label{sec:gpgpuOS}
-In this section we describe our work on OpenStack, from the development of an external module to handle remote GPUs to the physical setup and its validation.
+%In this section we describe our work on OpenStack, from the development of an external module to handle remote GPUs to the physical setup and its validation.
 
@@
 \subsection{Experimental Results}
-The following tests were executed on 2 sets of nodes interconnected by a 1Gb Ethernet network.
-All the nodes were equipped with an Intel Xeon E7420 quadcore processor, at 2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
+Following tests were executed on 2 sets of nodes using a 1Gb Ethernet network.
+All the nodes composed by an Intel Xeon E7420 quadcore processor, at 2.13 GHz, and 16 GB DDR2 RAM at 667 MHz.
 
@@
 The number of available GPU in “shared” mode will depend on the partition size. 
-In this particular case, we halved the GPU memory, resulting 8 partitions that can be addressed as independent GPUs.
+In this case, we halved the GPU memory, resulting 8 partitions that can be addressed as independent GPUs.
 For each group, we deployed virtual clusters of 1, 2 and 3 nodes, where the application processes were executed. 
@@
 Therefore, the number of GPUs determines the maximum number of processes we can launch.
-
 Figure~\ref{fig3} compares the execution time of the application with different configurations over different setups. 
@@
 The most remarkable achievement is the wide range of possible configurations and the flexibility to adapt a system to fit the user requirements. 
-
 In addition, with this virtualization technology, the requests for GPU devices can be fulfilled with small investment in infrastructure and maintenance.
@@
 The first item in the list of pending work is an upgrade of the network to an interconnect that is more prone to HPC.
-In particular, porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions for HPC.
+In particular, porting the setup and the tests to an infrastructure with an Infiniband network will shed light on the viability of this kind of solutions.
 Similar reasons, motivate us to try other Cloud vendors which better support for HPC.
-
-Looking for real situations where performance is less important than flexibility will drive us to explore alternative tools to easily deploy GPU-programming computer labs.
+Looking for situations where performance is less important than flexibility will drive us to explore alternative tools to easily deploy GPU-programming computer labs.
 
@@
 The initial version of rCUDA was jointly developed by Universitat Polit\`ecnica de 
-Val\`encia (UPV) and Universitat Jaume I de Castell\'on (UJI) until year 2010. 
+Val\`encia (UPV) and Universitat Jaume I de Castell\'on until year 2010. 
 This initial development was later split into two branches. Part of the 
-UPV version was used in this paper. The development of the UPV branch 
+UPV version was used in this paper and it  
 was supported by Generalitat Valenciana under Grants PROMETEO 2008/060 
